### 2022-09-22 paper/main.tex
Updated paper draft

\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}

\usepackage{amsfonts}

\title{Tempus: Probabilistic Network Latency Verifier}
\author{WIP}

\begin{document}
\begin{abstract}
    Test
\end{abstract}

\maketitle

\section{Introduction}
As society develops further in the information age and gets more dependent on interconnected 
digital services, the task of making those services reliable is of paramount priority.
The network that underlies those services is no exception to this.

Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
routing information so that a packet can reach their intended destination efficiently, even in the 
event of a failure.
Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
It is hard for a network engineer to reason about whether their configured network fulfilled some 
intended property in various possible states of the network. 
Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
to help them in this task, namely, network verifier.

In recent years, there have been many work done in this area. 
Data plane verifiers work on a single snapshot of the forwarding table of a given network to 
verify a property. 
However, if we want to verify how a given routing protocol will react to a specific failure, 
control plane verifier might be used instead. 
A control plane verifier will efficiently search for failures that might cause change in the 
property in question.

Earlier control plane verifiers are formulated to answer questions about a given property 
deterministically, in other words, yes or no answer. 
For example, k-link failures.
While useful to some extent, this kind of models are sometimes too restrictive since network 
operators are usually able to tolerate failures for a small fraction of time, particularly when 
reasoning about compliance with Service Level Agreement (SLA) where, for example, a given network 
only needs to be operational 99.999\% of the time. 
This kind of probabilistic properties have been the focus of a more recent works
\cite{steffen2020probabilistic}.

The property itself can also be divided into two kinds. 
Quantitative properties of the network, like reachability, waypointing, and loop existence, are 
the common properties that are studied by most of the existing verifiers. 
More recent works, like QARC, has also explored qualitative properties like link bandwidth 
violation for a given traffic.

In this work, we're exploring the other side of the network performance metric: latency.

\section{Proposed Approach}
We will start with the assertion that latency is a property that only make sense 
after connectivity between two nodes has been established. 
In other words, if two nodes in a network aren't connected (due to physical failure 
or ACL policy), then the latency between them will \textit{always} be infinite.
Because of this, we divide the problem of latency verification into two parts: 
verifying that two nodes are reachable (\textit{functional} property) and only if 
the functional property is fulfilled, we would verify whether the latency between 
two nodes fulfill some property (\textit{temporal} property)

\subsection{Topology Graph}
For functional property verification, previous work \cite{steffen2020probabilistic} 
has laid the way for verifying reachability between two nodes under failure in a 
quantifiable and efficient manner. 
In this framework, the physical network is encoded in an edge-labeled directed graph 
$G_T = (V_T, E_T)$ where $V_T$ represents the routers in the network and 
$E_T$ represents the connectivity between a pair of source and destination router. 
The function $r: E_T \rightarrow \{x \in \mathbb{R} \mid 0 \le x \le 1\}$ is the 
edge-label that represents the failure rate of a given connectivity link.

On top of this topology, we could define additional information that would be used
by a routing protocol to determine the valid path(s) between two routers 
$src, dst \in V_T$ given a particular link failure scenario. In OSPF for example, 
we define a function $w_{ospf}: E_T \rightarrow \mathbb{N}$ as the edge-label that 
represents the positive weight of a link.

\subsection{Latency Graph}
There are a lot of components in a network that may introduce some latency into 
a transmitting packet. 
The principal challenge for us, then, is to determine the appropriate level of abstraction; 
detailed enough in order to accomplish our verification goal in a correct and accurate manner but 
abstract enough to not introduce scalability problems. 
We settle on modeling two source of latency that we deem significant: propagation delay and 
queuing delay.

Propagation delay is the latency that is introduced by the links in the network. 
This delay is independent of the load in the system, i.e. no matter how many in-flight packets 
are transmitting, the introduced delay is relatively the same. 
Queuing delay is the latency that the queuing process in the node introduced. 
Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
packets there are in the queue, the more delay the node will introduce to a subsequent packets.

To take both of these forms of delay into account, we've decided to model the latency of each
components with a univariate continuous random variable.
For propagation delay, the semantic of this random variable is relatively straightforward: it is 
the distribution of latency that a given link will introduce.
For queuing delay however, this random variable represents the delay that a given queuing process 
will introduce, marginalized over various traffic pattern that a given network state might have 
resulted.

% TODO: change this to v_i
To convey this additional notion of latency distribution, we form a second edge-labeled directed 
graph (called a latency graph) $G_L = (V_L, E_L)$.
For each network node $v \in V_T$, we want to ``expand'' this node to be able to represent the 
inner working of the queuing process within each node. 
Each $v$ will be expanded into $v_{in}, v_{out1}, v_{out2}, ..., v_{outn} \in V_L$ where $n$ is 
the amount of outward neighbors that $v$ has in $G_T$. 
Every inward edge that $v$ has will get directed to $v_{in}$.
$v_{out}$ is used to represent each physical port that $v$ has and thus, each outward edge that 
$v$ has will be sourced from one of the $v_{out}$ instead.
We can then add additional edges $e_1, e_2, ..., e_n$ that connects from $v_{in}$ to each 
$v_{out}$.
While each $v_{out}$ represents a port, these edges represent the output queue connected to those 
ports.
Since we already convey the routing and failure information in $G_T$, we instead label each 
$e \in E_L$ with the random variable distribution that each of these components might introduce.

\subsection{Latency Verification}
From these two graphs, we then do the verification in the following way:

First, we will do the functional reachability verification similar to what NetDice have done. 
The difference, however, is that we also collect additional information from each state (e.g. 
convergent paths) in order to be used in the temporal verification step. 
After the functional verifier finished exploring all states / stopped at a given threshold, we 
then move to the next step.

Second, we will compute the total path latency distribution for each convergent path in the state.
This is done by doing the convolution operation over the latency distribution of each component 
in the path. 
Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
method for mixture distribution, DIRECT, which is able to convolve two distribution with a 
KL divergence error bound.

Third, after we compute the total path latency distribution, we calculate the probability of a 
given latency property being true in that state. For example, bounded reachability property 
(whether a packet can traverse from a source to destination below some time unit $T$) can be 
computed by integrating the PDF from $0$ to $T$. 
We combine all of the temporal probability of each state with the functional probability we 
calculated in the first step to get the probability of a given temporal property being true, 
which is what we want.

\section{Proposed Evaluation}
\begin{itemize}
    \item cross-validation with NetDice and DIRECT implementation in R
    \item performance evaluation with topology-zoo + dummy latency distribution
    \item performance evaluation with latency measurement data
\end{itemize}

\bibliography{cit}
\bibliographystyle{ieeetr}
\end{document}

### 2022-09-22 src/Direct.jl
Updated code

@@ -1,0 +1,59 @@
+# KL-divergence integrand of one distribution with 2 different inputs
+function kldivergence_integrand(d::Distribution, x1::Float64, x2::Float64)::Float64
+    px = pdf(d, x1)
+    log_px = logpdf(d, x1)
+    log_qx = logpdf(d, x2)
+
+    if log_px > -Inf && log_qx > -Inf
+        return px * (log_px - log_qx)
+    else
+        return 0.0
+    end
+end
+
+# Symmetric KL-divergence between one distribution with a shifted input
+function sym_kldivergence_shift(d::Distribution, shift::Float64)::Float64
+    # forward KL-divergence
+    forward = first(quadgk(x -> kldivergence_integrand(d, x, x-shift), extrema(d)...))
+    # backward KL-divergence
+    backward = first(quadgk(x -> kldivergence_integrand(d, x-shift, x), extrema(d)...))
+
+    return forward + backward
+end
+
+function direct(d1::UnivariateDistribution, d2::UnivariateDistribution, delta::Float64 = 0.01, epsilon::Float64 = 0.0001)
+    # Determine bin half-width from delta
+    step = sqrt(delta)
+    while sym_kldivergence_shift(d2, step) < delta
+        step = step * 2
+    end
+    step = find_zero(x -> sym_kldivergence_shift(d2, x) - delta, (0, step))
+
+    # Determine grid range from epsilon
+    mini = -1.0
+    maxi = 1.0
+    while cdf(d1, mini) > epsilon / 2
+        mini = mini * 2
+    end
+    while cdf(d1, maxi) < 1 - (epsilon / 2)
+        maxi = maxi * 2
+    end
+    mini = find_zero(x -> cdf(d1, x) - (epsilon / 2), (mini, maxi))
+    maxi = find_zero(x -> cdf(d1, x) - (1 - epsilon / 2), (mini, maxi))
+
+    # Determine reference points
+    k = ceil(Int, (maxi - mini)/(2 * step)) + 1
+    supports = [mini - ((2 * k * step) - (maxi - mini)) / 2 + 2 * x * step for x in 0:(k-1)]
+    
+    # Determine bin weights
+    margins = supports[2:length(supports)] .- step
+    weights = [cdf(d1, margins[i]) for i in 1:(k-1)]
+    weights = [weights; 1.0]
+    for i in k:-1:2
+        weights[i] = weights[i] - weights[i - 1]
+    end
+
[…]

### 2022-09-22 src/DirectDistributions.jl

@@ -1,0 +1,26 @@
+struct DirectDistribution{T<:Real} <: ContinuousUnivariateDistribution
+    supports::Vector{T}
+    weights::Vector{T}
+    d2::Distribution
+
+    delta::T
+    epsilon::T
+end
+
+cdf(d::DirectDistribution, x::Real) = sum(d.weights .* [cdf(d.d2, x - support) for support in d.supports])
+pdf(d::DirectDistribution, x::Real) = sum(d.weights .* [pdf(d.d2, x - support) for support in d.supports])
+logpdf(d::DirectDistribution, x::Real) = log(pdf(d, x))
+maximum(d::DirectDistribution) = Inf
+minimum(d::DirectDistribution) = -Inf
+
+function quantile(d::DirectDistribution, q::Real)
+    mini = -1.0
+    maxi = 1.0
+    while cdf(d, mini) > q
+        mini = mini * 2
+    end
+    while cdf(d, maxi) < q
+        maxi = maxi * 2
+    end
+    return find_zero(x -> cdf(d, x) - q, (mini, maxi))
+end

### 2022-09-22 src/ModifiedYen.jl

@@ -252,4 +252,8 @@ function allpaths_to_unique_links(allpaths::Vector{Vector{Symbol}})::Vector{Tupl
     return unique(links)
 end
 
+function get_hot_edges()
+    # Compute convergent paths 
+end
+
 end # module

### 2022-09-22 src/Tempus.jl

@@ -1,29 +1,38 @@
 module Tempus
 
-include("ConfigParser.jl")
-# include("Paths.jl")
-# include("ProbTraversal.jl")
-# using MetaGraphsNext
-# using Graphs
-include("ModifiedYen.jl")
+import JSON
+using Distributions
+import Distributions: cdf, pdf, logpdf, minimum, maximum, quantile, convolve
+
+using Graphs
+using MetaGraphsNext
+
+using QuadGK # Numerical integration
+using Roots
+
+include("graphs/TopologyGraphs.jl")
+
+include("DirectDistributions.jl")
+include("Direct.jl")
 
 # don't use 'using' or 'import' so that all globals are compartmentalized?
 
 function main()
-    topgraph, delaygraph, fwdtable, intent = ConfigParser.parseconf("artifacts/Highwinds.json")
-    a = ModifiedYen.tempus_yen_stateful(topgraph, Symbol(intent["src"]), Symbol(intent["dst"]))
-    println(a)
-    # dj = dijkstra_shortest_paths(topgraph, code_for(topgraph, Symbol(intent["src"])), weights(topgraph), allpaths=true)
-    # paths = enumerate_paths(dj, [code_for(topgraph, Symbol(intent["dst"])), code_for(topgraph, Symbol(intent["dst"]))])
-    # println(paths)
-    # prob = Paths.modified_yen(topgraph.graph, code_for(topgraph, Symbol(intent["src"])), code_for(topgraph, Symbol(intent["dst"])))
-    # println(ne(topgraph))
-    # reduciblepaths = Paths.get_reducible_paths_yen(topgraph, Symbol(intent["src"]), Symbol(intent["dst"]), 1000)
-    # reduciblepaths = Paths.get_reducible_paths_yen(delaygraph, Symbol(intent["src"] * "_in"), Symbol(intent["dst"] * "_out"), 500)
-    # println(size(reduciblepaths))
-    # ProbTraversal.traversestatic(delaygraph, reduciblepaths, intent)
-    # prob = ProbTraversal.traverse_topology_first(topgraph, intent)
-    # println(prob)
+    json::String = open("artifacts/7nodenew.json") do file
+        read(file, String)
+    end
+    config::Dict{String, Any} = JSON.parse(json)
+
+    tg::TopologyGraph = TopologyGraph(config["routers"], config["links"])
+    println(tg)
[…]

### 2022-09-22 src/graphs/TopologyGraphs.jl

@@ -1,0 +1,23 @@
+# EdgeData = (failProb, weight)
+tgweightfunction = x -> last(x)
+const TopologyGraph = typeof(MetaGraph(DiGraph(), EdgeData = Tuple{Real, UInt}, weight_function = tgweightfunction))
+
+function TopologyGraph(routers::Vector{Any}, links::Vector{Any})::TopologyGraph
+    tg::TopologyGraph = MetaGraph(DiGraph(), EdgeData = Tuple{Real, UInt}, weight_function = tgweightfunction)
+
+    # Make the nodes
+    for router::Dict{String, Any} in routers
+        tg[Symbol(router["name"])] = nothing
+    end
+
+    # Make the links
+    for link::Dict{String, Any} in links
+        if link["u"] == "src" || link["v"] == "src" || link["u"] == "dst" || link["v"] == "dst"
+            continue
+        end
+        tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], link["w_uv"])
+        tg[Symbol(link["v"]), Symbol(link["u"])] = (link["failProb"], link["w_vu"])
+    end
+
+    return tg
+end

### 2022-09-23 paper/main.tex
Added abstracts

@@
 \begin{abstract}
-    Test
+    To combat problems and bugs that a given network might have, network verifiers have emerged as 
+    one of the promising solutions. 
+    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
+    various scenarios of network failures, such as reachability under k-link failure. 
+    However, as modern networks evolved and performance need becomes more stringent -- often 
+    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
+    verifiers to also reason about quantitative performance properties. 
+    Works in quantitative network verifiers that has arose in recent years mainly focuses on one 
+    side of the network performance metric: bandwidth and load violation properties. 
+    Questions about the other side of network performance metric, latency, were left unanswered.
+
+    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
+    a given temporal properties being true given latency distributions of individual links and 
+    nodes in the network. 
+    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
+    that the average delay of packets traversing from a source to destination node is below $T$
+    time unit -- by only adding a fraction of the overhead introduced by the qualitative verifier 
+    its built upon.
 \end{abstract}
@@
 \section{Introduction}
-As society develops further in the information age and gets more dependent on interconnected 
-digital services, the task of making those services reliable is of paramount priority.
-The network that underlies those services is no exception to this.
-
 Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
@@
 Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network verifier.
+to help them in this task, namely, network control plane verifier.
 
-In recent years, there have been many work done in this area. 
-Data plane verifiers work on a single snapshot of the forwarding table of a given network to 
-verify a property. 
-However, if we want to verify how a given routing protocol will react to a specific failure, 
-control plane verifier might be used instead. 
-A control plane verifier will efficiently search for failures that might cause change in the 
-property in question.
-
-Earlier control plane verifiers are formulated to answer questions about a given property 
-deterministically, in other words, yes or no answer. 
-For example, k-link failures.
+Earlier control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
+questions about a given property deterministically. 
+In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
+expected to give a yes or no answer about whether a given property (e.g. two nodes are reachable) 
+is satisfied in all of those cases. 
 While useful to some extent, this kind of models are sometimes too restrictive since network 
@@
 only needs to be operational 99.999\% of the time. 
-This kind of probabilistic properties have been the focus of a more recent works
+This kind of probabilistic properties have been the focus of a more recent works like NetDice 
 \cite{steffen2020probabilistic}.
 
-The property itself can also be divided into two kinds. 
+Aside from the state exploration, the property itself can also be divided into two kinds. 
 Quantitative properties of the network, like reachability, waypointing, and loop existence, are 
 the common properties that are studied by most of the existing verifiers. 
-More recent works, like QARC, has also explored qualitative properties like link bandwidth 
-violation for a given traffic.
+More recent works, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
+properties like link bandwidth violation for a given traffic.
 
@@
 the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfill some property (\textit{temporal} property)
+two nodes fulfilled some property (\textit{temporal} property)
 
@@
 Third, after we compute the total path latency distribution, we calculate the probability of a 
-given latency property being true in that state. For example, bounded reachability property 
-(whether a packet can traverse from a source to destination below some time unit $T$) can be 
-computed by integrating the PDF from $0$ to $T$. 
+given latency property being true in that state. 
+For example, bounded reachability property (whether a packet can traverse from a source to 
+destination below some time unit $T$) can be computed by integrating the PDF from $0$ to $T$. 
 We combine all of the temporal probability of each state with the functional probability we 
@@
 
+% TODO: intro, reformalize graph formulation, citations
+% Examples of workshop paper?
+% Does a figure about topology graph and latency graph will be useful?
+% Multimedia context on older networked system Klara Nahrstedts
+% Graph
+
 \section{Proposed Evaluation}
-\begin{itemize}
-    \item cross-validation with NetDice and DIRECT implementation in R
-    \item performance evaluation with topology-zoo + dummy latency distribution
-    \item performance evaluation with latency measurement data
-\end{itemize}
+We want to evaluate two things: correctness and performance.
+
+On the correctness side, we want to cross-validate the result of the components that make up 
+the verifier, mainly the DIRECT algorithm, with the result of their publication. 
+We also wanted to validate whether the result of our verifier is equivalent to another verifier 
+that has lower fidelity (i.e. NetDice, that only verifies functional property) if we set the 
+temporal property to be virtually unconstrained (e.g. bounded reachability with a very high 
+$T$).
+
+As for performance, we will evaluate the verifier by measuring the amount of additional 
+overhead that the temporal verifier would introduce, measured by the amount of states that 
+needs to be temporally explored and / or the amount of convolution operation that needs to be 
+performed.
+We will evaluate them using two main datasets: using topologies from the topology zoo \cite{knight2011internet} with a
+realistic assumed temporal distribution and topologies that has node measurements that could 
+be used to form a output queue distribution.
 

### 2022-09-24 paper/main.tex
[Paper] Added citations and formatting

@@
 
-Earlier control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
+Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
 questions about a given property deterministically. 
 In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about whether a given property (e.g. two nodes are reachable) 
-is satisfied in all of those cases. 
+expected to give a yes or no answer about the property (e.g. two nodes are reachable).
 While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate failures for a small fraction of time, particularly when 
-reasoning about compliance with Service Level Agreement (SLA) where, for example, a given network 
-only needs to be operational 99.999\% of the time. 
+operators are usually able to tolerate small fraction of failure. 
+For example, a given network might have an availability SLA of 99.999\%. 
 This kind of probabilistic properties have been the focus of a more recent works like NetDice 
@@
 
-Aside from the state exploration, the property itself can also be divided into two kinds. 
-Quantitative properties of the network, like reachability, waypointing, and loop existence, are 
-the common properties that are studied by most of the existing verifiers. 
-More recent works, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
+The network property itself can also be divided into two kinds. 
+Quantitative properties like reachability and loop existence, are the common properties that are 
+studied by most of the existing verifiers. 
+More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
 properties like link bandwidth violation for a given traffic.
@@
 In this work, we're exploring the other side of the network performance metric: latency. 
+Certain network deployment often necessitates some latency requirement such as an ISP that has 
+latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
+We proposed a verification framework to probabilistically verify the latency property of packets 
+traversing from a source to destination node under various failure scenarios, by using latency 
+information of each components in the network.
 
@@
 the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some property (\textit{temporal} property)
+two nodes fulfilled some additional condition (\textit{temporal} property)
 
 \subsection{Topology Graph}
-For functional property verification, previous work \cite{steffen2020probabilistic} 
+For functional property verification, NetDice \cite{steffen2020probabilistic} 
 has laid the way for verifying reachability between two nodes under failure in a 
-quantifiable and efficient manner. 
+probabilistic and efficient manner. 
 In this framework, the physical network is encoded in an edge-labeled directed graph 
-$G_T = (V_T, E_T)$ where $V_T$ represents the routers in the network and 
-$E_T$ represents the connectivity between a pair of source and destination router. 
-The function $r: E_T \rightarrow \{x \in \mathbb{R} \mid 0 \le x \le 1\}$ is the 
+$G_t = (V_t, E_t)$ where $V_t$ represents the routers in the network and 
+$E_t$ represents the connectivity between a pair of source and destination router. 
+The function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x \le 1\}$ is the 
 edge-label that represents the failure rate of a given connectivity link.
@@
 On top of this topology, we could define additional information that would be used
-by a routing protocol to determine the valid path(s) between two routers 
-$src, dst \in V_T$ given a particular link failure scenario. In OSPF for example, 
-we define a function $w_{ospf}: E_T \rightarrow \mathbb{N}$ as the edge-label that 
-represents the positive weight of a link.
+by a routing protocol to determine the valid path(s) between two nodes 
+$src, dst \in V_t$ given a particular link failure scenario. 
+In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
+that represents the positive weight of a link.
 
@@
 detailed enough in order to accomplish our verification goal in a correct and accurate manner but 
-abstract enough to not introduce scalability problems. 
-We settle on modeling two source of latency that we deem significant: propagation delay and 
+abstract enough for the problem to remain tractable. 
+We settle on modeling two sources of latency that we deem significant: propagation delay and 
 queuing delay.
 
-Propagation delay is the latency that is introduced by the links in the network. 
-This delay is independent of the load in the system, i.e. no matter how many in-flight packets 
-are transmitting, the introduced delay is relatively the same. 
+Propagation delay is the latency that is introduced by the links in the network, which is 
+independent of the traffic load in the system.
 Queuing delay is the latency that the queuing process in the node introduced. 
@@
 To convey this additional notion of latency distribution, we form a second edge-labeled directed 
-graph (called a latency graph) $G_L = (V_L, E_L)$.
-For each network node $v \in V_T$, we want to ``expand'' this node to be able to represent the 
+graph (called a latency graph) $G_l = (V_l, E_l)$.
+For each network node $v_i \in V_t$, we want to ``expand'' this node to be able to represent the 
 inner working of the queuing process within each node.
-Each $v$ will be expanded into $v_{in}, v_{out1}, v_{out2}, ..., v_{outn} \in V_L$ where $n$ is 
-the amount of outward neighbors that $v$ has in $G_T$. 
-Every inward edge that $v$ has will get directed to $v_{in}$.
-$v_{out}$ is used to represent each physical port that $v$ has and thus, each outward edge that 
-$v$ has will be sourced from one of the $v_{out}$ instead.
-We can then add additional edges $e_1, e_2, ..., e_n$ that connects from $v_{in}$ to each 
-$v_{out}$.
-While each $v_{out}$ represents a port, these edges represent the output queue connected to those 
-ports.
-Since we already convey the routing and failure information in $G_T$, we instead label each 
-$e \in E_L$ with the random variable distribution that each of these components might introduce.
+
+\begin{itemize}
+    \item Each $v_i$ is expanded into $v_i^{in}, v_i^{out-1}, ..., v_i^{out-n} \in V_l$ 
+    where $n$ is the amount of outward neighbors that $v_i$ has in $G_t$.
+
+    \item Every inward edge that $v_i$ has in $G_t$ will all get directed to $v_i^{in}$.
+
+    \item Every outward edge that $v_i$ has in $G_t$ will be sourced from one of the 
+    $v_i^{out-j}$ instead.
+
+    \item Add additional edges $e_i^j$ that connects from $v_i^{in}$ to each $v_i^{out-j}$.
+
+    \item Finally, we label each $e \in E_l$ with the random variable as the latency distribution.
+\end{itemize}
+
+Semantically, $v_i^{out-j}$ represents a port, normal links that are copied from $G_t$ represents 
+a link with a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
+in node $v_i$ with a queuing delay.
 
@@
 
-First, we will do the functional reachability verification similar to what NetDice have done. 
-The difference, however, is that we also collect additional information from each state (e.g. 
-convergent paths) in order to be used in the temporal verification step. 
-After the functional verifier finished exploring all states / stopped at a given threshold, we 
-then move to the next step.
+\textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
+\cite{steffen2020probabilistic} have done. 
+In addition to that, we also collect additional information from each state (e.g. convergent 
+paths) to be used in the temporal verification step. 
+After the functional verifier finished exploring, we then move to the next step.
 
-Second, we will compute the total path latency distribution for each convergent path in the state.
-This is done by doing the convolution operation over the latency distribution of each component 
-in the path. 
+\textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
+state with $G_l$.
+This is done by convolving over the latency distribution of each component in the path. 
 Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-method for mixture distribution, DIRECT, which is able to convolve two distribution with a 
-KL divergence error bound.
+method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
+distribution with a KL divergence error bound.
 
-Third, after we compute the total path latency distribution, we calculate the probability of a 
-given latency property being true in that state. 
-For example, bounded reachability property (whether a packet can traverse from a source to 
-destination below some time unit $T$) can be computed by integrating the PDF from $0$ to $T$. 
-We combine all of the temporal probability of each state with the functional probability we 
-calculated in the first step to get the probability of a given temporal property being true, 
-which is what we want.
+\textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
+For example, we define \textbf{bounded reachability property} (probability of whether a packet can 
+traverse from a source to destination below some time unit $T$) which can be computed by 
+integrating the PDF from $0$ to $T$. 
+We then combine all of the \textit{temporal} and \textit{functional} in each state to get the 
+probability of a given temporal property being true in this network.
 
 % TODO: intro, reformalize graph formulation, citations
-% Examples of workshop paper?
 % Does a figure about topology graph and latency graph will be useful?
@@
 \section{Proposed Evaluation}
-We want to evaluate two things: correctness and performance.
+On the correctness side, we want to validate whether the result of our verifier is equivalent to 
+another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
+property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
 
-On the correctness side, we want to cross-validate the result of the components that make up 
-the verifier, mainly the DIRECT algorithm, with the result of their publication. 
-We also wanted to validate whether the result of our verifier is equivalent to another verifier 
-that has lower fidelity (i.e. NetDice, that only verifies functional property) if we set the 
-temporal property to be virtually unconstrained (e.g. bounded reachability with a very high 
-$T$).
-
-As for performance, we will evaluate the verifier by measuring the amount of additional 
+On the performance side, we will evaluate the verifier by measuring the amount of additional 
 overhead that the temporal verifier would introduce, measured by the amount of states that 
-needs to be temporally explored and / or the amount of convolution operation that needs to be 
+needs to be re-explored (in the second step) and / or the amount of convolution operation is 
 performed.
-We will evaluate them using two main datasets: using topologies from the topology zoo \cite{knight2011internet} with a
-realistic assumed temporal distribution and topologies that has node measurements that could 
-be used to form a output queue distribution.
+% We will use two main datasets: using topologies from the topology zoo \cite{knight2011internet} 
+% with a realistic assumed temporal distribution and topologies that has node measurements that 
+% could be used to form a output queue distribution.
 

### 2022-10-06 paper/main.tex
[Paper] expanded the verification part

@@
     verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focuses on one 
+    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
     side of the network performance metric: bandwidth and load violation properties. 
@@
     In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
-    a given temporal properties being true given latency distributions of individual links and 
+    a given temporal property being true given latency distributions of individual links and 
     nodes in the network. 
@@
     that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the overhead introduced by the qualitative verifier 
-    its built upon.
+    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
+    qualitative verifier it's built upon.
 \end{abstract}
@@
 routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a failure.
+event of a component failure.
 Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
@@
 
-\section{Proposed Approach}
+\section{Encoding}
 We will start with the assertion that latency is a property that only make sense 
@@
 \subsection{Latency Graph}
-There are a lot of components in a network that may introduce some latency into 
-a transmitting packet. 
+There are many source of latency in a network. 
 The principal challenge for us, then, is to determine the appropriate level of abstraction; 
-detailed enough in order to accomplish our verification goal in a correct and accurate manner but 
+detailed enough in order to accomplish our verification goal correctly but 
 abstract enough for the problem to remain tractable. 
-We settle on modeling two sources of latency that we deem significant: propagation delay and 
+We settle on modeling two significant sources of latency: propagation and 
 queuing delay.
@@
     \item Each $v_i$ is expanded into $v_i^{in}, v_i^{out-1}, ..., v_i^{out-n} \in V_l$ 
-    where $n$ is the amount of outward neighbors that $v_i$ has in $G_t$.
+    where $n$ is the amount $v_i$'s outward neighbors.
 
@@
 Semantically, $v_i^{out-j}$ represents a port, normal links that are copied from $G_t$ represents 
-a link with a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
-in node $v_i$ with a queuing delay.
+a link that has a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
+in node $v_i$ that has a queuing delay.
 
-\subsection{Latency Verification}
+\section{Verification}
 From these two graphs, we then do the verification in the following way:
 
-\textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-\cite{steffen2020probabilistic} have done. 
-In addition to that, we also collect additional information from each state (e.g. convergent 
-paths) to be used in the temporal verification step. 
-After the functional verifier finished exploring, we then move to the next step.
+\subsubsection{Functional Verification}
+Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of links that are 
+alive / up and $D$ is a set of links that are dead / down.
+We then define a set $S$ that represents all possible network states in a given topology. 
+Our goal is to identify a subset of $S$ where each element is a network state that can connect 
+the source and destination node given a convergent behavior of a routing protocol.
 
-\textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-state with $G_l$.
-This is done by convolving over the latency distribution of each component in the path. 
-Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-distribution with a KL divergence error bound.
+This framework fits nicely with the framework of probability theory, where $S$ represents the 
+sample space and the power set of $S$ represents the possible events, one of which is the one
+we care about: the event that a packet from a source node can reach a destination node.
+The way we assign probability to events is to simply do a summation over the product of all the 
+failure and success rate of each links in the network state.
 
-\textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-For example, we define \textbf{bounded reachability property} (probability of whether a packet can 
-traverse from a source to destination below some time unit $T$) which can be computed by 
-integrating the PDF from $0$ to $T$. 
-We then combine all of the \textit{temporal} and \textit{functional} in each state to get the 
-probability of a given temporal property being true in this network.
+The straightforward and naive way to compute the probability of this event is to iterate over all 
+the network state. 
+We then determine whether that network state can fulfill said property and which path it will 
+take. 
+If yes, then we would compute the probability from the product of the link's failure and success 
+rate, and add it to the running sum.
+If no, then we could skip the computation altogether.
+
+While natural, this brute-force algorithm is obviously doesn't scale since it will take $2^n$ 
+iteration ($n$ as the number of links in the topology). 
+NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could skip over some 
+events that are guaranteed not to change the convergent paths of the current network state, by 
+marginalizing over \textit{cold edges}.
+
+In NetDice, we start from a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$).
+We then determine whether that network state can fulfill said property and which path it will 
+take, similar to the brute-force method.
+If no, then it is safe to say that the source node can never reach the destination node in all 
+network state.
+If yes, then we could use the logic of the routing protocol to determine the set of cold edges, 
+(i.e. edges whose failure won't change the convergent paths). 
+Edges that are not part of this set is called \textit{hot edges} $h_i = (U, D)$.
+Instead of using $s_i$ to compute the probability of the current state, NetDice only use the set 
+of hot edges to compute the probability of them being alive ($h_i = (U, D)$ where $D = \emptyset$).
+This in effect computes the probability of multiple network state at once.
+
+Next, instead of examining other network state at random, we instead iterate over the links in 
+$h_i$.
+By definition, if we fail a link that belongs to $h_i$, then the convergent paths will change.
+To cover all of those possibilities efficiently, NetDice iterate over all the links in $h_i$, 
+disabled them, re-enabled them, and move to the next edge to do the same.
+If $h_i = ((e_1, e_2, ..., e_j), \emptyset)$, then we would enqueue $(\emptyset, (e_1))$, 
+$((e_1), (e_2))$, $((e_1, e_2), (e_3))$, ..., $((e_1, e_2, ..., e_{j-1}), (e_j))$.
+Thus, instead of exploring $2^j$ events, we will enqueue $j$ events instead.
+
+NetDice's exploration will effectively form a tree, where the root is the event where we have 
+empty hot edges, and the leaves are where the hot edges are full.
+
+\subsubsection{Temporal Verification}
+Since we assign each links and output queue in the network with a certain latency distribution, 
+computing the probability of the temporal property involves convolving the the paths we got from 
+functional verification and measuring the statistical property of the resulting distribution.
+However, while NetDice has shrunk the exploration space compared to brute-force iteration, for
+the purpose of temporal verification, further optimization can be done.
+
+The first is what we call \textbf{hot-edges union}. 
+The idea is, while each hot-edges events has the same convergent paths, it is not guaranteed to be 
+unique. 
+Two or more of these events can have the same convergent paths.
+Therefore, we could merge these two events and compute the convolution for a given path once.
+
+% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
+% \cite{steffen2020probabilistic} have done. 
+% In addition, we also collect additional information from each state (e.g. convergent 
+% paths) to be used in the temporal verification step.
+
+% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
+% state with $G_l$.
+% This is done by convolving over the latency distribution of each component in the path. 
+% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
+% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
+% distribution with a KL divergence error bound.
+% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
+% explored}.
+
+% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
+% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
+% traverse from a source to destination below some time unit $T$), we can compute its probability by 
+% integrating the PDF from $0$ to $T$. 
+% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
+% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
 
@@
 overhead that the temporal verifier would introduce, measured by the amount of states that 
-needs to be re-explored (in the second step) and / or the amount of convolution operation is 
-performed.
-% We will use two main datasets: using topologies from the topology zoo \cite{knight2011internet} 
-% with a realistic assumed temporal distribution and topologies that has node measurements that 
-% could be used to form a output queue distribution.
+need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
+Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
+produced by step 1, only 459 got re-explored on step 2.
 

### 2022-10-17 paper/main.tex
[Paper] Added 9 paragraphs of section 3

@@
 information of each components in the network.
+
+\section{Overview}
+To be written
+
+\section{On Latency Modeling}
+\subsection{Latency as Path Property} 
+We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
+a single packet.
+In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
+through the many components of the network (e.g. links, routers, firewall) and each of those 
+components might introduce some latency into the packet transmission process.
+Figuring out which of those components will actually introduce latency into the packet in question, 
+and by how much, is the next logical step that we must figure out.
+
+Obviously, a given packet doesn't need to visit all network components to reach its destination 
+host.
+A network engineer will configure the network in such a way that a packet will only need to be 
+routed via a specific subset of its components.
+That specific subset is dictated by two things: how those components are connected together 
+(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
+protocol, ACL).
+Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
+given packet appropriately.
+
+The goal of a classical control plane verifier then, is to use this forwarding table in some 
+form to verify certain properties. %TODO: examples?
+However, looking at the forwarding table alone will not give us a conclusive result regarding 
+which components are going to be visited by a given packet, making verification of latency 
+properties less clear.
+For example, the network might be configured with a load-balancing protocol in which a packet 
+departing from a source host might take multiple different \textit{paths} (with certain 
+probability weights) to arrive at the destination host, possibly resulting in a different 
+end-to-end latency measurement.
+
+Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
+the primary unit of reasoning, rather than forwarding table.
+By being more specific about our unit of reasoning, we could answer verification questions more 
+clearly, and we could design our verifier more efficiently since we could use it to represent 
+multiple different forwarding tables that shares the same path.
+
+\subsection{Relation to Classical Verifier}
+Before we analyze the latency of a given packet that propagates through a certain network 
+path however, we must make sure that that path exists in the first place.
+We note that latency is a property that only make sense after connectivity between two hosts 
+has been established. 
+In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
+link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
+making the verification task trivial.
+
+Fortunately, there are a rich body of work in the network verification literature regarding 
+functional reachability under failure. %TODO: cite
+We could then design our verifier on top of an existing classical control plane verifier.
+We use it to verify reachability property, and only if the reachability property is fulfilled, 
+we would verify whether the latency between two hosts fulfilled some additional condition.
+
+\subsection{Latency Distribution}
+Up to this point, we have talked about measuring the latency of a single packet by figuring out 
+its path; analyzing which exact components it has traversed through.
+However, when we try to generalize this framework and ask questions about the latency of multiple 
+packets, it is apparent that the path alone is not a determinative information, as the latency 
+of two packets propagating through the same path might be different due to a multitude of factors.
+
+The natural extension to the framework then, is instead of representing latency of a path with a 
+single number, we instead represent it with a continuous random variable that signifies the 
+possible delay that a given packet traversing through that path might have.
+This random variable will have a distribution that marginalizes over all other factors other than 
+the path.
+
+We can then use this latency distribution to verify some temporal properties in a probabilistic 
+manner.
+For example, we could verify the probability that a packet will be delivered in under a time 
+unit by taking the CDF of the distribution.
+
+\subsection{Path Decomposition and Convolution}
+Expand on these points:
+\begin{itemize}
+    \item How do we get the path latency distribution? Component-wise measurement data
+    \item What measurement? link delay and queuing delay $\rightarrow$ not all delays are 
+    significant
+    \item What to do with it? Simulation $\rightarrow$ lack of error guarantee
+    \item Convolution
+    \item Independence?
+\end{itemize}
 

### 2022-10-17 src/Exploration.jl
[Code] reorganized modules

@@ -1,0 +1,169 @@
+mutable struct MetaPath
+    path::Vector{Symbol}
+    dependencies::Vector{Symbol}        # TODO: technically should be a set
+end
+
+function explore(tg::AbstractGraph, src::Symbol, dst::Symbol)::Float64
+    # If src and dst are the same node
+    src == dst && return 1.0
+
+    # Create state tree and exploration queue
+    st = StateTree()
+    e = PriorityQueue(Base.Order.Reverse)
+
+    # Create the perfect state as the initial state
+    perfect = State([], [])
+    st[:perfect] = perfect
+    st[:perfect].p_state = get_p_state(st, :perfect)
+    e[:perfect] = st[:perfect].p_state
+    
+    # Prepare for functional state exploration
+    tgcopy = deepcopy(tg)
+    p_explored = 0.0
+
+    # Explore functional state
+    cnt = 0
+    while !isempty(e)
+        cnt += 1
+        l = dequeue!(e)
+
+        # Disable links of this current state
+        disabled_links::Vector{Tuple{Symbol, Symbol}} = get_disabled_with_dep(st, l)
+        for (u, v) in disabled_links
+            rem_edge!(tgcopy, code_for(tgcopy, u), code_for(tgcopy, v))
+            rem_edge!(tgcopy, code_for(tgcopy, v), code_for(tgcopy, u))
+        end
+
+        # Based on the current network state, compute hot edges (right now it's equal cost shortest paths)
+        st[l].converged_paths = dijkstra_mg_allpaths(tgcopy, src, dst)
+        hot_edges = allpaths_to_unique_links(st[l].converged_paths)
+        force_enabled_links::Vector{Tuple{Symbol, Symbol}} = get_enabled_with_dep(st, l)
+        hot_edges_new = filter(x -> !((x[1], x[2]) in force_enabled_links || (x[2], x[1]) in force_enabled_links), hot_edges)
+
+        # Enqueue new state based on (new) hot edges, with its p_state as prioritization
+        for i in 1:length(hot_edges_new)
+            # NOTE: this is different from NetDice's p_state, since it doesn't contain the shortest paths probability
+            s = State(hot_edges_new[1:i-1], [hot_edges_new[i]], 1)
+            lc::Symbol = Symbol(randstring(10)) # TODO: change this to something else that make sense
+            st[lc] = s
+            st[l, lc] = nothing
+            st[lc].p_state = get_p_state(st, lc)
+            
[…]

### 2022-10-17 src/StateTrees.jl

@@ -1,8 +1,3 @@
-module StateTrees
-
-using Graphs
-using MetaGraphsNext
-
 mutable struct State 
     # Pre-exploration
     force_enabled::Vector{Tuple{Symbol, Symbol}}
@@ -72,5 +67,3 @@ function get_p_state(st::StateTree, l::Symbol)::Float64
 
     return p_state
 end
-
-end # module

### 2022-10-17 src/Tempus.jl

@@ -1,6 +1,8 @@
 module Tempus
 
 import JSON
+using DataStructures
+using Random
 using Distributions
 import Distributions: cdf, pdf, logpdf, minimum, maximum, quantile, convolve
 
@@ -11,28 +13,34 @@ using QuadGK # Numerical integration
 using Roots
 
 include("graphs/TopologyGraphs.jl")
+include("graphs/LatencyGraphs.jl")
 
 include("DirectDistributions.jl")
 include("Direct.jl")
 
+include("StateTrees.jl")
+include("Exploration.jl")
+
 # don't use 'using' or 'import' so that all globals are compartmentalized?
 
 function main()
-    json::String = open("artifacts/7nodenew.json") do file
+    json::String = open("artifacts/Highwinds.json") do file
         read(file, String)
     end
     config::Dict{String, Any} = JSON.parse(json)
 
     tg::TopologyGraph = TopologyGraph(config["routers"], config["links"])
-    println(tg)
+    # lg::LatencyGraph = LatencyGraph(config["routers"], config["links"])
+    # println(lg[:S1_S2, :S2])
+
+    # a::Float64 = explore(tg, Symbol("3"), Symbol("10"))
+    # println(a)
     # topgraph, delaygraph, fwdtable, intent = ConfigParser.parseconf("artifacts/Highwinds.json")
     # a = ModifiedYen.tempus_yen_stateful(topgraph, Symbol(intent["src"]), Symbol(intent["dst"]))
     # println(a)
 
-    # x = Normal(1, 1)
-    # y = convolve(x, x)
-    # z = convolve(y, y)
-    # println(cdf(z, 4.0))
+    x = ecmpprob([[:S1, :S2, :S3, :S4], [:S1, :S2, :S5, :S4], [:S1, :S6, :S7, :S4]])
+    println(x)
 end
 
 if !isinteractive()

### 2022-10-17 src/graphs/LatencyGraphs.jl

@@ -1,0 +1,42 @@
+# EdgeData = Distribution
+lgweightfunction = x -> x
+const LatencyGraph = typeof(MetaGraph(DiGraph(), EdgeData = Distribution, weight_function = lgweightfunction))
+
+function LatencyGraph(routers::Vector{Any}, links::Vector{Any})::LatencyGraph
+    tg::LatencyGraph = MetaGraph(DiGraph(), EdgeData = Distribution, weight_function = lgweightfunction)
+
+    # Make the nodes
+    # TODO: src and dst
+    for router::Dict{String, Any} in routers
+        tg[Symbol(router["name"])] = nothing
+
+        for port::Dict{String, Any} in router["outQdelayModel"]
+            if port["to"] == "dst"
+                continue
+            end
+            
+            outQ::Symbol = Symbol(router["name"] * "_" * port["to"])
+            tg[outQ] = nothing
+
+            tg[Symbol(router["name"]), outQ] = gendist(port["delayModel"])
+        end
+    end
+
+    # Make the links
+    for link::Dict{String, Any} in links
+        if link["u"] == "src" || link["v"] == "src" || link["u"] == "dst" || link["v"] == "dst"
+            continue
+        end
+
+        tg[Symbol(link["u"] * "_" * link["v"]), Symbol(link["v"])] = gendist(link["delayModel"])
+        tg[Symbol(link["v"] * "_" * link["u"]), Symbol(link["u"])] = gendist(link["delayModel"])
+    end
+
+    return tg
+end
+
+function gendist(delaymodel::Dict{String, Any})::Distribution
+    # TODO: truncate to disallow negative delay
+    args = Tuple(delaymodel["args"])
+    return eval(Meta.parse(delaymodel["delayType"] * string(args)))
+end

### 2022-10-17 src/graphs/TopologyGraphs.jl

@@ -15,7 +15,7 @@ function TopologyGraph(routers::Vector{Any}, links::Vector{Any})::TopologyGraph
         if link["u"] == "src" || link["v"] == "src" || link["u"] == "dst" || link["v"] == "dst"
             continue
         end
-        tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], link["w_uv"])
+        tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], link["w_uv"]) # Link + output queue delay
         tg[Symbol(link["v"]), Symbol(link["u"])] = (link["failProb"], link["w_vu"])
     end
 

### 2022-10-17 src/netdiceparser/NetDiceTopologyParser.jl

@@ -5,19 +5,11 @@ import JSON
 function main()
     tempusConfig = Dict()
 
-    open("artifacts/Highwinds.in") do f 
+    open("artifacts/AttMpls.in") do f 
         
-        s = readline(f) # Read the first line containing the number of nodes
-        routers = [Dict(
-            "name" => string(i - 1), 
-            "failProb" => 0.0,
-            "delayModel" => Dict(
-                "delayType" => "Normal",
-                "args" => [1, 0]
-            )
-        ) for i in 1:parse(Int64, s)]
-
-        tempusConfig["routers"] = routers
+        r = readline(f) # Read the first line containing the number of nodes
+        
+        # outQ = Dict{String, Vector{String}}()
 
         links = []
         while !eof(f)
@@ -26,21 +18,36 @@ function main()
             push!(links, Dict(
                 "u" => array[1], 
                 "v" => array[2],
+                "w_uv" => 1,
+                "w_vu" => 1,
                 "failProb" => 0.1,
                 "delayModel" => Dict(
                     "delayType" => "Normal",
                     "args" => [1, 0]
                 )
             ))
+
+            # outQ[array[1]] = 
+            # outQ[array[2]] = 
         end
 
+        routers = [Dict(
+            "name" => string(i - 1), 
+            "failProb" => 0.0,
+            "delayModel" => Dict(
+                "delayType" => "Normal",
+                "args" => [1, 0]
+            )
+        ) for i in 1:parse(Int64, r)]
+
+        tempusConfig["routers"] = routers
         tempusConfig["links"] = links
 
         println(JSON.json(tempusConfig))
         # println(s)
     end
 
-    open("artifacts/Highwinds.json", "w") do f
+    open("artifacts/AttMpls.json", "w") do f
         write(f, JSON.json(tempusConfig))
     end
 end

### 2022-10-20 paper/main.tex
[Paper] completed latency modelling chapter

@@
 
-\subsection{Latency Distribution}
+\subsection{Path Latency Distribution}
 Up to this point, we have talked about measuring the latency of a single packet by figuring out 
@@
 \subsection{Path Decomposition and Convolution}
-Expand on these points:
-\begin{itemize}
-    \item How do we get the path latency distribution? Component-wise measurement data
-    \item What measurement? link delay and queuing delay $\rightarrow$ not all delays are 
-    significant
-    \item What to do with it? Simulation $\rightarrow$ lack of error guarantee
-    \item Convolution
-    \item Independence?
-\end{itemize}
+The final question that we had to decide on was how do we actually model path latency distribution
+based on real component measurements.
+Considering the complexity of factors that might determine the path latency distribution and the 
+availability of measurement data, we settle on the assumption that the path latency distribution 
+is composed of multiple independent distribution that corresponds to the latency each components 
+in the path might introduce.
+%TODO: confirm about independence?
+
+Since not all components in the path will introduce latency with a non-negligible value, we 
+specifically choose to model two source of latency in the path's components which we deem 
+significant: \textbf{link propagation latency} and \textbf{queuing latency}.
+%TODO: expand on how to get the data later
+
+Propagation delay is the latency that is introduced by the links in the network, which is 
+independent of the traffic load in the system.
+Queuing delay is the latency that the queuing process in the node introduced. 
+Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
+packets there are in the queue, the more delay the node will introduce to a subsequent packets.
+
+For propagation delay, the semantic of this random variable is relatively straightforward: it is 
+the distribution of latency that a given link will introduce.
+For queuing delay however, this random variable represents the delay that a given queuing process 
+will introduce, marginalized over various traffic pattern that a given network state might have 
+resulted.
+
+In order to obtain the overall path latency distribution from these per-component latency 
+distributions, we do a convolution operations over all the relevant component's latency 
+distributions.
+Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
+convolution technique with a guaranteed error bound. 
+We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
+the lack of a general technique to measure error made us opt for the formerly mentioned method.
 
 \section{Encoding}
-We will start with the assertion that latency is a property that only make sense 
-after connectivity between two nodes has been established. 
-In other words, if two nodes in a network aren't connected (due to physical failure 
-or ACL policy), then the latency between them will \textit{always} be infinite.
-Because of this, we divide the problem of latency verification into two parts: 
+We divide the problem of latency verification into two parts: 
 verifying that two nodes are reachable (\textit{functional} property) and only if 
 the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some additional condition (\textit{temporal} property)
+two nodes fulfilled some additional condition (\textit{temporal} property).
 
@@
 
+\subsection{Routing Label}
 On top of this topology, we could define additional information that would be used
@@
 
-\subsection{Latency Graph}
-There are many source of latency in a network. 
-The principal challenge for us, then, is to determine the appropriate level of abstraction; 
-detailed enough in order to accomplish our verification goal correctly but 
-abstract enough for the problem to remain tractable. 
-We settle on modeling two significant sources of latency: propagation and 
-queuing delay.
+\subsection{Latency Label}
+Expand on these points
+\begin{itemize}
+    \item distributions as label 
+    \item convolve the relevant output queue latency + link latency
+\end{itemize}
+% There are many source of latency in a network. 
+% The principal challenge for us, then, is to determine the appropriate level of abstraction; 
+% detailed enough in order to accomplish our verification goal correctly but 
+% abstract enough for the problem to remain tractable. 
+% We settle on modeling two significant sources of latency: propagation and 
+% queuing delay.
 
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
+% To take both of these forms of delay into account, we've decided to model the latency of each
+% components with a univariate continuous random variable.
 
-To take both of these forms of delay into account, we've decided to model the latency of each
-components with a univariate continuous random variable.
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
+% % TODO: change this to v_i
+% To convey this additional notion of latency distribution, we form a second edge-labeled directed 
+% graph (called a latency graph) $G_l = (V_l, E_l)$.
+% For each network node $v_i \in V_t$, we want to ``expand'' this node to be able to represent the 
+% inner working of the queuing process within each node.
 
-% TODO: change this to v_i
-To convey this additional notion of latency distribution, we form a second edge-labeled directed 
-graph (called a latency graph) $G_l = (V_l, E_l)$.
-For each network node $v_i \in V_t$, we want to ``expand'' this node to be able to represent the 
-inner working of the queuing process within each node.
+% \begin{itemize}
+%     \item Each $v_i$ is expanded into $v_i^{in}, v_i^{out-1}, ..., v_i^{out-n} \in V_l$ 
+%     where $n$ is the amount $v_i$'s outward neighbors.
 
-\begin{itemize}
-    \item Each $v_i$ is expanded into $v_i^{in}, v_i^{out-1}, ..., v_i^{out-n} \in V_l$ 
-    where $n$ is the amount $v_i$'s outward neighbors.
+%     \item Every inward edge that $v_i$ has in $G_t$ will all get directed to $v_i^{in}$.
 
-    \item Every inward edge that $v_i$ has in $G_t$ will all get directed to $v_i^{in}$.
+%     \item Every outward edge that $v_i$ has in $G_t$ will be sourced from one of the 
+%     $v_i^{out-j}$ instead.
 
-    \item Every outward edge that $v_i$ has in $G_t$ will be sourced from one of the 
-    $v_i^{out-j}$ instead.
+%     \item Add additional edges $e_i^j$ that connects from $v_i^{in}$ to each $v_i^{out-j}$.
 
-    \item Add additional edges $e_i^j$ that connects from $v_i^{in}$ to each $v_i^{out-j}$.
+%     \item Finally, we label each $e \in E_l$ with the random variable as the latency distribution.
+% \end{itemize}
 
-    \item Finally, we label each $e \in E_l$ with the random variable as the latency distribution.
-\end{itemize}
-
-Semantically, $v_i^{out-j}$ represents a port, normal links that are copied from $G_t$ represents 
-a link that has a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
-in node $v_i$ that has a queuing delay.
+% Semantically, $v_i^{out-j}$ represents a port, normal links that are copied from $G_t$ represents 
+% a link that has a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
+% in node $v_i$ that has a queuing delay.
 
@@
 
-\subsubsection{Functional Verification}
+\subsection{Functional Verification}
 Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of links that are 
@@
 
-\subsubsection{Temporal Verification}
+\subsection{Temporal Verification}
 Since we assign each links and output queue in the network with a certain latency distribution, 

### 2022-11-03 paper/main.tex
[Paper] Changed section 4 and trying to add figs

@@
 \usepackage{amsfonts} 
+\usepackage{graphicx}
+
+\graphicspath{ {./graph/} }
 
@@
 Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that that path exists in the first place.
+path however, we must make sure that said path exists in the first place.
 We note that latency is a property that only make sense after connectivity between two hosts 
@@
 
-\subsection{Path Decomposition and Convolution}
+\subsection{Path Decomposition and Convolution} \label{decomposition}
 The final question that we had to decide on was how do we actually model path latency distribution
@@
 In this framework, the physical network is encoded in an edge-labeled directed graph 
-$G_t = (V_t, E_t)$ where $V_t$ represents the routers in the network and 
-$E_t$ represents the connectivity between a pair of source and destination router. 
-The function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x \le 1\}$ is the 
-edge-label that represents the failure rate of a given connectivity link.
+$G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
+$E_t$ represents the functional connectivity between a source and a destination node. 
+A physical link is then represented as a pair of symmetrical edges that shares the same source 
+and destination node but with opposite direction (e.g. $v_1 \rightarrow v_2$ and $v_2 
+\rightarrow v_1$).
 
-\subsection{Routing Label}
-On top of this topology, we could define additional information that would be used
-by a routing protocol to determine the valid path(s) between two nodes 
-$src, dst \in V_t$ given a particular link failure scenario. 
+To represent the component's random failure, NetDice defined two failure rates.
+The first failure rate is the chance that a given physical link in the network will go down.
+The second failure rate is the chance that a given node in the network will go down.
+These rates are shared between all links / nodes in the network.
+Internally however, the node failure rate will be translated into the link failure rate 
+with a Bayesian Network model, since a node failure can be modeled with the failure of each 
+links connected to it.
+
+We refine NetDice's model slightly by allowing each links to have different failure rate.
+We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
+\le 1\}$ that represents the failure rate of a given physical link.
+As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
+opposite direction) will also share the same failure rate, and will be disabled in a coupled 
+fashion.
+
+\subsection{Routing}
+On top of this topology, NetDice also defined additional informations that would be used
+by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
+a particular link failure scenario. 
+These routing informations would also be used by their optimization algorithms to reduce the 
+amount of states that are going to be explored.
+
+NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
+They encode the relevant routing information by assigning some labels to the vertices or edges
+in the topology.
 In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
 that represents the positive weight of a link.
+They would later be used to compute the convergent paths of a given network state 
+(\ref{verification}).
 
 \subsection{Latency Label}
-Expand on these points
-\begin{itemize}
-    \item distributions as label 
-    \item convolve the relevant output queue latency + link latency
-\end{itemize}
-% There are many source of latency in a network. 
-% The principal challenge for us, then, is to determine the appropriate level of abstraction; 
-% detailed enough in order to accomplish our verification goal correctly but 
-% abstract enough for the problem to remain tractable. 
-% We settle on modeling two significant sources of latency: propagation and 
-% queuing delay.
+Similarly, we could also use this edge-labeling technique to represent latency. 
+As mentioned in \ref{decomposition}, our model will encode two kinds of latency: link propagation 
+and queuing in the node.
+We will encode these latencies by equipping the model with two additional labels.
 
-% To take both of these forms of delay into account, we've decided to model the latency of each
-% components with a univariate continuous random variable.
+Encoding link propagation latency is fairly straightforward. 
+We define a function $l_p: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$ is a set of 
+continuous univariate distribution that has a minimum value of $0$.
+This distribution signifies the time it would take for the respective physical link to transmit 
+a packet from one end to another.
+Because of this, just like $r$, two symmetrical edges will share the same distribution.
 
-% % TODO: change this to v_i
-% To convey this additional notion of latency distribution, we form a second edge-labeled directed 
-% graph (called a latency graph) $G_l = (V_l, E_l)$.
-% For each network node $v_i \in V_t$, we want to ``expand'' this node to be able to represent the 
-% inner working of the queuing process within each node.
+To encode node queuing latency, we first note that the queuing mechanism in modern switches 
+usually resides on the output port. %TODO: Cite PISA?
+Since a port in a switch is only connected to one other port, we could effectively assign the 
+latency to the connectivity between switches.
+To do this, we define a function $l_q: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$.
+Unlike $l_p$ however, two symmetrical edges will have different distribution since they 
+represent two different output queues.
 
-% \begin{itemize}
-%     \item Each $v_i$ is expanded into $v_i^{in}, v_i^{out-1}, ..., v_i^{out-n} \in V_l$ 
-%     where $n$ is the amount $v_i$'s outward neighbors.
-
-%     \item Every inward edge that $v_i$ has in $G_t$ will all get directed to $v_i^{in}$.
-
-%     \item Every outward edge that $v_i$ has in $G_t$ will be sourced from one of the 
-%     $v_i^{out-j}$ instead.
-
-%     \item Add additional edges $e_i^j$ that connects from $v_i^{in}$ to each $v_i^{out-j}$.
-
-%     \item Finally, we label each $e \in E_l$ with the random variable as the latency distribution.
-% \end{itemize}
-
-% Semantically, $v_i^{out-j}$ represents a port, normal links that are copied from $G_t$ represents 
-% a link that has a propagation delay, and newly introduced links $e_i^j$ represents an output queue 
-% in node $v_i$ that has a queuing delay.
-
-\section{Verification}
-From these two graphs, we then do the verification in the following way:
+\section{Verification} \label{verification}
+From a labeled graph that encodes a given network, we will then do functional verification 
+and temporal verification in the following way:
 
 \subsection{Functional Verification}
+For functional verification, we will essentially do the exploration technique employed by NetDice.
+However, we made some modifications to the algorithm in order for it to remember and pass 
+necessary informations to the next stage, which is the temporal verification.
+We will describe the high-level description of NetDice's algorithm in light of our modification
+as follows:
+
+\subsubsection{Network State}
 Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of links that are 
@@
 
-While natural, this brute-force algorithm is obviously doesn't scale since it will take $2^n$ 
-iteration ($n$ as the number of links in the topology). 
-NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could skip over some 
-events that are guaranteed not to change the convergent paths of the current network state, by 
+While natural, this brute-force algorithm is obviously doesn't scale since it will take 
+$2^{|E_t|}$ iterations.
+
+\subsubsection{Super-state}
+NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could prune over some 
+states that are guaranteed not to change the convergent paths of the current network state, by 
 marginalizing over \textit{cold edges}.
+Said another way, we could merge multiple network states into a single super-state that shares 
+the same convergent paths.
 
@@
 
-NetDice's exploration will effectively form a tree, where the root is the event where we have 
-empty hot edges, and the leaves are where the hot edges are full.
+NetDice's exploration will effectively form a tree, where each level of the tree represents 
+how many links we will explicitly put down.
+The root is the event where we have empty hot edges, and the leaves are where the hot edges 
+are full.
+
+\subsubsection{Convergent Paths}
+Remember instead of discarded
+
+\subsubsection{State and Functional Path Probability}
+State
 
@@
 
+\subsubsection{Mega-state}
 The first is what we call \textbf{hot-edges union}. 
@@
 Therefore, we could merge these two events and compute the convolution for a given path once.
+
+\subsubsection{Path Memoization}
 
@@
 
-\section{Proposed Evaluation}
-On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
+\section{Evaluation}
+\subsection{Scalability}
 
-On the performance side, we will evaluate the verifier by measuring the amount of additional 
-overhead that the temporal verifier would introduce, measured by the amount of states that 
-need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-produced by step 1, only 459 got re-explored on step 2.
+\subsubsection{States Explored}
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{states}
+    \caption{States explored for temporal verification and functional verification}
+    \label{fig:states}
+\end{figure}
+
+\subsubsection{Wall-clock Time}
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{wallclock}
+    \caption{Time took for temporal verification}
+    \label{fig:wallclock}
+\end{figure}
+
+\subsection{Optimization Effectiveness}
+
+% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
+% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
+% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
+
+% On the performance side, we will evaluate the verifier by measuring the amount of additional 
+% overhead that the temporal verifier would introduce, measured by the amount of states that 
+% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
+% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
+% produced by step 1, only 459 got re-explored on step 2.
+
+\section{Related Works}
+
+\section{Future Works}
+
+\section{Conclusion}
 

### 2022-11-10 paper/main.tex
[Paper] Finished 5.1

@@
 However, we made some modifications to the algorithm in order for it to remember and pass 
-necessary informations to the next stage, which is the temporal verification.
-We will describe the high-level description of NetDice's algorithm in light of our modification
-as follows:
+necessary informations to the temporal verification stage.
+To contextualize our modification, we will briefly describe the description of NetDice's algorithm 
+in three parts: Network States, Equivalence Classes, and Functional Exploration.
+For a more complete complete description of NetDice, please refer to the original paper.
 
-\subsubsection{Network State}
-Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of links that are 
-alive / up and $D$ is a set of links that are dead / down.
+\subsubsection{Network States}
+Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of all links that 
+are alive / up and $D$ is a set of all links that are dead / down.
+(Consequently, $E_t = U \cup D$).
 We then define a set $S$ that represents all possible network states in a given topology. 
-Our goal is to identify a subset of $S$ where each element is a network state that can connect 
-the source and destination node given a convergent behavior of a routing protocol.
+The goal of functional verification is to identify the largest subset of $S$ where each element 
+of this set is a network state that can connect the source and destination node given a convergent 
+behavior of the control plane.
+We will call this subset \textit{reachability subset}.
+While NetDice posseses the ability to verify other properties, we only focuses on basic 
+reachability, the most relevant property for latency verification.
 
 This framework fits nicely with the framework of probability theory, where $S$ represents the 
-sample space and the power set of $S$ represents the possible events, one of which is the one
-we care about: the event that a packet from a source node can reach a destination node.
-The way we assign probability to events is to simply do a summation over the product of all the 
-failure and success rate of each links in the network state.
+sample space and the power set of $S$ represents the possible probability events, one of which is 
+the previously mentioned reachability subset.
+The way we assign probability values to a subset is to simply do a summation over the probability 
+value of each network state, which in turn is the product of all the failure and success rate of 
+each links in the network state.
 
-The straightforward and naive way to compute the probability of this event is to iterate over all 
-the network state. 
-We then determine whether that network state can fulfill said property and which path it will 
-take. 
-If yes, then we would compute the probability from the product of the link's failure and success 
-rate, and add it to the running sum.
-If no, then we could skip the computation altogether.
-
-While natural, this brute-force algorithm is obviously doesn't scale since it will take 
+The naive approach to compute the probability of the reachability subset is to exhaustively iterate 
+over all the network state and check the reachability property. 
+While natural, this brute-force algorithm doesn't scale well since it will take 
 $2^{|E_t|}$ iterations.
 
-\subsubsection{Super-state}
-NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could prune over some 
-states that are guaranteed not to change the convergent paths of the current network state, by 
-marginalizing over \textit{cold edges}.
-Said another way, we could merge multiple network states into a single super-state that shares 
-the same convergent paths.
+\subsubsection{Equivalence Classes}
+NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could combine over 
+some states that are guaranteed not to change the convergent paths (and thus the reachability 
+property) of the current network state, by marginalizing over what is called \textit{cold edges}.
 
-In NetDice, we start from a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$).
-We then determine whether that network state can fulfill said property and which path it will 
-take, similar to the brute-force method.
-If no, then it is safe to say that the source node can never reach the destination node in all 
-network state.
-If yes, then we could use the logic of the routing protocol to determine the set of cold edges, 
-(i.e. edges whose failure won't change the convergent paths). 
-Edges that are not part of this set is called \textit{hot edges} $h_i = (U, D)$.
-Instead of using $s_i$ to compute the probability of the current state, NetDice only use the set 
-of hot edges to compute the probability of them being alive ($h_i = (U, D)$ where $D = \emptyset$).
-This in effect computes the probability of multiple network state at once.
+Cold edges are defined to be links whose failure is guaranteed not to change the property in 
+question.
+In the specific case of reachability, this essentially mean that the failure of such edge 
+won't change the convergent paths of the network state in question.
+By marginalizing over its cold edges and combining the state into one equivalence class, instead 
+of computing the probability of individual network states, we are left with computing the 
+probability from the remaining edges, which is called \textit{hot edges}.
 
-Next, instead of examining other network state at random, we instead iterate over the links in 
-$h_i$.
-By definition, if we fail a link that belongs to $h_i$, then the convergent paths will change.
-To cover all of those possibilities efficiently, NetDice iterate over all the links in $h_i$, 
-disabled them, re-enabled them, and move to the next edge to do the same.
-If $h_i = ((e_1, e_2, ..., e_j), \emptyset)$, then we would enqueue $(\emptyset, (e_1))$, 
-$((e_1), (e_2))$, $((e_1, e_2), (e_3))$, ..., $((e_1, e_2, ..., e_{j-1}), (e_j))$.
-Thus, instead of exploring $2^j$ events, we will enqueue $j$ events instead.
+Hot edges are the opposite of cold edges: their state (whether they are up or down) will change 
+the convergent path of the network state in question.
+Mostly, they consists of the links that make up the convergent paths itself.
 
-NetDice's exploration will effectively form a tree, where each level of the tree represents 
-how many links we will explicitly put down.
+By identifying the cold edges of a given network state, NetDice reduces the search-space by 
+merging multiple network states into fewer equivalence classes. 
+However, this optimization also serves another purpose: providing an efficient exploration 
+strategy.
+
+\subsubsection{Functional Exploration}
+To actually identify the equivalence classes that might emerge in a given network, NetDice employs
+a tree-based exploration method based on hot edges computation.
+
+We start with a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$ and $U = E_t$).
+We then compute the convergent paths and determine the hot edges set of that network state.
+Since, by definition, shutting down the link in this set will change the convergent paths, 
+instead of exhaustively exploring some other state at random, NetDice would continue the 
+exploration by systematically failing one of these links instead.
+
+% We then determine whether that network state can fulfill said property and which path it will 
+% take, similar to the brute-force method.
+% If no, then it is safe to say that the source node can never reach the destination node in all 
+% network state.
+% If yes, then we could use the logic of the routing protocol to determine the set of cold edges, 
+% (i.e. edges whose failure won't change the convergent paths). 
+% Edges that are not part of this set is called \textit{hot edges} $h_i = (U, D)$.
+% Instead of using $s_i$ to compute the probability of the current state, NetDice only use the set 
+% of hot edges to compute the probability of them being alive ($h_i = (U, D)$ where $D = \emptyset$).
+% This in effect computes the probability of multiple network state at once.
+
+% Next, instead of examining other network state at random, we instead iterate over the links in 
+% $h_i$.
+% By definition, if we fail a link that belongs to $h_i$, then the convergent paths will change.
+% To cover all of those possibilities efficiently, NetDice iterate over all the links in $h_i$, 
+% disabled them, re-enabled them, and move to the next edge to do the same.
+% If $h_i = ((e_1, e_2, ..., e_j), \emptyset)$, then we would enqueue $(\emptyset, (e_1))$, 
+% $((e_1), (e_2))$, $((e_1, e_2), (e_3))$, ..., $((e_1, e_2, ..., e_{j-1}), (e_j))$.
+% Thus, instead of exploring $2^j$ events, we will enqueue $j$ events instead.
+
+NetDice will recursively continue this process for the new state and effectively form an exploration 
+tree, where each level of the tree represents how many links we will explicitly put down.
 The root is the event where we have empty hot edges, and the leaves are where the hot edges 
@@
 
-\subsubsection{Convergent Paths}
-Remember instead of discarded
+\subsubsection{Making Exploration Explicit}
+After the original exploration algorithm finishes, NetDice will output the final probability (and 
+error range) of the reachability property in all the equivalence classes that has been explored. 
+However, this information alone isn't useful for us to reason about latency.
+Thus, we modified the original exploration algorithm in order to make it more explicit, so that its 
+output can be used effectively in the next stage.
 
-\subsubsection{State and Functional Path Probability}
-State
+We started by formalizing the notion of Equivalence Classes.
+An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
+$U_h$ and $D_h$ refers to a set of links in the hot edges that are up and down respectively.
+They are similar to $U$ and $D$ in network state, with a notable exception that 
+$E_t \neq U_h \cup D_h$, since they only represents hot edges, and not including cold edges.
+$P_{conv}$ refers to the convergent paths between src-dst pair when the links in $D_h$ fails.
+We explicitly store $P_{conv}$ in $\mathcal{E}$ (unlike the original algorithm) so that it could 
+be used to for the temporal verification stage.
+
+We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
+At the root, we have an Equivalence Class that corresponds to a perfect network, where $P_{conv}$ 
+is the path(s) that the control plane will produce given a perfect network condition and 
+$U_h = D_h = \emptyset$.
+Each Equivalence Class in this tree will have children which have one of its parent's link in 
+$P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
+If an Equivalence Class has an empty $P_conv$, then it would have no children.
+
+To compute the functional probability of a given Equivalence Class, we compute the product 
+of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the product of the 'down'
+probability of all the links in $D_h$.
+
+After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
+stage.
 
 \subsection{Temporal Verification}
+In the temporal verification stage, we will use the $P_{conv}$ within each Equivalence Class in 
+$\mathcal{T}$ and the latency distribution labels $l_p$ and $l_q$ to augment the functional 
+probability we had in the previous stage to actually answer some questions about latency property.
+
+\subsubsection{Convolution}
 Since we assign each links and output queue in the network with a certain latency distribution, 
-computing the probability of the temporal property involves convolving the the paths we got from 
-functional verification and measuring the statistical property of the resulting distribution.
-However, while NetDice has shrunk the exploration space compared to brute-force iteration, for
-the purpose of temporal verification, further optimization can be done.
+computing the probability of the temporal property involves convolving the the path(s) we got from 
+$P_{conv}$ and measuring the statistical property of the resulting distribution.
+This essentially introduced the need for the verifier to ``re-explore'' $\mathcal{T}$ in order to 
+compute the temporal probability of each Equivalence Class.
 
-\subsubsection{Mega-state}
+%TODO: explain the convolution, weighting, DIRECT
+
+However, while NetDice has shrunk the exploration space compared to exhaustive iteration, they still 
+contain some inefficiencies that could be optimized.
+
+\subsubsection{Equivalence Class Consolidation}
 The first is what we call \textbf{hot-edges union}. 

### 2022-11-10 paper/main.tex
[Paper] Continued section 5

@@
 
-\subsubsection{Convolution}
-Since we assign each links and output queue in the network with a certain latency distribution, 
-computing the probability of the temporal property involves convolving the the path(s) we got from 
-$P_{conv}$ and measuring the statistical property of the resulting distribution.
-This essentially introduced the need for the verifier to ``re-explore'' $\mathcal{T}$ in order to 
-compute the temporal probability of each Equivalence Class.
+\subsubsection{Path Convolution and Weighted Average}
+Given $P_{conv}$, $l_p$, and $l_q$, we will do the following:
+\begin{enumerate}
+    \item Split $P_{conv}$ into its individual path
+    \item For each path, compute its latency distribution by convolving the latency distribution 
+        $l_p$ and $l_q$ of each link
+    \item With the resulting path latency distribution, determine the probability of temporal 
+        property by computing the statistical property of said distribution
+    \item Combine the temporal property of each path by computing the weighted average, based 
+        on the load-balancing scheme of the control plane
+\end{enumerate}
 
-%TODO: explain the convolution, weighting, DIRECT
+We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
+overhead to the exploration algorithm (on top of functional exploration).
+In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
+those $n$ Equivalence Class.
+We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
+later.
 
-However, while NetDice has shrunk the exploration space compared to exhaustive iteration, they still 
-contain some inefficiencies that could be optimized.
+\subsubsection{Numerical Convolution}
+One problem with this algorithm, is that not every convolution between distributions have a 
+closed-form solution and could be done analytically.
+Therefore, we need to leverage a numerical convolution method in order to convolve two arbitrary
+distributions.
+
+%TODO: explain the convolution, weighting, DIRECT, changing order of DIRECT
+Expand points on 
+\begin{itemize}
+    \item DIRECT 
+    \item orderings DIRECT
+\end{itemize}
 
 \subsubsection{Equivalence Class Consolidation}
-The first is what we call \textbf{hot-edges union}. 
+The first is what we call \textbf{Equivalence Class Consolidation}. 
 The idea is, while each hot-edges events has the same convergent paths, it is not guaranteed to be 
@@
 \subsubsection{Path Memoization}
+
+\subsubsection{Convolution Grouping}
 

### 2022-11-15 paper/main.tex
[Paper] merged Encoding and Verification section

@@
 \section{Overview}
-To be written
-
-\section{On Latency Modeling}
+% \section{On Latency Modeling}
 \subsection{Latency as Path Property} 
@@
 
-\section{Encoding}
+\section{Verifier Design}
 We divide the problem of latency verification into two parts: 
-verifying that two nodes are reachable (\textit{functional} property) and only if 
+verifying that two nodes are reachable (\textbf{functional verification}) and only if 
 the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some additional condition (\textit{temporal} property).
+two nodes fulfilled some temporal condition (\textbf{temporal verification}).
 
-\subsection{Topology Graph}
+\subsection{Functional Verification}
 For functional property verification, NetDice \cite{steffen2020probabilistic} 
@@
 probabilistic and efficient manner.
-In this framework, the physical network is encoded in an edge-labeled directed graph 
+To contextualize our modification, we will briefly describe some parts of 
+NetDice's network encoding and exploration algorithm as a prelude to our alteration.
+For a more complete complete description of NetDice, please refer to the original 
+paper.
+
+\subsubsection{Topology Graph}
+In this framework, a network is encoded in an edge-labeled directed graph 
 $G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
@@
 
-To represent the component's random failure, NetDice defined two failure rates.
-The first failure rate is the chance that a given physical link in the network will go down.
-The second failure rate is the chance that a given node in the network will go down.
-These rates are shared between all links / nodes in the network.
-Internally however, the node failure rate will be translated into the link failure rate 
-with a Bayesian Network model, since a node failure can be modeled with the failure of each 
-links connected to it.
+% To represent the component's random failure, NetDice defined two failure rates.
+% The first failure rate is the chance that a given physical link in the network will go down.
+% The second failure rate is the chance that a given node in the network will go down.
+% These rates are shared between all links / nodes in the network.
+% Internally however, the node failure rate will be translated into the link failure rate 
+% with a Bayesian Network model, since a node failure can be modeled with the failure of each 
+% links connected to it.
 
+To represent random failure, NetDice defined a universal link failure rate (i.e. 
+the probability of \textit{any} link in the network randomly failing).
 We refine NetDice's model slightly by allowing each links to have different failure rate.
@@
 
-\subsection{Routing}
-On top of this topology, NetDice also defined additional informations that would be used
-by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
-a particular link failure scenario. 
-These routing informations would also be used by their optimization algorithms to reduce the 
-amount of states that are going to be explored.
+% \subsubsection{Routing}
+% On top of this topology, NetDice also defined additional informations that would be used
+% by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
+% a particular link failure scenario. 
+% These routing informations would also be used by their optimization algorithms to reduce the 
+% amount of states that are going to be explored.
 
-NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
-They encode the relevant routing information by assigning some labels to the vertices or edges
-in the topology.
-In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
-that represents the positive weight of a link.
-They would later be used to compute the convergent paths of a given network state 
-(\ref{verification}).
+% NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
+% They encode the relevant routing information by assigning some labels to the vertices or edges
+% in the topology.
+% In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
+% that represents the positive weight of a link.
+% They would later be used to compute the convergent paths of a given network state 
+% (\ref{verification}).
 
-\subsection{Latency Label}
-Similarly, we could also use this edge-labeling technique to represent latency. 
-As mentioned in \ref{decomposition}, our model will encode two kinds of latency: link propagation 
-and queuing in the node.
+\subsubsection{State Exploration}
+A \textbf{network state} is defined as a set of failed links.
+There are $2^{|E_t|}$ network states in a given network and a brute-force strategy 
+would need to explore all of the states to compute the reachability property of a 
+node pair in a given network. 
+NetDice however, will merge multiple network states into an \textbf{equivalence class}
+by marginalizing over \textit{cold edges}, links whose failure is guaranteed not to 
+change the convergent path of a highlighted network state.
+
+NetDice will systematically explore many equivalence classes in the network. 
+It will start from the equivalence class of a perfect network and will fail certain 
+\textit{hot edges} (i.e. edges that are not cold) to explore another equivalence 
+class.
+While not explicit in its description, NetDice's algorithm will effectively form an 
+\textbf{exploration tree}.
+
+In order to preserve some information that will be used in the temporal verification 
+stage, we modify this original exploration algorithm to make it more explicit.
+We started by formalizing the notion of Equivalence Classes.
+An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
+$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
+$P_{conv}$ refers to the convergent paths between src-dst pair that is produced by the
+control plane when the links in $D_h$ fails.
+Unlike the original algorithm, we explicitly store $P_{conv}$ in $\mathcal{E}$ so that 
+it could be used to for the temporal verification stage.
+
+We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
+At the root, we have an Equivalence Class that corresponds to a perfect network, where 
+$P_{conv}$ is the path(s) that the control plane will produce given a perfect network 
+condition and $U_h = D_h = \emptyset$.
+Each Equivalence Class in this tree will have children which have one of its parent's link in 
+$P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
+If an Equivalence Class has an empty $P_conv$, then it would have no children.
+
+To compute the functional probability of a given Equivalence Class, we compute the 
+product of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the 
+product of the 'down' probability of all the links in $D_h$.
+
+After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
+stage.
+
+\subsection{Temporal Verification}
+In the temporal verification stage, we will use the path information that we got from 
+functional verification stage and augment them with latency information to 
+produce the final temporal property probability.
+
+\subsubsection{Latency Label}
+Similar to link failure $r$, we use edge-labeling technique to represent latency that 
+will get introduced by the connectivity between two nodes. 
+Our model will model two significant sources of latency: \textbf{link propagation} 
+and \textbf{packet queuing in the node}.
 We will encode these latencies by equipping the model with two additional labels.
@@
 
-To encode node queuing latency, we first note that the queuing mechanism in modern switches 
-usually resides on the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively assign the 
-latency to the connectivity between switches.
-To do this, we define a function $l_q: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$.
-Unlike $l_p$ however, two symmetrical edges will have different distribution since they 
+To encode queuing latency, we first note that the queuing mechanism in modern switches 
+usually resides in the output port. %TODO: Cite PISA?
+Since a port in a switch is only connected to one other port, we could effectively 
+assign the latency to the connectivity between switches.
+To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
+This distribution signifies the output queue latency in the source node that is 
+going to be forwarded to the destination node.
+Unlike $l_p$, two symmetrical edges will have two different distributions since they 
 represent two different output queues.
 
-\section{Verification} \label{verification}
-From a labeled graph that encodes a given network, we will then do functional verification 
-and temporal verification in the following way:
+Given $P_{conv}$, $l_p$, and $l_q$, we could then compute the probability of a given 
+temporal property being fulfilled.
+We will describe our algorithm for computing this probability in the next few 
+sections.
 
-\subsection{Functional Verification}
-For functional verification, we will essentially do the exploration technique employed by NetDice.
-However, we made some modifications to the algorithm in order for it to remember and pass 
-necessary informations to the temporal verification stage.
-To contextualize our modification, we will briefly describe the description of NetDice's algorithm 
-in three parts: Network States, Equivalence Classes, and Functional Exploration.
-For a more complete complete description of NetDice, please refer to the original paper.
+\subsubsection{Weighted Average and Path Convolution}
+Since a convergent behavior of the forwarding plane $P_{conv}$ might contain more than 
+one paths, we will compute the temporal probability of $P_{conv}$ by computing the 
+weighted average of the temporal probability of each of the individual paths.
+The specific weight depends on the load-balancing method used.
+We have implemented the weight computation for ECMP load-balancing scheme.
 
-\subsubsection{Network States}
-Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of all links that 
-are alive / up and $D$ is a set of all links that are dead / down.
-(Consequently, $E_t = U \cup D$).
-We then define a set $S$ that represents all possible network states in a given topology. 
-The goal of functional verification is to identify the largest subset of $S$ where each element 
-of this set is a network state that can connect the source and destination node given a convergent 
-behavior of the control plane.
-We will call this subset \textit{reachability subset}.
-While NetDice posseses the ability to verify other properties, we only focuses on basic 
-reachability, the most relevant property for latency verification.
+To get the temporal probability of a single path, we will use the latency information 
+from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
+To do this, we resort to the methods of \textbf{convolution}.
+For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
+all together to get the path latency distribution.
 
-This framework fits nicely with the framework of probability theory, where $S$ represents the 
-sample space and the power set of $S$ represents the possible probability events, one of which is 
-the previously mentioned reachability subset.
-The way we assign probability values to a subset is to simply do a summation over the probability 
-value of each network state, which in turn is the product of all the failure and success rate of 
-each links in the network state.
+With this path latency distribution, we could get the temporal probability of a 
+path by computing the statistical property of said distribution.
+In this work, given a path latency distribution $\mathcal{L}$, we define two temporal 
+properties:
+\begin{itemize}
+    \item \textbf{Bounded Reachability}: the probability that a packet will get 
+        transmited below $t$ time unit. This will get computed as $cdf(\mathcal{L}, t)$
+    \item \textbf{Tail Reachability}: the probability that a packet will get 
+        transmited above $t$ time unit. This will get computed as 
+        $1 - cdf(\mathcal{L}, t)$
+\end{itemize}
 
-The naive approach to compute the probability of the reachability subset is to exhaustively iterate 
-over all the network state and check the reachability property. 
-While natural, this brute-force algorithm doesn't scale well since it will take 
-$2^{|E_t|}$ iterations.
+In short, given $P_{conv}$, $l_p$, and $l_q$, we will compute:
+\begin{enumerate}
+    \item Split $P_{conv}$ into its individual path
+    \item For each path, compute its latency distribution by convolving the latency distribution 
+        $l_p$ and $l_q$ of each link
+    \item With the resulting path latency distribution, determine the probability of temporal 
+        property by computing the statistical property of said distribution
+    \item Combine the temporal property of each path by computing the weighted average, based 
+        on the load-balancing scheme of the control plane
+\end{enumerate}
 
-\subsubsection{Equivalence Classes}
-NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could combine over 
-some states that are guaranteed not to change the convergent paths (and thus the reachability 
-property) of the current network state, by marginalizing over what is called \textit{cold edges}.
+We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
+overhead to the exploration algorithm (on top of functional exploration).
+In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
+those $n$ Equivalence Class.
+We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
+later.
 
-Cold edges are defined to be links whose failure is guaranteed not to change the property in 
-question.
-In the specific case of reachability, this essentially mean that the failure of such edge 
-won't change the convergent paths of the network state in question.
-By marginalizing over its cold edges and combining the state into one equivalence class, instead 
-of computing the probability of individual network states, we are left with computing the 
-probability from the remaining edges, which is called \textit{hot edges}.
+% \section{Verification} \label{verification}
+% From a labeled graph that encodes a given network, we will then do functional verification 
+% and temporal verification in the following way:
 
-Hot edges are the opposite of cold edges: their state (whether they are up or down) will change 
-the convergent path of the network state in question.
-Mostly, they consists of the links that make up the convergent paths itself.
+% \subsection{Functional Verification}
+% For functional verification, we will essentially do the exploration technique employed by NetDice.
+% However, we made some modifications to the algorithm in order for it to remember and pass 
+% necessary informations to the temporal verification stage.
+% To contextualize our modification, we will briefly describe the description of NetDice's algorithm 
+% in three parts: Network States, Equivalence Classes, and Functional Exploration.
+% For a more complete complete description of NetDice, please refer to the original paper.
 
-By identifying the cold edges of a given network state, NetDice reduces the search-space by 
-merging multiple network states into fewer equivalence classes. 
-However, this optimization also serves another purpose: providing an efficient exploration 
-strategy.
+% \subsubsection{Network States}
+% Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of all links that 
+% are alive / up and $D$ is a set of all links that are dead / down.
+% (Consequently, $E_t = U \cup D$).
+% We then define a set $S$ that represents all possible network states in a given topology. 
+% The goal of functional verification is to identify the largest subset of $S$ where each element 
+% of this set is a network state that can connect the source and destination node given a convergent 
+% behavior of the control plane.
+% We will call this subset \textit{reachability subset}.
+% While NetDice posseses the ability to verify other properties, we only focuses on basic 
+% reachability, the most relevant property for latency verification.
 
-\subsubsection{Functional Exploration}
-To actually identify the equivalence classes that might emerge in a given network, NetDice employs
-a tree-based exploration method based on hot edges computation.
+% This framework fits nicely with the framework of probability theory, where $S$ represents the 
+% sample space and the power set of $S$ represents the possible probability events, one of which is 
+% the previously mentioned reachability subset.
+% The way we assign probability values to a subset is to simply do a summation over the probability 
+% value of each network state, which in turn is the product of all the failure and success rate of 
+% each links in the network state.
 
-We start with a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$ and $U = E_t$).
-We then compute the convergent paths and determine the hot edges set of that network state.
-Since, by definition, shutting down the link in this set will change the convergent paths, 
-instead of exhaustively exploring some other state at random, NetDice would continue the 
-exploration by systematically failing one of these links instead.
+% The naive approach to compute the probability of the reachability subset is to exhaustively iterate 
+% over all the network state and check the reachability property. 
+% While natural, this brute-force algorithm doesn't scale well since it will take 
+% $2^{|E_t|}$ iterations.
+
+% \subsubsection{Equivalence Classes}
+% NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could combine over 
+% some states that are guaranteed not to change the convergent paths (and thus the reachability 
+% property) of the current network state, by marginalizing over what is called \textit{cold edges}.
+
+% Cold edges are defined to be links whose failure is guaranteed not to change the property in 
+% question.
+% In the specific case of reachability, this essentially mean that the failure of such edge 
+% won't change the convergent paths of the network state in question.
+% By marginalizing over its cold edges and combining the state into one equivalence class, instead 
+% of computing the probability of individual network states, we are left with computing the 
+% probability from the remaining edges, which is called \textit{hot edges}.
+
+% Hot edges are the opposite of cold edges: their state (whether they are up or down) will change 
+% the convergent path of the network state in question.
+% Mostly, they consists of the links that make up the convergent paths itself.
+
+% By identifying the cold edges of a given network state, NetDice reduces the search-space by 
+% merging multiple network states into fewer equivalence classes. 
+% However, this optimization also serves another purpose: providing an efficient exploration 
+% strategy.
+
+% \subsubsection{Functional Exploration}
+% To actually identify the equivalence classes that might emerge in a given network, NetDice employs
+% a tree-based exploration method based on hot edges computation.
+
+% We start with a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$ and $U = E_t$).
+% We then compute the convergent paths and determine the hot edges set of that network state.
+% Since, by definition, shutting down the link in this set will change the convergent paths, 
+% instead of exhaustively exploring some other state at random, NetDice would continue the 
+% exploration by systematically failing one of these links instead.
 
@@
 
-NetDice will recursively continue this process for the new state and effectively form an exploration 
-tree, where each level of the tree represents how many links we will explicitly put down.
-The root is the event where we have empty hot edges, and the leaves are where the hot edges 
-are full.
+% NetDice will recursively continue this process for the new state and effectively form an exploration 
+% tree, where each level of the tree represents how many links we will explicitly put down.
+% The root is the event where we have empty hot edges, and the leaves are where the hot edges 
+% are full.
 
-\subsubsection{Making Exploration Explicit}
-After the original exploration algorithm finishes, NetDice will output the final probability (and 
-error range) of the reachability property in all the equivalence classes that has been explored. 
-However, this information alone isn't useful for us to reason about latency.
-Thus, we modified the original exploration algorithm in order to make it more explicit, so that its 
-output can be used effectively in the next stage.
+% \subsubsection{Making Exploration Explicit}
+% After the original exploration algorithm finishes, NetDice will output the final probability (and 
+% error range) of the reachability property in all the equivalence classes that has been explored. 
+% However, this information alone isn't useful for us to reason about latency.
+% Thus, we modified the original exploration algorithm in order to make it more explicit, so that its 
+% output can be used effectively in the next stage.
 
-We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
-$U_h$ and $D_h$ refers to a set of links in the hot edges that are up and down respectively.
-They are similar to $U$ and $D$ in network state, with a notable exception that 
-$E_t \neq U_h \cup D_h$, since they only represents hot edges, and not including cold edges.
-$P_{conv}$ refers to the convergent paths between src-dst pair when the links in $D_h$ fails.
-We explicitly store $P_{conv}$ in $\mathcal{E}$ (unlike the original algorithm) so that it could 
-be used to for the temporal verification stage.
+% We started by formalizing the notion of Equivalence Classes.
+% An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
+% $U_h$ and $D_h$ refers to a set of links in the hot edges that are up and down respectively.
+% They are similar to $U$ and $D$ in network state, with a notable exception that 
+% $E_t \neq U_h \cup D_h$, since they only represents hot edges, and not including cold edges.
+% $P_{conv}$ refers to the convergent paths between src-dst pair when the links in $D_h$ fails.
+% We explicitly store $P_{conv}$ in $\mathcal{E}$ (unlike the original algorithm) so that it could 
+% be used to for the temporal verification stage.
 
-We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-At the root, we have an Equivalence Class that corresponds to a perfect network, where $P_{conv}$ 
-is the path(s) that the control plane will produce given a perfect network condition and 
-$U_h = D_h = \emptyset$.
-Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $P_conv$, then it would have no children.
+% We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
+% At the root, we have an Equivalence Class that corresponds to a perfect network, where $P_{conv}$ 
+% is the path(s) that the control plane will produce given a perfect network condition and 
+% $U_h = D_h = \emptyset$.
+% Each Equivalence Class in this tree will have children which have one of its parent's link in 
+% $P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
+% If an Equivalence Class has an empty $P_conv$, then it would have no children.
 
-To compute the functional probability of a given Equivalence Class, we compute the product 
-of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the product of the 'down'
-probability of all the links in $D_h$.
+% To compute the functional probability of a given Equivalence Class, we compute the product 
+% of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the product of the 'down'
+% probability of all the links in $D_h$.
 
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
+% After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
+% stage.
 
-\subsection{Temporal Verification}
-In the temporal verification stage, we will use the $P_{conv}$ within each Equivalence Class in 
-$\mathcal{T}$ and the latency distribution labels $l_p$ and $l_q$ to augment the functional 
-probability we had in the previous stage to actually answer some questions about latency property.
+% \subsection{Temporal Verification}
+% In the temporal verification stage, we will use the $P_{conv}$ within each Equivalence Class in 
+% $\mathcal{T}$ and the latency distribution labels $l_p$ and $l_q$ to augment the functional 
+% probability we had in the previous stage to actually answer some questions about latency property.
 
-\subsubsection{Path Convolution and Weighted Average}
-Given $P_{conv}$, $l_p$, and $l_q$, we will do the following:
-\begin{enumerate}
-    \item Split $P_{conv}$ into its individual path
-    \item For each path, compute its latency distribution by convolving the latency distribution 
-        $l_p$ and $l_q$ of each link
-    \item With the resulting path latency distribution, determine the probability of temporal 
-        property by computing the statistical property of said distribution
-    \item Combine the temporal property of each path by computing the weighted average, based 
-        on the load-balancing scheme of the control plane
-\end{enumerate}
+% \subsubsection{Path Convolution and Weighted Average}
+% Given $P_{conv}$, $l_p$, and $l_q$, we will do the following:
+% \begin{enumerate}
+%     \item Split $P_{conv}$ into its individual path
+%     \item For each path, compute its latency distribution by convolving the latency distribution 
+%         $l_p$ and $l_q$ of each link
+%     \item With the resulting path latency distribution, determine the probability of temporal 
+%         property by computing the statistical property of said distribution
+%     \item Combine the temporal property of each path by computing the weighted average, based 
+%         on the load-balancing scheme of the control plane
+% \end{enumerate}
 
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
-overhead to the exploration algorithm (on top of functional exploration).
-In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-those $n$ Equivalence Class.
-We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-later.
+% We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
+% overhead to the exploration algorithm (on top of functional exploration).
+% In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
+% those $n$ Equivalence Class.
+% We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
+% later.
 

### 2022-11-17 paper/main.tex
[Paper] Updated chapter 3

@@
 
-\subsubsection{Numerical Convolution}
-One problem with this algorithm, is that not every convolution between distributions have a 
-closed-form solution and could be done analytically.
-Therefore, we need to leverage a numerical convolution method in order to convolve two arbitrary
-distributions.
+\subsubsection{Numeric Convolution}
+One potential problem with a convolution-based technique is not every distribution pair can be convolved 
+analytically.
+A closed-form solution of a convolution is usually only available for two distributions of the same type.
+Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
+method.
 
-%TODO: explain the convolution, weighting, DIRECT, changing order of DIRECT
-Expand points on 
+We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
+%TODO: describe DIRECT in one sentence
+We choose this method due to their bounded error property: DIRECT guarantee that the computed 
+distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
+DIRECT has also been implemented in R's popular bayesmeta package.
+%TODO: cite and change font?
+
+One subtle detail about DIRECT is while convolution is defined to be a commutative operation, and 
+DIRECT also achieves this property, the order of operation matters for the algorithm's runtime 
+performance.
+In particular, if we had a chain of DIRECT convolutions, the result of a convolution should not be set 
+as the second input distribution in the later convolutions to avoid performance penalty.
+This is caused by a nested loop from the following two mechanisms:
 \begin{itemize}
-    \item DIRECT 
-    \item orderings DIRECT
+    \item DIRECT works by computing a list of support values by doing many PDF queries of the second input
+        distribution
+    \item Computing the PDF of a DIRECT distribution (the returned distribution of a DIRECT 
+        algorithm) itself involves iterating over the current support values
 \end{itemize}
 
+As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
+to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
+$direct(C, direct(A, B))$.
+
+While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
+DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
+process.
+
 \subsubsection{Equivalence Class Consolidation}
-The first is what we call \textbf{Equivalence Class Consolidation}. 
-The idea is, while each hot-edges events has the same convergent paths, it is not guaranteed to be 
-unique. 
-Two or more of these events can have the same convergent paths.
-Therefore, we could merge these two events and compute the convolution for a given path once.
+To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
+existing exploration algorithm so that we could reduce the exploration space even further.
+
+In the original NetDice exploration algorithm, multiple network states were merged into an equivalence 
+class based on its cold edges (edges whose failure won't change the convergent path(s)).
+In other words, network states within the same equivalence class will have the same convergent path(s).
+
+However, we observe that each of those equivalence classes are not \textit{unique}: while network states within 
+the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
+share the exact same convergent path(s)}.
+Since we define latency distribution to marginalize over all factors other than the path, we could 
+effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
+verification once.
+
+We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
+We will only do temporal probability computation once, when we first iterate over an equivalence class that 
+has a certain convergent path(s), and we cached the result should another equivalence class with the same 
+convergent path(s) emerged.
 
 \subsubsection{Path Memoization}
+After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
+bigger, but unique equivalence classes.
+As significant as it is, we could reduce the computation cost further by drawing connections between these 
+unique equivalence classes.
+
+In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
+through multiple possible paths.
+In the context of our framework, we say that an equivalence class might have more than one convergent paths.
+Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
+independent subset.
+In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
+
+%TODO: example AB and A, only need to do reweighting
+
+Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
+individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
+on the temporal probability of a given path.
+If we were to compute the temporal probability of an equivalence class with previously cached individual 
+path, we only need to calculate the weights that corresponds to the load balancing protocol without 
+doing any convolution.
 
 \subsubsection{Convolution Grouping}
+Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
+optimization on the \textit{convolution operation itself}.
+
+While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
+in our framework, it comes with two noticeable disadvantages.
+
+The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
+comes with some error. 
+While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
+chain convolution.
+
+The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
+previous subsection, DIRECT is still generally slower than analytical convolution.
+This is expected, since a closed form solution of a convolution tipically boils the operation down into 
+a few arithmetic calculation, rather than complex operation like integration.
+
+Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
+path latency distribution.
+This is done by \textit{grouping} subsets of the path components with the same probability distribution 
+type (e.g. exponential, normal, etc.). 
+We then analytically convolve the distribution within that subset and only numerically convolve the 
+returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
 

### 2022-11-18 paper/main.tex
[Paper] Added example topology
[Paper] added two TikZ graph

@@
 \usepackage{graphicx}
+\usepackage{tikz}
+\usepackage{tikzscale}
+\usetikzlibrary{positioning}
+\usetikzlibrary{shapes.multipart}
+\usetikzlibrary{trees}
 
@@
 \begin{document}
+% \shorthandoff{=}
 \begin{abstract}
@@
 
-\section{Verifier Design}
+% \section{Verifier Design}
 We divide the problem of latency verification into two parts: 
@@
 
-\subsection{Functional Verification}
+\section{Functional Verification}
 For functional property verification, NetDice \cite{steffen2020probabilistic} 
@@
 
-\subsubsection{Topology Graph}
+\begin{figure}[h]
+    \centering
+    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
+    \include{tikz/ex}
+    % \includegraphics{ex.eps}
+    \caption{Example topology, we want to check the reachability of A to Z}
+    \label{fig:ex}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
+    \includegraphics[width=\columnwidth]{../tikz/tree}
+    % \includegraphics{ex.eps}
+    \caption{Subset of Modified Exploration Tree, $\mathcal{E}_2$ and $\mathcal{E}_3$ 
+        will get consolidated, so does $\mathcal{E}_4$ and $\mathcal{E}_5$}
+    \label{fig:tree}
+\end{figure}
+
+\subsection{Topology Graph}
 In this framework, a network is encoded in an edge-labeled directed graph 
@@
 
-\subsubsection{State Exploration}
+\subsection{State Exploration}
 A \textbf{network state} is defined as a set of failed links.
@@
 While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree}.
+\textbf{exploration tree} (Fig. \ref{fig:tree}).
 
@@
 $P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $P_conv$, then it would have no children.
+If an Equivalence Class has an empty $P_{conv}$, then it would have no children.
 
@@
 
-\subsection{Temporal Verification}
+\section{Temporal Verification}
 In the temporal verification stage, we will use the path information that we got from 
@@
 
-\subsubsection{Latency Label}
+\subsection{Latency Label}
 Similar to link failure $r$, we use edge-labeling technique to represent latency that 
@@
 
-\subsubsection{Weighted Average and Path Convolution}
+\subsection{Weighted Average and Path Convolution}
 Since a convergent behavior of the forwarding plane $P_{conv}$ might contain more than 
@@
 
-\subsubsection{Numeric Convolution}
+\subsection{Numeric Convolution}
 One potential problem with a convolution-based technique is not every distribution pair can be convolved 
@@
 
-\subsubsection{Equivalence Class Consolidation}
+\section{Optimization}
+\subsection{Equivalence Class Consolidation}
 To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
@@
 
-\subsubsection{Path Memoization}
+\subsection{Path Memoization}
 After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
@@
 
-\subsubsection{Convolution Grouping}
+\subsection{Convolution Grouping}
 Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 

### 2022-11-23 paper/main.tex
[Paper] Added running example

@@
 \usetikzlibrary{shapes.multipart}
+\usetikzlibrary{graphs}
 \usetikzlibrary{trees}
+\usetikzlibrary{quotes}
 
@@
 \begin{document}
-% \shorthandoff{=}
 \begin{abstract}
@@
 information of each components in the network.
+
+\begin{figure*}[h]
+    \centering
+    \includegraphics{../tikz/process}
+    \caption{Overview of Tempus}
+    \label{fig:process}
+\end{figure*}
 
@@
 % \section{Verifier Design}
+
 We divide the problem of latency verification into two parts: 
@@
     % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \include{tikz/ex}
+    \includegraphics{../tikz/ex}
     % \includegraphics{ex.eps}
@@
     \label{fig:ex}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/tree}
-    % \includegraphics{ex.eps}
-    \caption{Subset of Modified Exploration Tree, $\mathcal{E}_2$ and $\mathcal{E}_3$ 
-        will get consolidated, so does $\mathcal{E}_4$ and $\mathcal{E}_5$}
-    \label{fig:tree}
 \end{figure}
@@
 \rightarrow v_1$).
+A routing protocol is then defined on top of this graph by some auxilary information.
+
+To explain it with further clarity, we use the topology in Fig. \ref{fig:ex} as a 
+running example for this paper. 
+We will assume that the nodes in the network runs OSPF with weight 1 for every edge and ECMP 
+as their load-balancing scheme.
+We want to check the temporal probability of packets that departs from A to Z.
 
@@
 fashion.
+
+For the sake of example, in our Fig. \ref{fig:ex} topology, we will assume that each link has a 
+10\% chance of failure.
 
@@
 While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree} (Fig. \ref{fig:tree}).
+\textbf{exploration tree}.
 
@@
 We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
+An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
 $U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
-$P_{conv}$ refers to the convergent paths between src-dst pair that is produced by the
+$paths$ refers to the convergent paths between src-dst pair that is produced by the
 control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $P_{conv}$ in $\mathcal{E}$ so that 
+Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
 it could be used to for the temporal verification stage.
@@
 At the root, we have an Equivalence Class that corresponds to a perfect network, where 
-$P_{conv}$ is the path(s) that the control plane will produce given a perfect network 
+$paths$ is the path(s) that the control plane will produce given a perfect network 
 condition and $U_h = D_h = \emptyset$.
 Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $P_{conv}$, then it would have no children.
+$paths$ appended to its $D_h$, essentially failing one of the link in the path.
+If an Equivalence Class has an empty $paths$, then it would have no children.
 
-To compute the functional probability of a given Equivalence Class, we compute the 
-product of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the 
+To compute the functional probability $P_f$ of a given Equivalence Class, we compute the 
+product of the 'up' probability of all the links in $U_h$ and $paths$ and the 
 product of the 'down' probability of all the links in $D_h$.
+
+In our running example, we could see the subset of the exploration tree in Fig. 
+\ref{fig:tree}. 
+We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
+shortest path between A to Z.  
+The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
+which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
+
+In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
+OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
+The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$
+($0.9^6 \cdot 0.1$).
 
@@
 produce the final temporal property probability.
+
+\begin{figure}[h]
+    \centering
+    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
+    \includegraphics[width=\columnwidth]{../tikz/tree}
+    % \includegraphics{ex.eps}
+    \caption{Subset of Modified Exploration Tree $\mathcal{T}$ for our example topology, 
+        $\mathcal{E}_2$ and $\mathcal{E}_3$ will get consolidated}
+    \label{fig:tree}
+\end{figure}
 
@@
 
-Given $P_{conv}$, $l_p$, and $l_q$, we could then compute the probability of a given 
-temporal property being fulfilled.
+In our running example, we will assume that all edges in the topology will have $l_p$
+and $l_q$ of exponential distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
+have a gamma distribution with $\alpha = 1$ and $\theta = 1$.
+
+Given $paths$, $l_p$, and $l_q$, we could then compute $P_t$, the probability of a 
+given temporal property being fulfilled.
 We will describe our algorithm for computing this probability in the next few 
@@
 \subsection{Weighted Average and Path Convolution}
-Since a convergent behavior of the forwarding plane $P_{conv}$ might contain more than 
-one paths, we will compute the temporal probability of $P_{conv}$ by computing the 
-weighted average of the temporal probability of each of the individual paths.
+Since a convergent behavior of the forwarding plane $paths$ might contain more than 
+one paths, we will compute the temporal probability of $paths$ by computing the 
+weighted average of the temporal probability of each of the individual path.
 The specific weight depends on the load-balancing method used.
-We have implemented the weight computation for ECMP load-balancing scheme.
+%We have implemented the weight computation for ECMP load-balancing scheme.
 
@@
 
-In short, given $P_{conv}$, $l_p$, and $l_q$, we will compute:
+In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
+equally probable path, $ABEZ$ and $ADFZ$.
+We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
+$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
+We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
+desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.
+
+In short, given $paths$, $l_p$, and $l_q$, we will do the following:
 \begin{enumerate}
-    \item Split $P_{conv}$ into its individual path
+    \item Split $paths$ into its individual path
     \item For each path, compute its latency distribution by convolving the latency distribution 
@@
 
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
-overhead to the exploration algorithm (on top of functional exploration).
+We do each of these steps to every Equivalence Class in $\mathcal{T}$, and do a sum-product operation 
+of $P_f$ and $P_t$ to get the final probability of the property in question.
+
+By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
+exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
 In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
@@
 
-% \section{Verification} \label{verification}
-% From a labeled graph that encodes a given network, we will then do functional verification 
-% and temporal verification in the following way:
+\subsection{Numeric Convolution}
+However, there is one major problem with a convolution-based technique for computing the latency distribution 
+of a given path: not every distribution pair can be convolved analytically.
+A closed-form solution of a convolution is usually only available for two distributions of the same type.
 
-% \subsection{Functional Verification}
-% For functional verification, we will essentially do the exploration technique employed by NetDice.
-% However, we made some modifications to the algorithm in order for it to remember and pass 
-% necessary informations to the temporal verification stage.
-% To contextualize our modification, we will briefly describe the description of NetDice's algorithm 
-% in three parts: Network States, Equivalence Classes, and Functional Exploration.
-% For a more complete complete description of NetDice, please refer to the original paper.
-
-% \subsubsection{Network States}
-% Let a network state $s_i$ be defined as a 2-tuple $(U, D)$ where $U$ is a set of all links that 
-% are alive / up and $D$ is a set of all links that are dead / down.
-% (Consequently, $E_t = U \cup D$).
-% We then define a set $S$ that represents all possible network states in a given topology. 
-% The goal of functional verification is to identify the largest subset of $S$ where each element 
-% of this set is a network state that can connect the source and destination node given a convergent 
-% behavior of the control plane.
-% We will call this subset \textit{reachability subset}.
-% While NetDice posseses the ability to verify other properties, we only focuses on basic 
-% reachability, the most relevant property for latency verification.
-
-% This framework fits nicely with the framework of probability theory, where $S$ represents the 
-% sample space and the power set of $S$ represents the possible probability events, one of which is 
-% the previously mentioned reachability subset.
-% The way we assign probability values to a subset is to simply do a summation over the probability 
-% value of each network state, which in turn is the product of all the failure and success rate of 
-% each links in the network state.
-
-% The naive approach to compute the probability of the reachability subset is to exhaustively iterate 
-% over all the network state and check the reachability property. 
-% While natural, this brute-force algorithm doesn't scale well since it will take 
-% $2^{|E_t|}$ iterations.
-
-% \subsubsection{Equivalence Classes}
-% NetDice \cite{steffen2020probabilistic} solves this issue by noticing that we could combine over 
-% some states that are guaranteed not to change the convergent paths (and thus the reachability 
-% property) of the current network state, by marginalizing over what is called \textit{cold edges}.
-
-% Cold edges are defined to be links whose failure is guaranteed not to change the property in 
-% question.
-% In the specific case of reachability, this essentially mean that the failure of such edge 
-% won't change the convergent paths of the network state in question.
-% By marginalizing over its cold edges and combining the state into one equivalence class, instead 
-% of computing the probability of individual network states, we are left with computing the 
-% probability from the remaining edges, which is called \textit{hot edges}.
-
-% Hot edges are the opposite of cold edges: their state (whether they are up or down) will change 
-% the convergent path of the network state in question.
-% Mostly, they consists of the links that make up the convergent paths itself.
-
-% By identifying the cold edges of a given network state, NetDice reduces the search-space by 
-% merging multiple network states into fewer equivalence classes. 
-% However, this optimization also serves another purpose: providing an efficient exploration 
-% strategy.
-
-% \subsubsection{Functional Exploration}
-% To actually identify the equivalence classes that might emerge in a given network, NetDice employs
-% a tree-based exploration method based on hot edges computation.
-
-% We start with a perfect network state (i.e. $s_i = (U, D)$ where $D = \emptyset$ and $U = E_t$).
-% We then compute the convergent paths and determine the hot edges set of that network state.
-% Since, by definition, shutting down the link in this set will change the convergent paths, 
-% instead of exhaustively exploring some other state at random, NetDice would continue the 
-% exploration by systematically failing one of these links instead.
-
-% We then determine whether that network state can fulfill said property and which path it will 
-% take, similar to the brute-force method.
-% If no, then it is safe to say that the source node can never reach the destination node in all 
-% network state.
-% If yes, then we could use the logic of the routing protocol to determine the set of cold edges, 
-% (i.e. edges whose failure won't change the convergent paths). 
-% Edges that are not part of this set is called \textit{hot edges} $h_i = (U, D)$.
-% Instead of using $s_i$ to compute the probability of the current state, NetDice only use the set 
-% of hot edges to compute the probability of them being alive ($h_i = (U, D)$ where $D = \emptyset$).
-% This in effect computes the probability of multiple network state at once.
-
-% Next, instead of examining other network state at random, we instead iterate over the links in 
-% $h_i$.
-% By definition, if we fail a link that belongs to $h_i$, then the convergent paths will change.
-% To cover all of those possibilities efficiently, NetDice iterate over all the links in $h_i$, 
-% disabled them, re-enabled them, and move to the next edge to do the same.
-% If $h_i = ((e_1, e_2, ..., e_j), \emptyset)$, then we would enqueue $(\emptyset, (e_1))$, 
-% $((e_1), (e_2))$, $((e_1, e_2), (e_3))$, ..., $((e_1, e_2, ..., e_{j-1}), (e_j))$.
-% Thus, instead of exploring $2^j$ events, we will enqueue $j$ events instead.
-
-% NetDice will recursively continue this process for the new state and effectively form an exploration 
-% tree, where each level of the tree represents how many links we will explicitly put down.
-% The root is the event where we have empty hot edges, and the leaves are where the hot edges 
-% are full.
-
-% \subsubsection{Making Exploration Explicit}
-% After the original exploration algorithm finishes, NetDice will output the final probability (and 
-% error range) of the reachability property in all the equivalence classes that has been explored. 
-% However, this information alone isn't useful for us to reason about latency.
-% Thus, we modified the original exploration algorithm in order to make it more explicit, so that its 
-% output can be used effectively in the next stage.
-
-% We started by formalizing the notion of Equivalence Classes.
-% An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, P_{conv})$.
-% $U_h$ and $D_h$ refers to a set of links in the hot edges that are up and down respectively.
-% They are similar to $U$ and $D$ in network state, with a notable exception that 
-% $E_t \neq U_h \cup D_h$, since they only represents hot edges, and not including cold edges.
-% $P_{conv}$ refers to the convergent paths between src-dst pair when the links in $D_h$ fails.
-% We explicitly store $P_{conv}$ in $\mathcal{E}$ (unlike the original algorithm) so that it could 
-% be used to for the temporal verification stage.
-
-% We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-% At the root, we have an Equivalence Class that corresponds to a perfect network, where $P_{conv}$ 
-% is the path(s) that the control plane will produce given a perfect network condition and 
-% $U_h = D_h = \emptyset$.
-% Each Equivalence Class in this tree will have children which have one of its parent's link in 
-% $P_{conv}$ appended to its $D_h$, essentially failing one of the link in the path.
-% If an Equivalence Class has an empty $P_conv$, then it would have no children.
-
-% To compute the functional probability of a given Equivalence Class, we compute the product 
-% of the 'up' probability of all the links in $U_h$ and $P_{conv}$ and the product of the 'down'
-% probability of all the links in $D_h$.
-
-% After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-% stage.
-
-% \subsection{Temporal Verification}
-% In the temporal verification stage, we will use the $P_{conv}$ within each Equivalence Class in 
-% $\mathcal{T}$ and the latency distribution labels $l_p$ and $l_q$ to augment the functional 
-% probability we had in the previous stage to actually answer some questions about latency property.
-
-% \subsubsection{Path Convolution and Weighted Average}
-% Given $P_{conv}$, $l_p$, and $l_q$, we will do the following:
-% \begin{enumerate}
-%     \item Split $P_{conv}$ into its individual path
-%     \item For each path, compute its latency distribution by convolving the latency distribution 
-%         $l_p$ and $l_q$ of each link
-%     \item With the resulting path latency distribution, determine the probability of temporal 
-%         property by computing the statistical property of said distribution
-%     \item Combine the temporal property of each path by computing the weighted average, based 
-%         on the load-balancing scheme of the control plane
-% \end{enumerate}
-
-% We do each of these steps to every Equivalence Class in $\mathcal{T}$, and thus adding additional 
-% overhead to the exploration algorithm (on top of functional exploration).
-% In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-% those $n$ Equivalence Class.
-% We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-% later.
-
-\subsection{Numeric Convolution}
-One potential problem with a convolution-based technique is not every distribution pair can be convolved 
-analytically.
-A closed-form solution of a convolution is usually only available for two distributions of the same type.
+In our example, we define $l_q(EZ)$ to be gamma distribution, which doesn't have a closed form convolution 
+solution with the other latency distribution in the network, which is of exponential form.
 Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
@@
 
-In the original NetDice exploration algorithm, multiple network states were merged into an equivalence 
-class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same equivalence class will have the same convergent path(s).
+In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
+Class based on its cold edges (edges whose failure won't change the convergent path(s)).
+In other words, network states within the same Equivalence Class will have the same convergent path(s).
 
-However, we observe that each of those equivalence classes are not \textit{unique}: while network states within 
+However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
 the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}.
+share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
+
 Since we define latency distribution to marginalize over all factors other than the path, we could 
@@
 
-%TODO: example AB and A, only need to do reweighting
-
 Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
@@
 
+In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
+an identical path, $ADFZ$.
+We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{{E}_2}$ and just bring the 
+cached result when we explore $\mathcal{{E}_4}$.
+
+With this optimization, we will do a chain convolution by the amount of path available in the network between 
+the source and destination (in the case of Fig. \ref{fig:ex}, 3).
+
 \subsection{Convolution Grouping}
@@
 optimization on the \textit{convolution operation itself}.
+
+\begin{figure}[h]
+    \centering
+    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
+    \includegraphics[width=\columnwidth]{../tikz/group}
+    % \includegraphics{ex.eps}
+    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
+    we want to start introducing numerical convolution as late as possible}
+    \label{fig:grouping}
+\end{figure}
 
@@
 a few arithmetic calculation, rather than complex operation like integration.
-
 Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
 path latency distribution.
-This is done by \textit{grouping} subsets of the path components with the same probability distribution 
+
+We did this by \textit{grouping} subsets of the path components with the same probability distribution 
 type (e.g. exponential, normal, etc.). 
@@
 returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
+
+In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
+convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (exponential) 
+and then numerically convolve the resulting distribution with $l_q(EZ)$.
+Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
 

### 2022-11-29 paper/main.tex
[Code] Finished evaluation

@@
 In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of exponential distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a gamma distribution with $\alpha = 1$ and $\theta = 1$.
+and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
+have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
 
@@
 
-In our example, we define $l_q(EZ)$ to be gamma distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of exponential form.
+In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
+solution with the other latency distribution in the network, which is of gamma distribution form.
 Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
@@
 an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{{E}_2}$ and just bring the 
-cached result when we explore $\mathcal{{E}_4}$.
+We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
+cached result when we explore $\mathcal{E}_4$.
 
 With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (in the case of Fig. \ref{fig:ex}, 3).
+the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
+and $ADFZ$).
 
@@
 In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (exponential) 
+convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
 and then numerically convolve the resulting distribution with $l_q(EZ)$.
@@
 \section{Evaluation}
-\subsection{Scalability}
 
-\subsubsection{States Explored}
+Spec
+\begin{itemize}
+    \item Julia
+    \item i7-1065G7 16GB RAM
+\end{itemize}
+
+Network model
+\begin{itemize}
+    \item OSPF weight 1
+    \item 0.1\% link failure rate
+    \item $10^{-8}$ inaccuracy
+    \item Reachability between two edge nodes
+    \item Queuing latency: LogNormal distribution
+    \item Link latency: either Gamma or Chisq (chosen at random)
+    \item Bounded reachability with t=1000.0
+\end{itemize}
+
+\subsection{Runtime Scalability}
+
 \begin{figure}[h]
     \centering
-    \includegraphics[scale=0.5]{states}
-    \caption{States explored for temporal verification and functional verification}
-    \label{fig:states}
+    \includegraphics[scale=0.5]{scalability}
+    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
+    \label{fig:scalability}
 \end{figure}
 
-\subsubsection{Wall-clock Time}
+\begin{itemize}
+    \item The convolution step takes the most amount of time
+    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
+    \item The amount of convolution is the result of the optimizations that we've done
+\end{itemize}
+
+\subsection{Equivalence Class Reduction}
+
 \begin{figure}[h]
     \centering
-    \includegraphics[scale=0.5]{wallclock}
-    \caption{Time took for temporal verification}
-    \label{fig:wallclock}
+    \includegraphics[scale=0.5]{ec}
+    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
+    \label{fig:ec}
 \end{figure}
 
+\begin{itemize}
+    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
+    \item Corresponds to the consolidation strategy
+    \item \textbf{More than 50\% reduction}
+    \item Doesn't work equally on all topologies, depend on 
+    \begin{itemize}
+        \item How many functional EC are explored (for a given inaccuracy level)
+        \item How many possible paths a given node pair have in those ECs
+    \end{itemize}
+\end{itemize}
+
 \subsection{Optimization Effectiveness}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{optimization}
+    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
+    \label{fig:opt}
+\end{figure}
+
+\begin{itemize}
+    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
+        the amount of convolution, we also need to consider the length of each of those path
+    \item By considering the both amount of path and its length to counting the convolution, we could 
+        compare consolidation and memoization strategy to the unoptimized version
+    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
+        convolution by 98\%} (in the case of GtsCe)
+\end{itemize}
+
+\subsection{Effect of Latency Distribution Type}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{distribution}
+    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
+    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
+    \label{fig:dist}
+\end{figure}
+
+\begin{itemize}
+    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
+        that there's still some variations
+    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
+    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
+        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
+    \item In the worst case (all numerical), the grouping optimization would have no effect
+\end{itemize}
 
@@
 \section{Future Works}
+\begin{itemize}
+    \item Addressing independence: convolution treats each probability distribution as independent 
+        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
+        interesting future works
+    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
+        has already been consolidated would be interesting future works
+    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
+        latency verification, future work needs to be done in translating measurement data into latency 
+        distributions. Julia supports distribution fitting given a set of data. 
+\end{itemize}
 
 \section{Conclusion}
+\begin{itemize}
+    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
+    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
+        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
+        length.
+    \item The grouping optimization strategy is effective only if most distributions in the network can be 
+        convolved analytically. In the worst case, if most of them can't (like LogNormal only) then this 
+        optimization would have no improvement.
+\end{itemize}
 

### 2022-11-29 src/Direct.jl

@@ -65,7 +65,8 @@ function direct(d1::UnivariateDistribution, d2::UnivariateDistribution, delta::F
     while sym_kldivergence_shift(d2, step) < delta
         step = step * 2
     end
-    step = find_zero(x -> sym_kldivergence_shift(d2, x) - delta, (0, step))
+    # This line took the longest to finish
+    step = find_zero(x -> sym_kldivergence_shift(d2, x) - delta, (0.0, step), A42())
 
     # Determine grid range from epsilon
     mini = -1.0

### 2022-11-29 src/Exploration.jl

@@ -4,6 +4,8 @@ mutable struct MetaPath
 end
 
 function explore(tg::AbstractGraph, src::Symbol, dst::Symbol, threshold::Real)::Float64
+    start = now()
+
     # If src and dst are the same node
     src == dst && return 1.0
 
@@ -23,14 +25,14 @@ function explore(tg::AbstractGraph, src::Symbol, dst::Symbol, threshold::Real)::
 
     # Explore functional state
     cnt = 0
-    while !isempty(e) && p_explored < 0.999999999999
+    while !isempty(e) && p_explored < 0.99999999
         cnt += 1
         l = dequeue!(e)
 
         # Disable links of this current state
         # offlinks is undirected, offlinksdata is directed
         offlinks::Vector{Tuple{Symbol, Symbol}} = get_disabled_with_dep(st, l)
-        offlinksdata = Dict{Tuple{Symbol, Symbol}, Tuple{Real, Distribution, UInt}}()
+        offlinksdata = Dict{Tuple{Symbol, Symbol}, Tuple{Real, Distribution, Distribution, UInt}}()
         for (u, v) in offlinks
             offlinksdata[(u, v)] = tgcopy[u, v]
             rem_edge!(tgcopy, code_for(tgcopy, u), code_for(tgcopy, v))
@@ -67,19 +69,21 @@ function explore(tg::AbstractGraph, src::Symbol, dst::Symbol, threshold::Real)::
             tgcopy[key[1], key[2]] = val
         end
     end
+    println(now() - start)
 
     # Calculate p_paths_temporal of all states
     start = now()
-    p_paths_temporals = Dict{Vector{Vector{Symbol}}, Float64}()
+    p_paths_temporals = Dict{Set{Vector{Symbol}}, Float64}()
     p_path_temporals = Dict{Vector{Symbol}, Float64}()
     cnt_reexploration = 0
     cnt_path = 0
-    cnt_conv = 0
+    cnt_aconv = 0
+    cnt_nconv = 0
     for x in map(y -> label_for(st, y), vertices(st))
         isempty(st[x].converged_paths) && continue
 
         # If p_paths_temporal has not been computed before, re-explore state
-        if !haskey(p_paths_temporals, st[x].converged_paths)
+        if !haskey(p_paths_temporals, Set{Vector{Symbol}}(st[x].converged_paths))
             # Compute probability of each path being taken
             weights = ecmpprob(st[x].converged_paths)
[…]

### 2022-11-29 src/Tempus.jl

@@ -18,8 +18,6 @@ include("TopologyGraphs.jl")
 include("StateTrees.jl")
 include("Exploration.jl")
 
-# don't use 'using' or 'import' so that all globals are compartmentalized?
-
 function main()
     isempty(ARGS) && (println("Please provide a file name as an input"); return 1)
 
@@ -31,6 +29,14 @@ function main()
     tg::TopologyGraph = TopologyGraph(config["routers"], config["links"])
     prob::Float64 = explore(tg, Symbol(config["intent"]["src"]), Symbol(config["intent"]["dst"]), config["intent"]["threshold"])
     println(prob)
+
+    # start = now()
+    # x = LogNormal()
+    # y = Exponential()
+    # # z = convolve(convolve(x, x), y)
+    # z = convolve(x, x)
+    # println(now() - start)
+    # println(z)
 end
 
 if !isinteractive()

### 2022-11-29 src/TopologyGraphs.jl

@@ -1,9 +1,9 @@
 # EdgeData = (failprob, latencydist, ospfweight)
 tgweightfunction = x -> last(x)
-const TopologyGraph = typeof(MetaGraph(DiGraph(), EdgeData = Tuple{Real, Distribution, UInt}, weight_function = tgweightfunction))
+const TopologyGraph = typeof(MetaGraph(DiGraph(), EdgeData = Tuple{Real, Distribution, Distribution, UInt}, weight_function = tgweightfunction))
 
 function TopologyGraph(routers::Vector{Any}, links::Vector{Any})::TopologyGraph
-    tg::TopologyGraph = MetaGraph(DiGraph(), EdgeData = Tuple{Real, Distribution, UInt}, weight_function = tgweightfunction)
+    tg::TopologyGraph = MetaGraph(DiGraph(), EdgeData = Tuple{Real, Distribution, Distribution, UInt}, weight_function = tgweightfunction)
 
     # Make the nodes
     for router::Dict{String, Any} in routers
@@ -17,11 +17,12 @@ function TopologyGraph(routers::Vector{Any}, links::Vector{Any})::TopologyGraph
             continue
         end
 
+        # println(link)
         linkdist = gendist(link["delayModel"])
         uvdist = gendist(getdelaymodel(routers, link["u"], link["v"]))
-        tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], convolve(linkdist, uvdist), link["w_uv"])
+        tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], linkdist, uvdist, link["w_uv"])
         vudist = gendist(getdelaymodel(routers, link["v"], link["u"]))
-        tg[Symbol(link["v"]), Symbol(link["u"])] = (link["failProb"], convolve(linkdist, vudist), link["w_vu"])
+        tg[Symbol(link["v"]), Symbol(link["u"])] = (link["failProb"], linkdist, vudist, link["w_vu"])
         # Link failure should nullify both connections
     end
 
@@ -37,5 +38,9 @@ function gendist(delaymodel::Dict{String, Any})::Distribution
     args = Tuple(delaymodel["args"])
     rawdist = eval(Meta.parse(delaymodel["delayType"] * string(args)))
     # Truncate to disallow negative delay
-    return truncated(rawdist; lower=0.0)
+    if cdf(rawdist, 0.0) <= 0.0 # TODO: change this threshold
+        return rawdist 
[…]

### 2022-11-29 src/netdiceparser/NetDiceTopologyParser.jl

@@ -1,11 +1,12 @@
 module TopologyParser
 
 import JSON
+using Random
 
 function main()
     tempusConfig = Dict()
 
-    open("artifacts/Highwinds.in") do f 
+    open("artifacts/AttMpls.in") do f 
         
         r = readline(f) # Read the first line containing the number of nodes
         
@@ -21,9 +22,12 @@ function main()
                 "w_uv" => parse(Int, array[3]),
                 "w_vu" => parse(Int, array[4]),
                 "failProb" => 0.001,
-                "delayModel" => Dict(
-                    "delayType" => "Normal",
-                    "args" => [1, 0]
+                "delayModel" => rand(Bool) ? Dict(
+                    "delayType" => "Gamma",
+                    "args" => []
+                ) : Dict(
+                    "delayType" => "Chisq",
+                    "args" => [3]
                 )
             ))
         end
@@ -36,9 +40,12 @@ function main()
                 "failProb" => 0.0,
                 "outQdelayModel" => [Dict(
                     "to" => link["u"] == name ? link["v"] : link["u"],
-                    "delayModel" => Dict(
-                        "delayType" => "Normal",
-                        "args" => [1, 0]
+                    "delayModel" => rand(Bool) ? Dict(
+                        "delayType" => "Gamma",
+                        "args" => []
+                    ) : Dict(
+                        "delayType" => "Chisq",
+                        "args" => [3]
                     )
                 ) for link in filter(x -> x["u"] == name || x["v"] == name, links)]
             ))
@@ -47,8 +54,8 @@ function main()
         tempusConfig["routers"] = routers
         tempusConfig["links"] = links
         tempusConfig["intent"] = Dict(
-            "src" => "64",
-            "dst" => "68",
+            "src" => "30",
+            "dst" => "46",
             "threshold" => 1000.0
         )
 
@@ -56,7 +63,7 @@ function main()
         # println(s)
     end
 
-    open("artifacts/Highwinds.json", "w") do f
+    open("artifacts/AttMplsGammaChisq.json", "w") do f
         write(f, JSON.json(tempusConfig))
     end
 end

### 2022-11-29 paper/main.tex
[Paper] minor update to points

@@
         overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-    \item In the worst case (all numerical), the grouping optimization would have no effect
 \end{itemize}
@@
         length.
-    \item The grouping optimization strategy is effective only if most distributions in the network can be 
-        convolved analytically. In the worst case, if most of them can't (like LogNormal only) then this 
-        optimization would have no improvement.
+    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
+        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
 \end{itemize}

### 2022-12-02 paper/main.tex
[Paper] added explanation for both runtime and opt

@@
 
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
+To evaluate our idea, we implemented Tempus in $\sim600$ lines of Julia code.
+This code will take a topology configuration, source and destination node, and the 
+temporal property in question and its threshold.
+It will then output the final probability of the network upholding that temporal property.
+We ran our prototype on a machine with i7-1065G7 CPU and 16 GB of RAM.
 
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
+We tested Tempus on 5 real-world WAN topologies from the Topology Zoo and 3 datacenter topologies 
+(fat-tree, with various amount of pods).
+We use OSPF as the routing protocol with uniform link weights of 1 and 0.1\% failure rate in all 
+links, in accordance to the failure rate in previous studies. %TODO: cite
+We then run the functional verification with $10^{-8}$ inaccuracy level.
+
+To evaluate the performance and scalability of Tempus, we first evaluate the running time of the 
+temporal verification on different topologies given a mixture of latency distributions. 
+We then examine the effectiveness of our optimization technique by comparing the number of 
+convolution of the former results with the unoptimized version.
+Next, we compare the result of the consolidation strategy in terms of the number of equivalence 
+classes, instead of convolution.
+Finally, we compare the effect of latency distribution type on the performance of Tempus.
+
+Since the CDF operation that we define for temporal property verification is a relatively cheap 
+operation, we expect our result to generalize with any threshold we chose. 
+In other words, while changing the threshold will change the final probability, it won't change 
+the runtime performance of Tempus.
+Thus, we set an arbitrary threshold for our evaluation.
 
@@
 
+To begin our evaluation, we first define the $l_p$ and $l_q$ for each component in 
+the topology with varying probability distributions to ensure the robustness of our method.
+Some distributions (like LogNormal) don't have a closed-form solution for their convolution, even 
+against itself, while others (e.g. Gamma, Chi-Squared) do.
+Since our algorithm relies on convolution, having a diversity of distributions, especially accross
+these categories, is important for the representativeness of our evaluation.
+In light of this, we set all the queuing latency $l_q$ of our topologies' component to be LogNormal 
+and the link latency $l_p$ to be randomly chosen between Gamma and Chi-Squared.
+
+We measure the runtime performance of Tempus by verifying the bounded reachability property of two 
+edge routers.
+Fig. \ref{fig:scalability} shows the temporal verification runtime performance (excluding functional 
+verification) and the amount of convolution operation performed in each network.
+We note that we can precisely verify the temporal property of networks with hundreds of links in 
+the order of minutes.
+
+The convolution step of Tempus takes the most amount of time in the temporal verification process, 
+hence we can see in the same figure that the runtime performance of Tempus and the total amount 
+of convolution is highly correlated.
+
+We conclude that \textbf{the performance of Tempus is primarily bottlenecked by the convolution 
+operation} and for a reasonably low imprecision level, \textbf{Tempus runs in the order of 
+minutes}.
+
+\subsection{Optimization Effectiveness}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{optimization}
+    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
+    \label{fig:opt}
+\end{figure}
+
+We have established that the convolution operation is a relatively expensive operation within 
+Tempus.
+Now, we inspect the effectiveness of our optimization techniques in reducing them.
+
+Fig. \ref{fig:opt} shows the effect of the two optimization techniques, consolidation and 
+memoization, in reducing the amount of convolution (in logarithmic scale).
+We exclude the third optimization technique (grouping) in this evaluation because they are not 
+designed to reduce the amount of convolution and thus need to be evaluated differently in later 
+subsection.
+
+We could see that in the best case, the optimization strategies reduce the amount of 
+convolution by \textbf{98\%} (in the case of GtsCe).
+
+%TODO: factors of the network node-pair? path amount and length?
+
 \begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
+    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
+        the amount of convolution, we also need to consider the length of each of those path
+    \item By considering the both amount of path and its length to counting the convolution, we could 
+        compare consolidation and memoization strategy to the unoptimized version
+    \item 
 \end{itemize}
@@
     \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
 \end{itemize}
@@
 \section{Related Works}
+\begin{itemize}
+    \item Control plane and data plane verifier
+    \item Probabilistic Verifier 
+    \item Quantitative Verifier
+\end{itemize}
 

### 2023-01-18 paper/1_intro.tex
[Paper] Updated section 7, 8

--- 2022-11-29 paper/main.tex
@@
-\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{tikzscale}
-\usetikzlibrary{positioning}
-\usetikzlibrary{shapes.multipart}
-\usetikzlibrary{graphs}
-\usetikzlibrary{trees}
-\usetikzlibrary{quotes}
-
-\graphicspath{ {./graph/} }
-
-\title{Tempus: Probabilistic Network Latency Verifier}
-\author{WIP}
-
-\begin{document}
-\begin{abstract}
-    To combat problems and bugs that a given network might have, network verifiers have emerged as 
-    one of the promising solutions. 
-    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under k-link failure. 
-    However, as modern networks evolved and performance need becomes more stringent -- often 
-    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
-    verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
-    side of the network performance metric: bandwidth and load violation properties. 
-    Questions about the other side of network performance metric, latency, were left unanswered.
-
-    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
-    a given temporal property being true given latency distributions of individual links and 
-    nodes in the network. 
-    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
-    that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
-    qualitative verifier it's built upon.
-\end{abstract}
-
-\maketitle
-
 \section{Introduction}
@@
 
-Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
+Popular control plane verifiers, like ARC \cite{arc}, are formulated to answer 
 questions about a given property deterministically. 
@@
 \end{figure*}
-
-\section{Overview}
-% \section{On Latency Modeling}
-\subsection{Latency as Path Property} 
-We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
-a single packet.
-In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
-through the many components of the network (e.g. links, routers, firewall) and each of those 
-components might introduce some latency into the packet transmission process.
-Figuring out which of those components will actually introduce latency into the packet in question, 
-and by how much, is the next logical step that we must figure out.
-
-Obviously, a given packet doesn't need to visit all network components to reach its destination 
-host.
-A network engineer will configure the network in such a way that a packet will only need to be 
-routed via a specific subset of its components.
-That specific subset is dictated by two things: how those components are connected together 
-(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
-protocol, ACL).
-Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
-given packet appropriately.
-
-The goal of a classical control plane verifier then, is to use this forwarding table in some 
-form to verify certain properties. %TODO: examples?
-However, looking at the forwarding table alone will not give us a conclusive result regarding 
-which components are going to be visited by a given packet, making verification of latency 
-properties less clear.
-For example, the network might be configured with a load-balancing protocol in which a packet 
-departing from a source host might take multiple different \textit{paths} (with certain 
-probability weights) to arrive at the destination host, possibly resulting in a different 
-end-to-end latency measurement.
-
-Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
-the primary unit of reasoning, rather than forwarding table.
-By being more specific about our unit of reasoning, we could answer verification questions more 
-clearly, and we could design our verifier more efficiently since we could use it to represent 
-multiple different forwarding tables that shares the same path.
-
-\subsection{Relation to Classical Verifier}
-Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that said path exists in the first place.
-We note that latency is a property that only make sense after connectivity between two hosts 
-has been established. 
-In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
-link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
-making the verification task trivial.
-
-Fortunately, there are a rich body of work in the network verification literature regarding 
-functional reachability under failure. %TODO: cite
-We could then design our verifier on top of an existing classical control plane verifier.
-We use it to verify reachability property, and only if the reachability property is fulfilled, 
-we would verify whether the latency between two hosts fulfilled some additional condition.
-
-\subsection{Path Latency Distribution}
-Up to this point, we have talked about measuring the latency of a single packet by figuring out 
-its path; analyzing which exact components it has traversed through.
-However, when we try to generalize this framework and ask questions about the latency of multiple 
-packets, it is apparent that the path alone is not a determinative information, as the latency 
-of two packets propagating through the same path might be different due to a multitude of factors.
-
-The natural extension to the framework then, is instead of representing latency of a path with a 
-single number, we instead represent it with a continuous random variable that signifies the 
-possible delay that a given packet traversing through that path might have.
-This random variable will have a distribution that marginalizes over all other factors other than 
-the path.
-
-We can then use this latency distribution to verify some temporal properties in a probabilistic 
-manner.
-For example, we could verify the probability that a packet will be delivered in under a time 
-unit by taking the CDF of the distribution.
-
-\subsection{Path Decomposition and Convolution} \label{decomposition}
-The final question that we had to decide on was how do we actually model path latency distribution
-based on real component measurements.
-Considering the complexity of factors that might determine the path latency distribution and the 
-availability of measurement data, we settle on the assumption that the path latency distribution 
-is composed of multiple independent distribution that corresponds to the latency each components 
-in the path might introduce.
-%TODO: confirm about independence?
-
-Since not all components in the path will introduce latency with a non-negligible value, we 
-specifically choose to model two source of latency in the path's components which we deem 
-significant: \textbf{link propagation latency} and \textbf{queuing latency}.
-%TODO: expand on how to get the data later
-
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
-
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
-
-In order to obtain the overall path latency distribution from these per-component latency 
-distributions, we do a convolution operations over all the relevant component's latency 
-distributions.
-Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
-convolution technique with a guaranteed error bound. 
-We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
-the lack of a general technique to measure error made us opt for the formerly mentioned method.
-
-% \section{Verifier Design}
-
-We divide the problem of latency verification into two parts: 
-verifying that two nodes are reachable (\textbf{functional verification}) and only if 
-the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some temporal condition (\textbf{temporal verification}).
-
-\section{Functional Verification}
-For functional property verification, NetDice \cite{steffen2020probabilistic} 
-has laid the way for verifying reachability between two nodes under failure in a 
-probabilistic and efficient manner.
-To contextualize our modification, we will briefly describe some parts of 
-NetDice's network encoding and exploration algorithm as a prelude to our alteration.
-For a more complete complete description of NetDice, please refer to the original 
-paper.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics{../tikz/ex}
-    % \includegraphics{ex.eps}
-    \caption{Example topology, we want to check the reachability of A to Z}
-    \label{fig:ex}
-\end{figure}
-
-\subsection{Topology Graph}
-In this framework, a network is encoded in an edge-labeled directed graph 
-$G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
-$E_t$ represents the functional connectivity between a source and a destination node. 
-A physical link is then represented as a pair of symmetrical edges that shares the same source 
-and destination node but with opposite direction (e.g. $v_1 \rightarrow v_2$ and $v_2 
-\rightarrow v_1$).
-A routing protocol is then defined on top of this graph by some auxilary information.
-
-To explain it with further clarity, we use the topology in Fig. \ref{fig:ex} as a 
-running example for this paper. 
-We will assume that the nodes in the network runs OSPF with weight 1 for every edge and ECMP 
-as their load-balancing scheme.
-We want to check the temporal probability of packets that departs from A to Z.
-
-% To represent the component's random failure, NetDice defined two failure rates.
-% The first failure rate is the chance that a given physical link in the network will go down.
-% The second failure rate is the chance that a given node in the network will go down.
-% These rates are shared between all links / nodes in the network.
-% Internally however, the node failure rate will be translated into the link failure rate 
-% with a Bayesian Network model, since a node failure can be modeled with the failure of each 
-% links connected to it.
-
-To represent random failure, NetDice defined a universal link failure rate (i.e. 
-the probability of \textit{any} link in the network randomly failing).
-We refine NetDice's model slightly by allowing each links to have different failure rate.
-We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
-\le 1\}$ that represents the failure rate of a given physical link.
-As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
-opposite direction) will also share the same failure rate, and will be disabled in a coupled 
-fashion.
-
-For the sake of example, in our Fig. \ref{fig:ex} topology, we will assume that each link has a 
-10\% chance of failure.
-
-% \subsubsection{Routing}
-% On top of this topology, NetDice also defined additional informations that would be used
-% by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
-% a particular link failure scenario. 
-% These routing informations would also be used by their optimization algorithms to reduce the 
-% amount of states that are going to be explored.
-
-% NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
-% They encode the relevant routing information by assigning some labels to the vertices or edges
-% in the topology.
-% In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
-% that represents the positive weight of a link.
-% They would later be used to compute the convergent paths of a given network state 
-% (\ref{verification}).
-
-\subsection{State Exploration}
-A \textbf{network state} is defined as a set of failed links.
-There are $2^{|E_t|}$ network states in a given network and a brute-force strategy 
-would need to explore all of the states to compute the reachability property of a 
-node pair in a given network. 
-NetDice however, will merge multiple network states into an \textbf{equivalence class}
-by marginalizing over \textit{cold edges}, links whose failure is guaranteed not to 
-change the convergent path of a highlighted network state.
-
-NetDice will systematically explore many equivalence classes in the network. 
-It will start from the equivalence class of a perfect network and will fail certain 
-\textit{hot edges} (i.e. edges that are not cold) to explore another equivalence 
-class.
-While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree}.
-
-In order to preserve some information that will be used in the temporal verification 
-stage, we modify this original exploration algorithm to make it more explicit.
-We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
-$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
-$paths$ refers to the convergent paths between src-dst pair that is produced by the
-control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
-it could be used to for the temporal verification stage.
-
-We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-At the root, we have an Equivalence Class that corresponds to a perfect network, where 
-$paths$ is the path(s) that the control plane will produce given a perfect network 
-condition and $U_h = D_h = \emptyset$.
-Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$paths$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $paths$, then it would have no children.
-
-To compute the functional probability $P_f$ of a given Equivalence Class, we compute the 
-product of the 'up' probability of all the links in $U_h$ and $paths$ and the 
-product of the 'down' probability of all the links in $D_h$.
-
-In our running example, we could see the subset of the exploration tree in Fig. 
-\ref{fig:tree}. 
-We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
-shortest path between A to Z.  
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
-which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
-
-In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
-OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$
-($0.9^6 \cdot 0.1$).
-
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
-
-\section{Temporal Verification}
-In the temporal verification stage, we will use the path information that we got from 
-functional verification stage and augment them with latency information to 
-produce the final temporal property probability.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/tree}
-    % \includegraphics{ex.eps}
-    \caption{Subset of Modified Exploration Tree $\mathcal{T}$ for our example topology, 
-        $\mathcal{E}_2$ and $\mathcal{E}_3$ will get consolidated}
-    \label{fig:tree}
-\end{figure}
-
-\subsection{Latency Label}
-Similar to link failure $r$, we use edge-labeling technique to represent latency that 
-will get introduced by the connectivity between two nodes. 
-Our model will model two significant sources of latency: \textbf{link propagation} 
-and \textbf{packet queuing in the node}.
-We will encode these latencies by equipping the model with two additional labels.
-
-Encoding link propagation latency is fairly straightforward. 
-We define a function $l_p: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$ is a set of 
-continuous univariate distribution that has a minimum value of $0$.
-This distribution signifies the time it would take for the respective physical link to transmit 
-a packet from one end to another.
-Because of this, just like $r$, two symmetrical edges will share the same distribution.
-
-To encode queuing latency, we first note that the queuing mechanism in modern switches 
-usually resides in the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively 
-assign the latency to the connectivity between switches.
-To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
-This distribution signifies the output queue latency in the source node that is 
-going to be forwarded to the destination node.
-Unlike $l_p$, two symmetrical edges will have two different distributions since they 
-represent two different output queues.
-
-In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
-
-Given $paths$, $l_p$, and $l_q$, we could then compute $P_t$, the probability of a 
-given temporal property being fulfilled.
-We will describe our algorithm for computing this probability in the next few 
-sections.
-
-\subsection{Weighted Average and Path Convolution}
-Since a convergent behavior of the forwarding plane $paths$ might contain more than 
-one paths, we will compute the temporal probability of $paths$ by computing the 
-weighted average of the temporal probability of each of the individual path.
-The specific weight depends on the load-balancing method used.
-%We have implemented the weight computation for ECMP load-balancing scheme.
-
-To get the temporal probability of a single path, we will use the latency information 
-from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
-To do this, we resort to the methods of \textbf{convolution}.
-For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
-all together to get the path latency distribution.
-
-With this path latency distribution, we could get the temporal probability of a 
-path by computing the statistical property of said distribution.
-In this work, given a path latency distribution $\mathcal{L}$, we define two temporal 
-properties:
-\begin{itemize}
-    \item \textbf{Bounded Reachability}: the probability that a packet will get 
-        transmited below $t$ time unit. This will get computed as $cdf(\mathcal{L}, t)$
-    \item \textbf{Tail Reachability}: the probability that a packet will get 
-        transmited above $t$ time unit. This will get computed as 
-        $1 - cdf(\mathcal{L}, t)$
-\end{itemize}
-
-In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
-equally probable path, $ABEZ$ and $ADFZ$.
-We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
-$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
-We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
-desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.
-
-In short, given $paths$, $l_p$, and $l_q$, we will do the following:
-\begin{enumerate}
-    \item Split $paths$ into its individual path
-    \item For each path, compute its latency distribution by convolving the latency distribution 
-        $l_p$ and $l_q$ of each link
-    \item With the resulting path latency distribution, determine the probability of temporal 
-        property by computing the statistical property of said distribution
-    \item Combine the temporal property of each path by computing the weighted average, based 
-        on the load-balancing scheme of the control plane
-\end{enumerate}
-
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and do a sum-product operation 
-of $P_f$ and $P_t$ to get the final probability of the property in question.
-
-By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
-exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
-In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-those $n$ Equivalence Class.
-We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-later.
-
-\subsection{Numeric Convolution}
-However, there is one major problem with a convolution-based technique for computing the latency distribution 
-of a given path: not every distribution pair can be convolved analytically.
-A closed-form solution of a convolution is usually only available for two distributions of the same type.
-
-In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of gamma distribution form.
-Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
-method.
-
-We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
-%TODO: describe DIRECT in one sentence
-We choose this method due to their bounded error property: DIRECT guarantee that the computed 
-distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
-DIRECT has also been implemented in R's popular bayesmeta package.
-%TODO: cite and change font?
-
-One subtle detail about DIRECT is while convolution is defined to be a commutative operation, and 
-DIRECT also achieves this property, the order of operation matters for the algorithm's runtime 
-performance.
-In particular, if we had a chain of DIRECT convolutions, the result of a convolution should not be set 
-as the second input distribution in the later convolutions to avoid performance penalty.
-This is caused by a nested loop from the following two mechanisms:
-\begin{itemize}
-    \item DIRECT works by computing a list of support values by doing many PDF queries of the second input
-        distribution
-    \item Computing the PDF of a DIRECT distribution (the returned distribution of a DIRECT 
-        algorithm) itself involves iterating over the current support values
-\end{itemize}
-
-As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
-to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
-$direct(C, direct(A, B))$.
-
-While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
-DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
-process.
-
-\section{Optimization}
-\subsection{Equivalence Class Consolidation}
-To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
-existing exploration algorithm so that we could reduce the exploration space even further.
-
-In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
-Class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same Equivalence Class will have the same convergent path(s).
-
-However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
-the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
-
-Since we define latency distribution to marginalize over all factors other than the path, we could 
-effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
-verification once.
-
-We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
-We will only do temporal probability computation once, when we first iterate over an equivalence class that 
-has a certain convergent path(s), and we cached the result should another equivalence class with the same 
-convergent path(s) emerged.
-
-\subsection{Path Memoization}
-After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
-bigger, but unique equivalence classes.
-As significant as it is, we could reduce the computation cost further by drawing connections between these 
-unique equivalence classes.
-
-In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
-through multiple possible paths.
-In the context of our framework, we say that an equivalence class might have more than one convergent paths.
-Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
-independent subset.
-In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
-
-Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
-individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
-on the temporal probability of a given path.
-If we were to compute the temporal probability of an equivalence class with previously cached individual 
-path, we only need to calculate the weights that corresponds to the load balancing protocol without 
-doing any convolution.
-
-In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
-cached result when we explore $\mathcal{E}_4$.
-
-With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
-
-\subsection{Convolution Grouping}
-Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
-optimization on the \textit{convolution operation itself}.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/group}
-    % \includegraphics{ex.eps}
-    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
-    we want to start introducing numerical convolution as late as possible}
-    \label{fig:grouping}
-\end{figure}
-
-While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
-in our framework, it comes with two noticeable disadvantages.
-
-The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
-comes with some error. 
-While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
-chain convolution.
-
-The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
-previous subsection, DIRECT is still generally slower than analytical convolution.
-This is expected, since a closed form solution of a convolution tipically boils the operation down into 
-a few arithmetic calculation, rather than complex operation like integration.
-Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
-path latency distribution.
-
-We did this by \textit{grouping} subsets of the path components with the same probability distribution 
-type (e.g. exponential, normal, etc.). 
-We then analytically convolve the distribution within that subset and only numerically convolve the 
-returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
-
-In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
-and then numerically convolve the resulting distribution with $l_q(EZ)$.
-Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
-
-% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{steffen2020probabilistic} have done. 
-% In addition, we also collect additional information from each state (e.g. convergent 
-% paths) to be used in the temporal verification step.
-
-% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-% state with $G_l$.
-% This is done by convolving over the latency distribution of each component in the path. 
-% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-% distribution with a KL divergence error bound.
-% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
-% explored}.
-
-% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
-% traverse from a source to destination below some time unit $T$), we can compute its probability by 
-% integrating the PDF from $0$ to $T$. 
-% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
-% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
-
-% TODO: intro, reformalize graph formulation, citations
-% Does a figure about topology graph and latency graph will be useful?
-% Multimedia context on older networked system Klara Nahrstedts
-% Graph
-
-\section{Evaluation}
-
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
-
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
-
-\subsection{Runtime Scalability}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{scalability}
-    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
-    \label{fig:scalability}
-\end{figure}
-
-\begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
-\end{itemize}
-
-\subsection{Equivalence Class Reduction}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{ec}
-    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
-    \label{fig:ec}
-\end{figure}
-
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
-\end{itemize}
-
-\subsection{Effect of Latency Distribution Type}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{distribution}
-    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-    \label{fig:dist}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
-
-\section{Related Works}
-
-\section{Future Works}
-\begin{itemize}
-    \item Addressing independence: convolution treats each probability distribution as independent 
-        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
-        interesting future works
-    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
-        has already been consolidated would be interesting future works
-    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
-        latency verification, future work needs to be done in translating measurement data into latency 
-        distributions. Julia supports distribution fitting given a set of data. 
-\end{itemize}
-
-\section{Conclusion}
-\begin{itemize}
-    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
-    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
-        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
-        length.
-    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
-        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
-\end{itemize}
-
-\bibliography{cit}
-\bibliographystyle{ieeetr}
-\end{document}

### 2023-01-18 paper/2_overview.tex
[Paper] Updated section 7, 8

--- 2022-11-29 paper/main.tex
@@
-\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{tikzscale}
-\usetikzlibrary{positioning}
-\usetikzlibrary{shapes.multipart}
-\usetikzlibrary{graphs}
-\usetikzlibrary{trees}
-\usetikzlibrary{quotes}
-
-\graphicspath{ {./graph/} }
-
-\title{Tempus: Probabilistic Network Latency Verifier}
-\author{WIP}
-
-\begin{document}
-\begin{abstract}
-    To combat problems and bugs that a given network might have, network verifiers have emerged as 
-    one of the promising solutions. 
-    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under k-link failure. 
-    However, as modern networks evolved and performance need becomes more stringent -- often 
-    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
-    verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
-    side of the network performance metric: bandwidth and load violation properties. 
-    Questions about the other side of network performance metric, latency, were left unanswered.
-
-    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
-    a given temporal property being true given latency distributions of individual links and 
-    nodes in the network. 
-    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
-    that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
-    qualitative verifier it's built upon.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
-routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a component failure.
-Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
-It is hard for a network engineer to reason about whether their configured network fulfilled some 
-intended property in various possible states of the network. 
-Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network control plane verifier.
-
-Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
-questions about a given property deterministically. 
-In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about the property (e.g. two nodes are reachable).
-While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate small fraction of failure. 
-For example, a given network might have an availability SLA of 99.999\%. 
-This kind of probabilistic properties have been the focus of a more recent works like NetDice 
-\cite{steffen2020probabilistic}.
-
-The network property itself can also be divided into two kinds. 
-Quantitative properties like reachability and loop existence, are the common properties that are 
-studied by most of the existing verifiers. 
-More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
-properties like link bandwidth violation for a given traffic.
-
-In this work, we're exploring the other side of the network performance metric: latency. 
-Certain network deployment often necessitates some latency requirement such as an ISP that has 
-latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
-We proposed a verification framework to probabilistically verify the latency property of packets 
-traversing from a source to destination node under various failure scenarios, by using latency 
-information of each components in the network.
-
-\begin{figure*}[h]
-    \centering
-    \includegraphics{../tikz/process}
-    \caption{Overview of Tempus}
-    \label{fig:process}
-\end{figure*}
-
 \section{Overview}
@@
 two nodes fulfilled some temporal condition (\textbf{temporal verification}).
-
-\section{Functional Verification}
-For functional property verification, NetDice \cite{steffen2020probabilistic} 
-has laid the way for verifying reachability between two nodes under failure in a 
-probabilistic and efficient manner.
-To contextualize our modification, we will briefly describe some parts of 
-NetDice's network encoding and exploration algorithm as a prelude to our alteration.
-For a more complete complete description of NetDice, please refer to the original 
-paper.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics{../tikz/ex}
-    % \includegraphics{ex.eps}
-    \caption{Example topology, we want to check the reachability of A to Z}
-    \label{fig:ex}
-\end{figure}
-
-\subsection{Topology Graph}
-In this framework, a network is encoded in an edge-labeled directed graph 
-$G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
-$E_t$ represents the functional connectivity between a source and a destination node. 
-A physical link is then represented as a pair of symmetrical edges that shares the same source 
-and destination node but with opposite direction (e.g. $v_1 \rightarrow v_2$ and $v_2 
-\rightarrow v_1$).
-A routing protocol is then defined on top of this graph by some auxilary information.
-
-To explain it with further clarity, we use the topology in Fig. \ref{fig:ex} as a 
-running example for this paper. 
-We will assume that the nodes in the network runs OSPF with weight 1 for every edge and ECMP 
-as their load-balancing scheme.
-We want to check the temporal probability of packets that departs from A to Z.
-
-% To represent the component's random failure, NetDice defined two failure rates.
-% The first failure rate is the chance that a given physical link in the network will go down.
-% The second failure rate is the chance that a given node in the network will go down.
-% These rates are shared between all links / nodes in the network.
-% Internally however, the node failure rate will be translated into the link failure rate 
-% with a Bayesian Network model, since a node failure can be modeled with the failure of each 
-% links connected to it.
-
-To represent random failure, NetDice defined a universal link failure rate (i.e. 
-the probability of \textit{any} link in the network randomly failing).
-We refine NetDice's model slightly by allowing each links to have different failure rate.
-We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
-\le 1\}$ that represents the failure rate of a given physical link.
-As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
-opposite direction) will also share the same failure rate, and will be disabled in a coupled 
-fashion.
-
-For the sake of example, in our Fig. \ref{fig:ex} topology, we will assume that each link has a 
-10\% chance of failure.
-
-% \subsubsection{Routing}
-% On top of this topology, NetDice also defined additional informations that would be used
-% by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
-% a particular link failure scenario. 
-% These routing informations would also be used by their optimization algorithms to reduce the 
-% amount of states that are going to be explored.
-
-% NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
-% They encode the relevant routing information by assigning some labels to the vertices or edges
-% in the topology.
-% In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
-% that represents the positive weight of a link.
-% They would later be used to compute the convergent paths of a given network state 
-% (\ref{verification}).
-
-\subsection{State Exploration}
-A \textbf{network state} is defined as a set of failed links.
-There are $2^{|E_t|}$ network states in a given network and a brute-force strategy 
-would need to explore all of the states to compute the reachability property of a 
-node pair in a given network. 
-NetDice however, will merge multiple network states into an \textbf{equivalence class}
-by marginalizing over \textit{cold edges}, links whose failure is guaranteed not to 
-change the convergent path of a highlighted network state.
-
-NetDice will systematically explore many equivalence classes in the network. 
-It will start from the equivalence class of a perfect network and will fail certain 
-\textit{hot edges} (i.e. edges that are not cold) to explore another equivalence 
-class.
-While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree}.
-
-In order to preserve some information that will be used in the temporal verification 
-stage, we modify this original exploration algorithm to make it more explicit.
-We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
-$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
-$paths$ refers to the convergent paths between src-dst pair that is produced by the
-control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
-it could be used to for the temporal verification stage.
-
-We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-At the root, we have an Equivalence Class that corresponds to a perfect network, where 
-$paths$ is the path(s) that the control plane will produce given a perfect network 
-condition and $U_h = D_h = \emptyset$.
-Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$paths$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $paths$, then it would have no children.
-
-To compute the functional probability $P_f$ of a given Equivalence Class, we compute the 
-product of the 'up' probability of all the links in $U_h$ and $paths$ and the 
-product of the 'down' probability of all the links in $D_h$.
-
-In our running example, we could see the subset of the exploration tree in Fig. 
-\ref{fig:tree}. 
-We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
-shortest path between A to Z.  
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
-which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
-
-In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
-OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$
-($0.9^6 \cdot 0.1$).
-
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
-
-\section{Temporal Verification}
-In the temporal verification stage, we will use the path information that we got from 
-functional verification stage and augment them with latency information to 
-produce the final temporal property probability.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/tree}
-    % \includegraphics{ex.eps}
-    \caption{Subset of Modified Exploration Tree $\mathcal{T}$ for our example topology, 
-        $\mathcal{E}_2$ and $\mathcal{E}_3$ will get consolidated}
-    \label{fig:tree}
-\end{figure}
-
-\subsection{Latency Label}
-Similar to link failure $r$, we use edge-labeling technique to represent latency that 
-will get introduced by the connectivity between two nodes. 
-Our model will model two significant sources of latency: \textbf{link propagation} 
-and \textbf{packet queuing in the node}.
-We will encode these latencies by equipping the model with two additional labels.
-
-Encoding link propagation latency is fairly straightforward. 
-We define a function $l_p: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$ is a set of 
-continuous univariate distribution that has a minimum value of $0$.
-This distribution signifies the time it would take for the respective physical link to transmit 
-a packet from one end to another.
-Because of this, just like $r$, two symmetrical edges will share the same distribution.
-
-To encode queuing latency, we first note that the queuing mechanism in modern switches 
-usually resides in the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively 
-assign the latency to the connectivity between switches.
-To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
-This distribution signifies the output queue latency in the source node that is 
-going to be forwarded to the destination node.
-Unlike $l_p$, two symmetrical edges will have two different distributions since they 
-represent two different output queues.
-
-In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
-
-Given $paths$, $l_p$, and $l_q$, we could then compute $P_t$, the probability of a 
-given temporal property being fulfilled.
-We will describe our algorithm for computing this probability in the next few 
-sections.
-
-\subsection{Weighted Average and Path Convolution}
-Since a convergent behavior of the forwarding plane $paths$ might contain more than 
-one paths, we will compute the temporal probability of $paths$ by computing the 
-weighted average of the temporal probability of each of the individual path.
-The specific weight depends on the load-balancing method used.
-%We have implemented the weight computation for ECMP load-balancing scheme.
-
-To get the temporal probability of a single path, we will use the latency information 
-from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
-To do this, we resort to the methods of \textbf{convolution}.
-For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
-all together to get the path latency distribution.
-
-With this path latency distribution, we could get the temporal probability of a 
-path by computing the statistical property of said distribution.
-In this work, given a path latency distribution $\mathcal{L}$, we define two temporal 
-properties:
-\begin{itemize}
-    \item \textbf{Bounded Reachability}: the probability that a packet will get 
-        transmited below $t$ time unit. This will get computed as $cdf(\mathcal{L}, t)$
-    \item \textbf{Tail Reachability}: the probability that a packet will get 
-        transmited above $t$ time unit. This will get computed as 
-        $1 - cdf(\mathcal{L}, t)$
-\end{itemize}
-
-In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
-equally probable path, $ABEZ$ and $ADFZ$.
-We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
-$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
-We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
-desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.
-
-In short, given $paths$, $l_p$, and $l_q$, we will do the following:
-\begin{enumerate}
-    \item Split $paths$ into its individual path
-    \item For each path, compute its latency distribution by convolving the latency distribution 
-        $l_p$ and $l_q$ of each link
-    \item With the resulting path latency distribution, determine the probability of temporal 
-        property by computing the statistical property of said distribution
-    \item Combine the temporal property of each path by computing the weighted average, based 
-        on the load-balancing scheme of the control plane
-\end{enumerate}
-
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and do a sum-product operation 
-of $P_f$ and $P_t$ to get the final probability of the property in question.
-
-By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
-exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
-In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-those $n$ Equivalence Class.
-We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-later.
-
-\subsection{Numeric Convolution}
-However, there is one major problem with a convolution-based technique for computing the latency distribution 
-of a given path: not every distribution pair can be convolved analytically.
-A closed-form solution of a convolution is usually only available for two distributions of the same type.
-
-In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of gamma distribution form.
-Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
-method.
-
-We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
-%TODO: describe DIRECT in one sentence
-We choose this method due to their bounded error property: DIRECT guarantee that the computed 
-distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
-DIRECT has also been implemented in R's popular bayesmeta package.
-%TODO: cite and change font?
-
-One subtle detail about DIRECT is while convolution is defined to be a commutative operation, and 
-DIRECT also achieves this property, the order of operation matters for the algorithm's runtime 
-performance.
-In particular, if we had a chain of DIRECT convolutions, the result of a convolution should not be set 
-as the second input distribution in the later convolutions to avoid performance penalty.
-This is caused by a nested loop from the following two mechanisms:
-\begin{itemize}
-    \item DIRECT works by computing a list of support values by doing many PDF queries of the second input
-        distribution
-    \item Computing the PDF of a DIRECT distribution (the returned distribution of a DIRECT 
-        algorithm) itself involves iterating over the current support values
-\end{itemize}
-
-As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
-to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
-$direct(C, direct(A, B))$.
-
-While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
-DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
-process.
-
-\section{Optimization}
-\subsection{Equivalence Class Consolidation}
-To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
-existing exploration algorithm so that we could reduce the exploration space even further.
-
-In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
-Class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same Equivalence Class will have the same convergent path(s).
-
-However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
-the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
-
-Since we define latency distribution to marginalize over all factors other than the path, we could 
-effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
-verification once.
-
-We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
-We will only do temporal probability computation once, when we first iterate over an equivalence class that 
-has a certain convergent path(s), and we cached the result should another equivalence class with the same 
-convergent path(s) emerged.
-
-\subsection{Path Memoization}
-After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
-bigger, but unique equivalence classes.
-As significant as it is, we could reduce the computation cost further by drawing connections between these 
-unique equivalence classes.
-
-In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
-through multiple possible paths.
-In the context of our framework, we say that an equivalence class might have more than one convergent paths.
-Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
-independent subset.
-In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
-
-Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
-individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
-on the temporal probability of a given path.
-If we were to compute the temporal probability of an equivalence class with previously cached individual 
-path, we only need to calculate the weights that corresponds to the load balancing protocol without 
-doing any convolution.
-
-In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
-cached result when we explore $\mathcal{E}_4$.
-
-With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
-
-\subsection{Convolution Grouping}
-Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
-optimization on the \textit{convolution operation itself}.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/group}
-    % \includegraphics{ex.eps}
-    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
-    we want to start introducing numerical convolution as late as possible}
-    \label{fig:grouping}
-\end{figure}
-
-While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
-in our framework, it comes with two noticeable disadvantages.
-
-The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
-comes with some error. 
-While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
-chain convolution.
-
-The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
-previous subsection, DIRECT is still generally slower than analytical convolution.
-This is expected, since a closed form solution of a convolution tipically boils the operation down into 
-a few arithmetic calculation, rather than complex operation like integration.
-Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
-path latency distribution.
-
-We did this by \textit{grouping} subsets of the path components with the same probability distribution 
-type (e.g. exponential, normal, etc.). 
-We then analytically convolve the distribution within that subset and only numerically convolve the 
-returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
-
-In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
-and then numerically convolve the resulting distribution with $l_q(EZ)$.
-Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
-
-% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{steffen2020probabilistic} have done. 
-% In addition, we also collect additional information from each state (e.g. convergent 
-% paths) to be used in the temporal verification step.
-
-% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-% state with $G_l$.
-% This is done by convolving over the latency distribution of each component in the path. 
-% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-% distribution with a KL divergence error bound.
-% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
-% explored}.
-
-% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
-% traverse from a source to destination below some time unit $T$), we can compute its probability by 
-% integrating the PDF from $0$ to $T$. 
-% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
-% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
-
-% TODO: intro, reformalize graph formulation, citations
-% Does a figure about topology graph and latency graph will be useful?
-% Multimedia context on older networked system Klara Nahrstedts
-% Graph
-
-\section{Evaluation}
-
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
-
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
-
-\subsection{Runtime Scalability}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{scalability}
-    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
-    \label{fig:scalability}
-\end{figure}
-
-\begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
-\end{itemize}
-
-\subsection{Equivalence Class Reduction}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{ec}
-    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
-    \label{fig:ec}
-\end{figure}
-
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
-\end{itemize}
-
-\subsection{Effect of Latency Distribution Type}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{distribution}
-    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-    \label{fig:dist}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
-
-\section{Related Works}
-
-\section{Future Works}
-\begin{itemize}
-    \item Addressing independence: convolution treats each probability distribution as independent 
-        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
-        interesting future works
-    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
-        has already been consolidated would be interesting future works
-    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
-        latency verification, future work needs to be done in translating measurement data into latency 
-        distributions. Julia supports distribution fitting given a set of data. 
-\end{itemize}
-
-\section{Conclusion}
-\begin{itemize}
-    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
-    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
-        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
-        length.
-    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
-        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
-\end{itemize}
-
-\bibliography{cit}
-\bibliographystyle{ieeetr}
-\end{document}

### 2023-01-18 paper/3_functional.tex
[Paper] Updated section 7, 8

--- 2022-11-29 paper/main.tex
@@
-\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{tikzscale}
-\usetikzlibrary{positioning}
-\usetikzlibrary{shapes.multipart}
-\usetikzlibrary{graphs}
-\usetikzlibrary{trees}
-\usetikzlibrary{quotes}
-
-\graphicspath{ {./graph/} }
-
-\title{Tempus: Probabilistic Network Latency Verifier}
-\author{WIP}
-
-\begin{document}
-\begin{abstract}
-    To combat problems and bugs that a given network might have, network verifiers have emerged as 
-    one of the promising solutions. 
-    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under k-link failure. 
-    However, as modern networks evolved and performance need becomes more stringent -- often 
-    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
-    verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
-    side of the network performance metric: bandwidth and load violation properties. 
-    Questions about the other side of network performance metric, latency, were left unanswered.
-
-    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
-    a given temporal property being true given latency distributions of individual links and 
-    nodes in the network. 
-    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
-    that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
-    qualitative verifier it's built upon.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
-routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a component failure.
-Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
-It is hard for a network engineer to reason about whether their configured network fulfilled some 
-intended property in various possible states of the network. 
-Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network control plane verifier.
-
-Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
-questions about a given property deterministically. 
-In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about the property (e.g. two nodes are reachable).
-While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate small fraction of failure. 
-For example, a given network might have an availability SLA of 99.999\%. 
-This kind of probabilistic properties have been the focus of a more recent works like NetDice 
-\cite{steffen2020probabilistic}.
-
-The network property itself can also be divided into two kinds. 
-Quantitative properties like reachability and loop existence, are the common properties that are 
-studied by most of the existing verifiers. 
-More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
-properties like link bandwidth violation for a given traffic.
-
-In this work, we're exploring the other side of the network performance metric: latency. 
-Certain network deployment often necessitates some latency requirement such as an ISP that has 
-latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
-We proposed a verification framework to probabilistically verify the latency property of packets 
-traversing from a source to destination node under various failure scenarios, by using latency 
-information of each components in the network.
-
-\begin{figure*}[h]
-    \centering
-    \includegraphics{../tikz/process}
-    \caption{Overview of Tempus}
-    \label{fig:process}
-\end{figure*}
-
-\section{Overview}
-% \section{On Latency Modeling}
-\subsection{Latency as Path Property} 
-We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
-a single packet.
-In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
-through the many components of the network (e.g. links, routers, firewall) and each of those 
-components might introduce some latency into the packet transmission process.
-Figuring out which of those components will actually introduce latency into the packet in question, 
-and by how much, is the next logical step that we must figure out.
-
-Obviously, a given packet doesn't need to visit all network components to reach its destination 
-host.
-A network engineer will configure the network in such a way that a packet will only need to be 
-routed via a specific subset of its components.
-That specific subset is dictated by two things: how those components are connected together 
-(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
-protocol, ACL).
-Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
-given packet appropriately.
-
-The goal of a classical control plane verifier then, is to use this forwarding table in some 
-form to verify certain properties. %TODO: examples?
-However, looking at the forwarding table alone will not give us a conclusive result regarding 
-which components are going to be visited by a given packet, making verification of latency 
-properties less clear.
-For example, the network might be configured with a load-balancing protocol in which a packet 
-departing from a source host might take multiple different \textit{paths} (with certain 
-probability weights) to arrive at the destination host, possibly resulting in a different 
-end-to-end latency measurement.
-
-Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
-the primary unit of reasoning, rather than forwarding table.
-By being more specific about our unit of reasoning, we could answer verification questions more 
-clearly, and we could design our verifier more efficiently since we could use it to represent 
-multiple different forwarding tables that shares the same path.
-
-\subsection{Relation to Classical Verifier}
-Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that said path exists in the first place.
-We note that latency is a property that only make sense after connectivity between two hosts 
-has been established. 
-In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
-link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
-making the verification task trivial.
-
-Fortunately, there are a rich body of work in the network verification literature regarding 
-functional reachability under failure. %TODO: cite
-We could then design our verifier on top of an existing classical control plane verifier.
-We use it to verify reachability property, and only if the reachability property is fulfilled, 
-we would verify whether the latency between two hosts fulfilled some additional condition.
-
-\subsection{Path Latency Distribution}
-Up to this point, we have talked about measuring the latency of a single packet by figuring out 
-its path; analyzing which exact components it has traversed through.
-However, when we try to generalize this framework and ask questions about the latency of multiple 
-packets, it is apparent that the path alone is not a determinative information, as the latency 
-of two packets propagating through the same path might be different due to a multitude of factors.
-
-The natural extension to the framework then, is instead of representing latency of a path with a 
-single number, we instead represent it with a continuous random variable that signifies the 
-possible delay that a given packet traversing through that path might have.
-This random variable will have a distribution that marginalizes over all other factors other than 
-the path.
-
-We can then use this latency distribution to verify some temporal properties in a probabilistic 
-manner.
-For example, we could verify the probability that a packet will be delivered in under a time 
-unit by taking the CDF of the distribution.
-
-\subsection{Path Decomposition and Convolution} \label{decomposition}
-The final question that we had to decide on was how do we actually model path latency distribution
-based on real component measurements.
-Considering the complexity of factors that might determine the path latency distribution and the 
-availability of measurement data, we settle on the assumption that the path latency distribution 
-is composed of multiple independent distribution that corresponds to the latency each components 
-in the path might introduce.
-%TODO: confirm about independence?
-
-Since not all components in the path will introduce latency with a non-negligible value, we 
-specifically choose to model two source of latency in the path's components which we deem 
-significant: \textbf{link propagation latency} and \textbf{queuing latency}.
-%TODO: expand on how to get the data later
-
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
-
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
-
-In order to obtain the overall path latency distribution from these per-component latency 
-distributions, we do a convolution operations over all the relevant component's latency 
-distributions.
-Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
-convolution technique with a guaranteed error bound. 
-We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
-the lack of a general technique to measure error made us opt for the formerly mentioned method.
-
-% \section{Verifier Design}
-
-We divide the problem of latency verification into two parts: 
-verifying that two nodes are reachable (\textbf{functional verification}) and only if 
-the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some temporal condition (\textbf{temporal verification}).
-
 \section{Functional Verification}
@@
 stage.
-
-\section{Temporal Verification}
-In the temporal verification stage, we will use the path information that we got from 
-functional verification stage and augment them with latency information to 
-produce the final temporal property probability.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/tree}
-    % \includegraphics{ex.eps}
-    \caption{Subset of Modified Exploration Tree $\mathcal{T}$ for our example topology, 
-        $\mathcal{E}_2$ and $\mathcal{E}_3$ will get consolidated}
-    \label{fig:tree}
-\end{figure}
-
-\subsection{Latency Label}
-Similar to link failure $r$, we use edge-labeling technique to represent latency that 
-will get introduced by the connectivity between two nodes. 
-Our model will model two significant sources of latency: \textbf{link propagation} 
-and \textbf{packet queuing in the node}.
-We will encode these latencies by equipping the model with two additional labels.
-
-Encoding link propagation latency is fairly straightforward. 
-We define a function $l_p: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$ is a set of 
-continuous univariate distribution that has a minimum value of $0$.
-This distribution signifies the time it would take for the respective physical link to transmit 
-a packet from one end to another.
-Because of this, just like $r$, two symmetrical edges will share the same distribution.
-
-To encode queuing latency, we first note that the queuing mechanism in modern switches 
-usually resides in the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively 
-assign the latency to the connectivity between switches.
-To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
-This distribution signifies the output queue latency in the source node that is 
-going to be forwarded to the destination node.
-Unlike $l_p$, two symmetrical edges will have two different distributions since they 
-represent two different output queues.
-
-In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
-
-Given $paths$, $l_p$, and $l_q$, we could then compute $P_t$, the probability of a 
-given temporal property being fulfilled.
-We will describe our algorithm for computing this probability in the next few 
-sections.
-
-\subsection{Weighted Average and Path Convolution}
-Since a convergent behavior of the forwarding plane $paths$ might contain more than 
-one paths, we will compute the temporal probability of $paths$ by computing the 
-weighted average of the temporal probability of each of the individual path.
-The specific weight depends on the load-balancing method used.
-%We have implemented the weight computation for ECMP load-balancing scheme.
-
-To get the temporal probability of a single path, we will use the latency information 
-from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
-To do this, we resort to the methods of \textbf{convolution}.
-For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
-all together to get the path latency distribution.
-
-With this path latency distribution, we could get the temporal probability of a 
-path by computing the statistical property of said distribution.
-In this work, given a path latency distribution $\mathcal{L}$, we define two temporal 
-properties:
-\begin{itemize}
-    \item \textbf{Bounded Reachability}: the probability that a packet will get 
-        transmited below $t$ time unit. This will get computed as $cdf(\mathcal{L}, t)$
-    \item \textbf{Tail Reachability}: the probability that a packet will get 
-        transmited above $t$ time unit. This will get computed as 
-        $1 - cdf(\mathcal{L}, t)$
-\end{itemize}
-
-In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
-equally probable path, $ABEZ$ and $ADFZ$.
-We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
-$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
-We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
-desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.
-
-In short, given $paths$, $l_p$, and $l_q$, we will do the following:
-\begin{enumerate}
-    \item Split $paths$ into its individual path
-    \item For each path, compute its latency distribution by convolving the latency distribution 
-        $l_p$ and $l_q$ of each link
-    \item With the resulting path latency distribution, determine the probability of temporal 
-        property by computing the statistical property of said distribution
-    \item Combine the temporal property of each path by computing the weighted average, based 
-        on the load-balancing scheme of the control plane
-\end{enumerate}
-
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and do a sum-product operation 
-of $P_f$ and $P_t$ to get the final probability of the property in question.
-
-By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
-exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
-In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-those $n$ Equivalence Class.
-We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-later.
-
-\subsection{Numeric Convolution}
-However, there is one major problem with a convolution-based technique for computing the latency distribution 
-of a given path: not every distribution pair can be convolved analytically.
-A closed-form solution of a convolution is usually only available for two distributions of the same type.
-
-In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of gamma distribution form.
-Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
-method.
-
-We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
-%TODO: describe DIRECT in one sentence
-We choose this method due to their bounded error property: DIRECT guarantee that the computed 
-distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
-DIRECT has also been implemented in R's popular bayesmeta package.
-%TODO: cite and change font?
-
-One subtle detail about DIRECT is while convolution is defined to be a commutative operation, and 
-DIRECT also achieves this property, the order of operation matters for the algorithm's runtime 
-performance.
-In particular, if we had a chain of DIRECT convolutions, the result of a convolution should not be set 
-as the second input distribution in the later convolutions to avoid performance penalty.
-This is caused by a nested loop from the following two mechanisms:
-\begin{itemize}
-    \item DIRECT works by computing a list of support values by doing many PDF queries of the second input
-        distribution
-    \item Computing the PDF of a DIRECT distribution (the returned distribution of a DIRECT 
-        algorithm) itself involves iterating over the current support values
-\end{itemize}
-
-As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
-to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
-$direct(C, direct(A, B))$.
-
-While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
-DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
-process.
-
-\section{Optimization}
-\subsection{Equivalence Class Consolidation}
-To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
-existing exploration algorithm so that we could reduce the exploration space even further.
-
-In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
-Class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same Equivalence Class will have the same convergent path(s).
-
-However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
-the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
-
-Since we define latency distribution to marginalize over all factors other than the path, we could 
-effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
-verification once.
-
-We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
-We will only do temporal probability computation once, when we first iterate over an equivalence class that 
-has a certain convergent path(s), and we cached the result should another equivalence class with the same 
-convergent path(s) emerged.
-
-\subsection{Path Memoization}
-After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
-bigger, but unique equivalence classes.
-As significant as it is, we could reduce the computation cost further by drawing connections between these 
-unique equivalence classes.
-
-In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
-through multiple possible paths.
-In the context of our framework, we say that an equivalence class might have more than one convergent paths.
-Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
-independent subset.
-In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
-
-Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
-individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
-on the temporal probability of a given path.
-If we were to compute the temporal probability of an equivalence class with previously cached individual 
-path, we only need to calculate the weights that corresponds to the load balancing protocol without 
-doing any convolution.
-
-In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
-cached result when we explore $\mathcal{E}_4$.
-
-With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
-
-\subsection{Convolution Grouping}
-Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
-optimization on the \textit{convolution operation itself}.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/group}
-    % \includegraphics{ex.eps}
-    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
-    we want to start introducing numerical convolution as late as possible}
-    \label{fig:grouping}
-\end{figure}
-
-While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
-in our framework, it comes with two noticeable disadvantages.
-
-The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
-comes with some error. 
-While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
-chain convolution.
-
-The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
-previous subsection, DIRECT is still generally slower than analytical convolution.
-This is expected, since a closed form solution of a convolution tipically boils the operation down into 
-a few arithmetic calculation, rather than complex operation like integration.
-Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
-path latency distribution.
-
-We did this by \textit{grouping} subsets of the path components with the same probability distribution 
-type (e.g. exponential, normal, etc.). 
-We then analytically convolve the distribution within that subset and only numerically convolve the 
-returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
-
-In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
-and then numerically convolve the resulting distribution with $l_q(EZ)$.
-Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
-
-% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{steffen2020probabilistic} have done. 
-% In addition, we also collect additional information from each state (e.g. convergent 
-% paths) to be used in the temporal verification step.
-
-% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-% state with $G_l$.
-% This is done by convolving over the latency distribution of each component in the path. 
-% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-% distribution with a KL divergence error bound.
-% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
-% explored}.
-
-% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
-% traverse from a source to destination below some time unit $T$), we can compute its probability by 
-% integrating the PDF from $0$ to $T$. 
-% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
-% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
-
-% TODO: intro, reformalize graph formulation, citations
-% Does a figure about topology graph and latency graph will be useful?
-% Multimedia context on older networked system Klara Nahrstedts
-% Graph
-
-\section{Evaluation}
-
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
-
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
-
-\subsection{Runtime Scalability}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{scalability}
-    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
-    \label{fig:scalability}
-\end{figure}
-
-\begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
-\end{itemize}
-
-\subsection{Equivalence Class Reduction}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{ec}
-    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
-    \label{fig:ec}
-\end{figure}
-
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
-\end{itemize}
-
-\subsection{Effect of Latency Distribution Type}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{distribution}
-    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-    \label{fig:dist}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
-
-\section{Related Works}
-
-\section{Future Works}
-\begin{itemize}
-    \item Addressing independence: convolution treats each probability distribution as independent 
-        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
-        interesting future works
-    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
-        has already been consolidated would be interesting future works
-    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
-        latency verification, future work needs to be done in translating measurement data into latency 
-        distributions. Julia supports distribution fitting given a set of data. 
-\end{itemize}
-
-\section{Conclusion}
-\begin{itemize}
-    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
-    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
-        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
-        length.
-    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
-        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
-\end{itemize}
-
-\bibliography{cit}
-\bibliographystyle{ieeetr}
-\end{document}

### 2023-01-18 paper/4_temporal.tex
[Paper] Updated section 7, 8

--- 2022-11-29 paper/main.tex
@@
-\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}
-
-\usepackage{amsfonts}
-\usepackage{graphicx}
-\usepackage{tikz}
-\usepackage{tikzscale}
-\usetikzlibrary{positioning}
-\usetikzlibrary{shapes.multipart}
-\usetikzlibrary{graphs}
-\usetikzlibrary{trees}
-\usetikzlibrary{quotes}
-
-\graphicspath{ {./graph/} }
-
-\title{Tempus: Probabilistic Network Latency Verifier}
-\author{WIP}
-
-\begin{document}
-\begin{abstract}
-    To combat problems and bugs that a given network might have, network verifiers have emerged as 
-    one of the promising solutions. 
-    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under k-link failure. 
-    However, as modern networks evolved and performance need becomes more stringent -- often 
-    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
-    verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
-    side of the network performance metric: bandwidth and load violation properties. 
-    Questions about the other side of network performance metric, latency, were left unanswered.
-
-    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
-    a given temporal property being true given latency distributions of individual links and 
-    nodes in the network. 
-    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
-    that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
-    qualitative verifier it's built upon.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
-routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a component failure.
-Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
-It is hard for a network engineer to reason about whether their configured network fulfilled some 
-intended property in various possible states of the network. 
-Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network control plane verifier.
-
-Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
-questions about a given property deterministically. 
-In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about the property (e.g. two nodes are reachable).
-While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate small fraction of failure. 
-For example, a given network might have an availability SLA of 99.999\%. 
-This kind of probabilistic properties have been the focus of a more recent works like NetDice 
-\cite{steffen2020probabilistic}.
-
-The network property itself can also be divided into two kinds. 
-Quantitative properties like reachability and loop existence, are the common properties that are 
-studied by most of the existing verifiers. 
-More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
-properties like link bandwidth violation for a given traffic.
-
-In this work, we're exploring the other side of the network performance metric: latency. 
-Certain network deployment often necessitates some latency requirement such as an ISP that has 
-latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
-We proposed a verification framework to probabilistically verify the latency property of packets 
-traversing from a source to destination node under various failure scenarios, by using latency 
-information of each components in the network.
-
-\begin{figure*}[h]
-    \centering
-    \includegraphics{../tikz/process}
-    \caption{Overview of Tempus}
-    \label{fig:process}
-\end{figure*}
-
-\section{Overview}
-% \section{On Latency Modeling}
-\subsection{Latency as Path Property} 
-We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
-a single packet.
-In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
-through the many components of the network (e.g. links, routers, firewall) and each of those 
-components might introduce some latency into the packet transmission process.
-Figuring out which of those components will actually introduce latency into the packet in question, 
-and by how much, is the next logical step that we must figure out.
-
-Obviously, a given packet doesn't need to visit all network components to reach its destination 
-host.
-A network engineer will configure the network in such a way that a packet will only need to be 
-routed via a specific subset of its components.
-That specific subset is dictated by two things: how those components are connected together 
-(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
-protocol, ACL).
-Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
-given packet appropriately.
-
-The goal of a classical control plane verifier then, is to use this forwarding table in some 
-form to verify certain properties. %TODO: examples?
-However, looking at the forwarding table alone will not give us a conclusive result regarding 
-which components are going to be visited by a given packet, making verification of latency 
-properties less clear.
-For example, the network might be configured with a load-balancing protocol in which a packet 
-departing from a source host might take multiple different \textit{paths} (with certain 
-probability weights) to arrive at the destination host, possibly resulting in a different 
-end-to-end latency measurement.
-
-Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
-the primary unit of reasoning, rather than forwarding table.
-By being more specific about our unit of reasoning, we could answer verification questions more 
-clearly, and we could design our verifier more efficiently since we could use it to represent 
-multiple different forwarding tables that shares the same path.
-
-\subsection{Relation to Classical Verifier}
-Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that said path exists in the first place.
-We note that latency is a property that only make sense after connectivity between two hosts 
-has been established. 
-In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
-link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
-making the verification task trivial.
-
-Fortunately, there are a rich body of work in the network verification literature regarding 
-functional reachability under failure. %TODO: cite
-We could then design our verifier on top of an existing classical control plane verifier.
-We use it to verify reachability property, and only if the reachability property is fulfilled, 
-we would verify whether the latency between two hosts fulfilled some additional condition.
-
-\subsection{Path Latency Distribution}
-Up to this point, we have talked about measuring the latency of a single packet by figuring out 
-its path; analyzing which exact components it has traversed through.
-However, when we try to generalize this framework and ask questions about the latency of multiple 
-packets, it is apparent that the path alone is not a determinative information, as the latency 
-of two packets propagating through the same path might be different due to a multitude of factors.
-
-The natural extension to the framework then, is instead of representing latency of a path with a 
-single number, we instead represent it with a continuous random variable that signifies the 
-possible delay that a given packet traversing through that path might have.
-This random variable will have a distribution that marginalizes over all other factors other than 
-the path.
-
-We can then use this latency distribution to verify some temporal properties in a probabilistic 
-manner.
-For example, we could verify the probability that a packet will be delivered in under a time 
-unit by taking the CDF of the distribution.
-
-\subsection{Path Decomposition and Convolution} \label{decomposition}
-The final question that we had to decide on was how do we actually model path latency distribution
-based on real component measurements.
-Considering the complexity of factors that might determine the path latency distribution and the 
-availability of measurement data, we settle on the assumption that the path latency distribution 
-is composed of multiple independent distribution that corresponds to the latency each components 
-in the path might introduce.
-%TODO: confirm about independence?
-
-Since not all components in the path will introduce latency with a non-negligible value, we 
-specifically choose to model two source of latency in the path's components which we deem 
-significant: \textbf{link propagation latency} and \textbf{queuing latency}.
-%TODO: expand on how to get the data later
-
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
-
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
-
-In order to obtain the overall path latency distribution from these per-component latency 
-distributions, we do a convolution operations over all the relevant component's latency 
-distributions.
-Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
-convolution technique with a guaranteed error bound. 
-We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
-the lack of a general technique to measure error made us opt for the formerly mentioned method.
-
-% \section{Verifier Design}
-
-We divide the problem of latency verification into two parts: 
-verifying that two nodes are reachable (\textbf{functional verification}) and only if 
-the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some temporal condition (\textbf{temporal verification}).
-
-\section{Functional Verification}
-For functional property verification, NetDice \cite{steffen2020probabilistic} 
-has laid the way for verifying reachability between two nodes under failure in a 
-probabilistic and efficient manner.
-To contextualize our modification, we will briefly describe some parts of 
-NetDice's network encoding and exploration algorithm as a prelude to our alteration.
-For a more complete complete description of NetDice, please refer to the original 
-paper.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics{../tikz/ex}
-    % \includegraphics{ex.eps}
-    \caption{Example topology, we want to check the reachability of A to Z}
-    \label{fig:ex}
-\end{figure}
-
-\subsection{Topology Graph}
-In this framework, a network is encoded in an edge-labeled directed graph 
-$G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
-$E_t$ represents the functional connectivity between a source and a destination node. 
-A physical link is then represented as a pair of symmetrical edges that shares the same source 
-and destination node but with opposite direction (e.g. $v_1 \rightarrow v_2$ and $v_2 
-\rightarrow v_1$).
-A routing protocol is then defined on top of this graph by some auxilary information.
-
-To explain it with further clarity, we use the topology in Fig. \ref{fig:ex} as a 
-running example for this paper. 
-We will assume that the nodes in the network runs OSPF with weight 1 for every edge and ECMP 
-as their load-balancing scheme.
-We want to check the temporal probability of packets that departs from A to Z.
-
-% To represent the component's random failure, NetDice defined two failure rates.
-% The first failure rate is the chance that a given physical link in the network will go down.
-% The second failure rate is the chance that a given node in the network will go down.
-% These rates are shared between all links / nodes in the network.
-% Internally however, the node failure rate will be translated into the link failure rate 
-% with a Bayesian Network model, since a node failure can be modeled with the failure of each 
-% links connected to it.
-
-To represent random failure, NetDice defined a universal link failure rate (i.e. 
-the probability of \textit{any} link in the network randomly failing).
-We refine NetDice's model slightly by allowing each links to have different failure rate.
-We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
-\le 1\}$ that represents the failure rate of a given physical link.
-As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
-opposite direction) will also share the same failure rate, and will be disabled in a coupled 
-fashion.
-
-For the sake of example, in our Fig. \ref{fig:ex} topology, we will assume that each link has a 
-10\% chance of failure.
-
-% \subsubsection{Routing}
-% On top of this topology, NetDice also defined additional informations that would be used
-% by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
-% a particular link failure scenario. 
-% These routing informations would also be used by their optimization algorithms to reduce the 
-% amount of states that are going to be explored.
-
-% NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
-% They encode the relevant routing information by assigning some labels to the vertices or edges
-% in the topology.
-% In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
-% that represents the positive weight of a link.
-% They would later be used to compute the convergent paths of a given network state 
-% (\ref{verification}).
-
-\subsection{State Exploration}
-A \textbf{network state} is defined as a set of failed links.
-There are $2^{|E_t|}$ network states in a given network and a brute-force strategy 
-would need to explore all of the states to compute the reachability property of a 
-node pair in a given network. 
-NetDice however, will merge multiple network states into an \textbf{equivalence class}
-by marginalizing over \textit{cold edges}, links whose failure is guaranteed not to 
-change the convergent path of a highlighted network state.
-
-NetDice will systematically explore many equivalence classes in the network. 
-It will start from the equivalence class of a perfect network and will fail certain 
-\textit{hot edges} (i.e. edges that are not cold) to explore another equivalence 
-class.
-While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree}.
-
-In order to preserve some information that will be used in the temporal verification 
-stage, we modify this original exploration algorithm to make it more explicit.
-We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
-$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
-$paths$ refers to the convergent paths between src-dst pair that is produced by the
-control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
-it could be used to for the temporal verification stage.
-
-We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-At the root, we have an Equivalence Class that corresponds to a perfect network, where 
-$paths$ is the path(s) that the control plane will produce given a perfect network 
-condition and $U_h = D_h = \emptyset$.
-Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$paths$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $paths$, then it would have no children.
-
-To compute the functional probability $P_f$ of a given Equivalence Class, we compute the 
-product of the 'up' probability of all the links in $U_h$ and $paths$ and the 
-product of the 'down' probability of all the links in $D_h$.
-
-In our running example, we could see the subset of the exploration tree in Fig. 
-\ref{fig:tree}. 
-We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
-shortest path between A to Z.  
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
-which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
-
-In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
-OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$
-($0.9^6 \cdot 0.1$).
-
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
-
 \section{Temporal Verification}
@@
 % Graph
-
-\section{Evaluation}
-
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
-
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
-
-\subsection{Runtime Scalability}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{scalability}
-    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
-    \label{fig:scalability}
-\end{figure}
-
-\begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
-\end{itemize}
-
-\subsection{Equivalence Class Reduction}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{ec}
-    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
-    \label{fig:ec}
-\end{figure}
-
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
-\end{itemize}
-
-\subsection{Effect of Latency Distribution Type}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{distribution}
-    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-    \label{fig:dist}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
-
-\section{Related Works}
-
-\section{Future Works}
-\begin{itemize}
-    \item Addressing independence: convolution treats each probability distribution as independent 
-        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
-        interesting future works
-    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
-        has already been consolidated would be interesting future works
-    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
-        latency verification, future work needs to be done in translating measurement data into latency 
-        distributions. Julia supports distribution fitting given a set of data. 
-\end{itemize}
-
-\section{Conclusion}
-\begin{itemize}
-    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
-    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
-        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
-        length.
-    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
-        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
-\end{itemize}
-
-\bibliography{cit}
-\bibliographystyle{ieeetr}
-\end{document}

### 2023-01-18 paper/5_eval.tex
[Paper] Updated section 7, 8

\section{Evaluation}

We implemented Tempus in $\sim600$ lines of Julia \cite{julia} code.
This code will take a topology configuration, source and destination node, and the 
temporal property in question and its threshold.
It will then output the final probability of the network upholding that temporal property.
We ran our prototype on a machine with 2 Intel E5-2630 v3 8-core CPU running up to 2.4 GHz 
and 128 GB of ECC RAM, provided by CloudLab \cite{cloudlab}.

To appraise the performance and scalability of Tempus, we seek to answer 5 evaluation questions:
\begin{enumerate}
    \item How long does it take to run the verifier overall? How does it scale?
    \item What is the bottleneck step in the verification process?
    \item How effective is the optimization technique in reducing the verification time?
    \item How effective is the optimization technique in reducing the explored equivalence classes?
    \item How does different topologies affect optimization effectiveness?
\end{enumerate}

\subsection{Experiment Setup}
We tested Tempus on 5 real-world WAN topologies from the Topology Zoo and 3 datacenter topologies 
(fat-tree, with various amount of pods).
%TODO: stats of the topologies

\subsubsection{Functional Verification}
As verifying the routing protocol functional behavior is not the primary contribution of our research, 
we simply use OSPF as the routing protocol in our experiment with uniform link weights of 1.
We set 0.1\% failure rate in all links, in accordance to the failure rate in previous studies. %TODO: cite
We then run the functional verification with $10^{-8}$ inaccuracy level.

In order for our evaluation to be realistic, we set the source and destination node to be one of the 
edge routers (node with the smallest degree / smallest centrality?).
%TODO: edge-to-edge routers

\subsubsection{Temporal Verification}
For latency distribution, we show that we can use an empirical measurement as a distribution by 
using the reported queue length from DCTCP \cite{dctcp}.
By multiplying the queue length distribution with the line-rate, we could approximate the queuing 
latency distribution for a given router.

Since the CDF operation that we define for temporal property verification is relatively cheap, 
we expect our result to generalize with any threshold we chose. 
In other words, while changing the threshold will change the final probability, it won't change 
the runtime performance of Tempus.
Thus, we set an arbitrary threshold for our evaluation.

\subsection{Runtime Profiling}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{scalability}
    \caption{Runtime performance of the verification process, red dots represent the number of convolutions}
    \label{fig:scalability}
\end{figure}

\subsubsection{Performance and Scalability}
To begin our evaluation, we first measured the running duration of the verification process (and its steps) 
on various topologies.
Our results in Fig. \ref{fig:scalability} shows that our verification technique finished within a reasonable timeframe.
The verifier finished in the order of minutes, even for topologies with hundred of nodes.

From the same result, we also note that our verifier scales gracefully over the size of the network.
The verification duration ranges from 6 seconds to around 15 minutes. 
% TODO: fat-tree scalability
To drive this point more precisely, we also evaluate the running time of fat-tree topology in various sizes.
Our results in Fig. X shows that by keeping the same type of topology and scaling them up, we increased 
our verification time almost linearly?

\subsubsection{Bottleneck}
By looking at the proportion of time spent on each step in Fig. \ref{fig:scalability}, we could see that 
the combined \textbf{convolution procedure is the bottleneck} step in our verifier.
The convolution procedure takes 95\% - 99\% of the duration of the overall verification.
Hence we can see in the same figure that the runtime performance of Tempus and the total amount 
of convolution is highly correlated.

We conclude that for a reasonably low imprecision level, \textbf{Tempus runs in the order of 
minutes} and the performance of Tempus is primarily \textbf{bottlenecked by the total convolution 
operations}.

\subsection{Optimization Effectiveness}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{optimization}
    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
    \label{fig:opt}
\end{figure}

We have established that the convolution procedure is a relatively expensive operation within 
our verifier.
Next, we inspected the effectiveness of our optimization techniques in reducing them.
To do that, we measured the overall running duration of the verifier while selectively disabling the 
optimization.

Our results in Fig. \ref{fig:opt} shows the effect of the two optimization techniques we introduced -- 
consolidation and memoization -- in reducing the verification duration and amount of convolution (both in log scale).
For baseline, we ran the verifier without any optimization, resulting in the left bar.
The middle bar represents the result where we enable only the consolidation strategy.
Finally, the right bar represents the result where we enable both the consolidation and memoization strategy.

We note that for most of the topologies, the combination of both optimization strategies resulted in 
\textbf{79\% - 98\% improvement in performance}.
Compared to the baseline performance, the consolidation strategy contributes to 30\% - 94\% of the improvement and
the memoization strategy contributes to 67\% - 96\% on top of that. 
%TODO: factors of the network node-pair? path amount and length?

We conclude that \textbf{consolidation and memoization are both effective optimization strategies} in our verification
framework.

\subsection{Equivalence Class Reduction}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.5]{ec}
    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
    \label{fig:ec}
\end{figure}

While we cannot directly compare Tempus with other verifiers (due to difference in verification goal), 
we could use the amount of equivalence class as an indirect proxy of the verifier's behavior and performance.
Therefore, we measured the amount of additional equivalence classes introduced by Tempus in order to 
indirectly compare its overhead in addition to its functional counterparts.
This result is only influenced by the consolidation strategy, since the memoization strategy operates on 
a smaller granularity than equivalence classes.

Our results in Fig. \ref{fig:ec} shows the ratio between the amount of temporal equivalence classes that is 
being re-explored and the amount of functional equivalence class.
We note that we only need to \textbf{re-explore 4\% to 46\% equivalence classes} as an additional overhead 
compared to functional verification.

We conclude that \textbf{consolidation is effective in reducing the amount of equivalence classes that 
needs to be re-explored} in our verification framework.

\subsection{Optimization Analysis}

Up to this point, we have demonstrated that consolidation and memoization are two effective strategies in 
reducing the verification duration and amount of convolution procedure that needs to be done.
However, the results also show that their effectiveness varies depending on the topology.

To examine the factors that affects the effectiveness of each optimization strategy, we compute some properties 
of the topology that relates to the nature of the optimization.

% \begin{figure}[h]
%     \centering
%     \includegraphics[scale=0.5]{distribution}
%     \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
%     performance in seconds and red dots represents the amount of \textit{numerical} convolution}
%     \label{fig:dist}
% \end{figure}

\begin{itemize}
    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
        that there's still some variations
    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
\end{itemize}

\begin{itemize}
    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
        the amount of convolution, we also need to consider the length of each of those path
    \item By considering the both amount of path and its length to counting the convolution, we could 
        compare consolidation and memoization strategy to the unoptimized version
\end{itemize}

\begin{itemize}
    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
    \item Corresponds to the consolidation strategy
    \item \textbf{More than 50\% reduction}
    \item Doesn't work equally on all topologies, depend on 
    \begin{itemize}
        \item How many functional EC are explored (for a given inaccuracy level)
        \item How many possible paths a given node pair have in those ECs
    \end{itemize}
\end{itemize}

% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).

% On the performance side, we will evaluate the verifier by measuring the amount of additional 
% overhead that the temporal verifier would introduce, measured by the amount of states that 
% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
% produced by step 1, only 459 got re-explored on step 2.

### 2023-01-18 paper/6_related.tex
[Paper] Updated section 7, 8

\section{Related Works}
\textbf{Deterministic Data and Control Plane Verifier}
There has been a substantial body of work concerning the functional behavior of routers
with a given Forwarding Information Base (FIB) or Routing Information Base (RIB).

Prior works like HSA \cite{hsa}, NetKAT \cite{netkat}, and VeriFlow \cite{veriflow} focuses on verifying the 
current data plane forwarding behavior against some property.
While useful in some real-time cases, these tools are not built to explore the many
possible forwarding table given an enumeration of failure cases.

Control plane verifiers such as ARC \cite{arc}, Minesweeper \cite{minesweeper}, and 
Tiramisu \cite{tiramisu} focuses on verifying the various data plane states that 
could be produced by a control plane configuration over multiple failure scenarios.
Many researchers has built on top of this idea, such as DNA \cite{dna} where they 
focuses on verifying the change in property in the event of configuration update.

As comprehensive as it is, both approaches only considers the functional behavior 
of the network (e.g. reachability, loop detection) and they traditionally verify 
the stated properties in a black or white fashion, only determining whether a given 
property is violated or not.

As a result, unlike Tempus, its usefulness does not extend into the use cases of 
verifying SLA, where such binary guarantees are often rightfully avoided in favor 
of a probabilistic and/or quantitative agreement.

\textbf{Probabilistic Verifier}
More recent works have tried to address this very issue.
Works like ProbNetKAT \cite{probnetkat} (a probabilistic verifier based on NetKAT 
\cite{netkat}), NetDice \cite{steffen2020probabilistic}, and SRE \cite{sre} provides 
the user ability to define probabilistic failure model, where a given component 
of the network can fail with a given probability.

However, most of them still considers the functional property of the network as the 
main focus. 
While they are useful in verifying certain types of SLA (e.g. availability), they don't 
focus on more quantitative aspect of the network (e.g. bandwidth, latency).

\textbf{Quantitative Verifier}
Some recent works are an exception to this trend, however.
Works like QARC \cite{subramanian2020detecting} (a quantitative verifier based on 
ARC \cite{arc}) has developed a framework to verify one particular quantitative 
property: excess bandwidth on a certain link.
While QARC is a deterministic verifier, like Tempus, they focuses on a more quantitative 
property. 
Unlike Tempus however, this work is orthogonal to verifying latency and thus complementary 
to Tempus.

%TODO: ICNA paper?

\textbf{Network Calculus}
CCAC?

### 2023-01-18 paper/7_future.tex
[Paper] Updated section 7, 8

\section{Future Works}
\textbf{Addressing Distribution Independence}
The convolution of two probability distributions assumes that the two distributions are \textit{independent}
of each other.
While we accept this as a limitation of our model, latency in real networks (particularly the latency of two 
queues in the network) might be dependent of each other.

Queuing theory has laid some ground work to describe the asymptotic behavior of such queuing network.
However, they assume some information about the incoming traffic and can only describe a network with a 
particular traffic or delay distribution, like poisson.

Thus, eliminating the independence assumption without demanding an unwieldy amount of information from 
the user while maintaining (or even improving) the accuracy of the model remains an open challenge and an 
interesting future direction.

\textbf{Optimizing Functional Verification}
In our verifier, we built our temporal verifier on top of an existing functional verifier design.
While it makes the design of the verifier simpler, since there is clear separation of concern, there 
might be some performance benefit to gain if we integrate these two parts further.

While we have proposed and evaluated optimization techniques for the temporal verification step, 
integrating this optimization to the functional verification step -- perhaps by consolidating the 
equivalence classes before it is brought to the temporal verification stage in some way -- might make 
our approach more scalable.

While our evaluation shows that the functional verification stage is not the bottleneck of our design, 
discovering the symmetry in the exploration state that might accomplish this task might bring an interesting 
insight to the network verification community in general.

### 2023-01-18 paper/main.tex
[Paper] Updated section 7, 8

--- 2022-11-29 paper/main.tex
@@
 
-\graphicspath{ {./graph/} }
+\graphicspath{ {../viz/} }
 
@@
 
-\section{Introduction}
-Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
-routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a component failure.
-Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
-It is hard for a network engineer to reason about whether their configured network fulfilled some 
-intended property in various possible states of the network. 
-Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network control plane verifier.
+\input{1_intro.tex}
 
-Popular control plane verifiers, like ARC \cite{gember2016fast}, are formulated to answer 
-questions about a given property deterministically. 
-In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about the property (e.g. two nodes are reachable).
-While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate small fraction of failure. 
-For example, a given network might have an availability SLA of 99.999\%. 
-This kind of probabilistic properties have been the focus of a more recent works like NetDice 
-\cite{steffen2020probabilistic}.
+\input{2_overview.tex}
 
-The network property itself can also be divided into two kinds. 
-Quantitative properties like reachability and loop existence, are the common properties that are 
-studied by most of the existing verifiers. 
-More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
-properties like link bandwidth violation for a given traffic.
+\input{3_functional.tex}
 
-In this work, we're exploring the other side of the network performance metric: latency. 
-Certain network deployment often necessitates some latency requirement such as an ISP that has 
-latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
-We proposed a verification framework to probabilistically verify the latency property of packets 
-traversing from a source to destination node under various failure scenarios, by using latency 
-information of each components in the network.
+\input{4_temporal.tex}
 
-\begin{figure*}[h]
-    \centering
-    \includegraphics{../tikz/process}
-    \caption{Overview of Tempus}
-    \label{fig:process}
-\end{figure*}
+\input{5_eval.tex}
 
-\section{Overview}
-% \section{On Latency Modeling}
-\subsection{Latency as Path Property} 
-We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
-a single packet.
-In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
-through the many components of the network (e.g. links, routers, firewall) and each of those 
-components might introduce some latency into the packet transmission process.
-Figuring out which of those components will actually introduce latency into the packet in question, 
-and by how much, is the next logical step that we must figure out.
+\input{6_related.tex}
 
-Obviously, a given packet doesn't need to visit all network components to reach its destination 
-host.
-A network engineer will configure the network in such a way that a packet will only need to be 
-routed via a specific subset of its components.
-That specific subset is dictated by two things: how those components are connected together 
-(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
-protocol, ACL).
-Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
-given packet appropriately.
-
-The goal of a classical control plane verifier then, is to use this forwarding table in some 
-form to verify certain properties. %TODO: examples?
-However, looking at the forwarding table alone will not give us a conclusive result regarding 
-which components are going to be visited by a given packet, making verification of latency 
-properties less clear.
-For example, the network might be configured with a load-balancing protocol in which a packet 
-departing from a source host might take multiple different \textit{paths} (with certain 
-probability weights) to arrive at the destination host, possibly resulting in a different 
-end-to-end latency measurement.
-
-Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
-the primary unit of reasoning, rather than forwarding table.
-By being more specific about our unit of reasoning, we could answer verification questions more 
-clearly, and we could design our verifier more efficiently since we could use it to represent 
-multiple different forwarding tables that shares the same path.
-
-\subsection{Relation to Classical Verifier}
-Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that said path exists in the first place.
-We note that latency is a property that only make sense after connectivity between two hosts 
-has been established. 
-In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
-link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
-making the verification task trivial.
-
-Fortunately, there are a rich body of work in the network verification literature regarding 
-functional reachability under failure. %TODO: cite
-We could then design our verifier on top of an existing classical control plane verifier.
-We use it to verify reachability property, and only if the reachability property is fulfilled, 
-we would verify whether the latency between two hosts fulfilled some additional condition.
-
-\subsection{Path Latency Distribution}
-Up to this point, we have talked about measuring the latency of a single packet by figuring out 
-its path; analyzing which exact components it has traversed through.
-However, when we try to generalize this framework and ask questions about the latency of multiple 
-packets, it is apparent that the path alone is not a determinative information, as the latency 
-of two packets propagating through the same path might be different due to a multitude of factors.
-
-The natural extension to the framework then, is instead of representing latency of a path with a 
-single number, we instead represent it with a continuous random variable that signifies the 
-possible delay that a given packet traversing through that path might have.
-This random variable will have a distribution that marginalizes over all other factors other than 
-the path.
-
-We can then use this latency distribution to verify some temporal properties in a probabilistic 
-manner.
-For example, we could verify the probability that a packet will be delivered in under a time 
-unit by taking the CDF of the distribution.
-
-\subsection{Path Decomposition and Convolution} \label{decomposition}
-The final question that we had to decide on was how do we actually model path latency distribution
-based on real component measurements.
-Considering the complexity of factors that might determine the path latency distribution and the 
-availability of measurement data, we settle on the assumption that the path latency distribution 
-is composed of multiple independent distribution that corresponds to the latency each components 
-in the path might introduce.
-%TODO: confirm about independence?
-
-Since not all components in the path will introduce latency with a non-negligible value, we 
-specifically choose to model two source of latency in the path's components which we deem 
-significant: \textbf{link propagation latency} and \textbf{queuing latency}.
-%TODO: expand on how to get the data later
-
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
-
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
-
-In order to obtain the overall path latency distribution from these per-component latency 
-distributions, we do a convolution operations over all the relevant component's latency 
-distributions.
-Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
-convolution technique with a guaranteed error bound. 
-We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
-the lack of a general technique to measure error made us opt for the formerly mentioned method.
-
-% \section{Verifier Design}
-
-We divide the problem of latency verification into two parts: 
-verifying that two nodes are reachable (\textbf{functional verification}) and only if 
-the functional property is fulfilled, we would verify whether the latency between 
-two nodes fulfilled some temporal condition (\textbf{temporal verification}).
-
-\section{Functional Verification}
-For functional property verification, NetDice \cite{steffen2020probabilistic} 
-has laid the way for verifying reachability between two nodes under failure in a 
-probabilistic and efficient manner.
-To contextualize our modification, we will briefly describe some parts of 
-NetDice's network encoding and exploration algorithm as a prelude to our alteration.
-For a more complete complete description of NetDice, please refer to the original 
-paper.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics{../tikz/ex}
-    % \includegraphics{ex.eps}
-    \caption{Example topology, we want to check the reachability of A to Z}
-    \label{fig:ex}
-\end{figure}
-
-\subsection{Topology Graph}
-In this framework, a network is encoded in an edge-labeled directed graph 
-$G_t = (V_t, E_t)$ where $V_t$ represents the nodes in the network and 
-$E_t$ represents the functional connectivity between a source and a destination node. 
-A physical link is then represented as a pair of symmetrical edges that shares the same source 
-and destination node but with opposite direction (e.g. $v_1 \rightarrow v_2$ and $v_2 
-\rightarrow v_1$).
-A routing protocol is then defined on top of this graph by some auxilary information.
-
-To explain it with further clarity, we use the topology in Fig. \ref{fig:ex} as a 
-running example for this paper. 
-We will assume that the nodes in the network runs OSPF with weight 1 for every edge and ECMP 
-as their load-balancing scheme.
-We want to check the temporal probability of packets that departs from A to Z.
-
-% To represent the component's random failure, NetDice defined two failure rates.
-% The first failure rate is the chance that a given physical link in the network will go down.
-% The second failure rate is the chance that a given node in the network will go down.
-% These rates are shared between all links / nodes in the network.
-% Internally however, the node failure rate will be translated into the link failure rate 
-% with a Bayesian Network model, since a node failure can be modeled with the failure of each 
-% links connected to it.
-
-To represent random failure, NetDice defined a universal link failure rate (i.e. 
-the probability of \textit{any} link in the network randomly failing).
-We refine NetDice's model slightly by allowing each links to have different failure rate.
-We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
-\le 1\}$ that represents the failure rate of a given physical link.
-As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
-opposite direction) will also share the same failure rate, and will be disabled in a coupled 
-fashion.
-
-For the sake of example, in our Fig. \ref{fig:ex} topology, we will assume that each link has a 
-10\% chance of failure.
-
-% \subsubsection{Routing}
-% On top of this topology, NetDice also defined additional informations that would be used
-% by a routing protocol to determine the valid path(s) between two nodes $src, dst \in V_t$ given 
-% a particular link failure scenario. 
-% These routing informations would also be used by their optimization algorithms to reduce the 
-% amount of states that are going to be explored.
-
-% NetDice implemented iBGP and OSPF (with ECMP) routing protocols.
-% They encode the relevant routing information by assigning some labels to the vertices or edges
-% in the topology.
-% In OSPF for example, we define a function $w_{ospf}: E_t \rightarrow \mathbb{N}$ as the edge-label 
-% that represents the positive weight of a link.
-% They would later be used to compute the convergent paths of a given network state 
-% (\ref{verification}).
-
-\subsection{State Exploration}
-A \textbf{network state} is defined as a set of failed links.
-There are $2^{|E_t|}$ network states in a given network and a brute-force strategy 
-would need to explore all of the states to compute the reachability property of a 
-node pair in a given network. 
-NetDice however, will merge multiple network states into an \textbf{equivalence class}
-by marginalizing over \textit{cold edges}, links whose failure is guaranteed not to 
-change the convergent path of a highlighted network state.
-
-NetDice will systematically explore many equivalence classes in the network. 
-It will start from the equivalence class of a perfect network and will fail certain 
-\textit{hot edges} (i.e. edges that are not cold) to explore another equivalence 
-class.
-While not explicit in its description, NetDice's algorithm will effectively form an 
-\textbf{exploration tree}.
-
-In order to preserve some information that will be used in the temporal verification 
-stage, we modify this original exploration algorithm to make it more explicit.
-We started by formalizing the notion of Equivalence Classes.
-An Equivalence Class $\mathcal{E}$ is defined as a 3-tuple $(U_h, D_h, paths)$.
-$U_h$ and $D_h$ refers to a set of hot edges that are up and down respectively.
-$paths$ refers to the convergent paths between src-dst pair that is produced by the
-control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
-it could be used to for the temporal verification stage.
-
-We then define the Exploration Tree $\mathcal{T}$ as a tree of Equivalence Classes.
-At the root, we have an Equivalence Class that corresponds to a perfect network, where 
-$paths$ is the path(s) that the control plane will produce given a perfect network 
-condition and $U_h = D_h = \emptyset$.
-Each Equivalence Class in this tree will have children which have one of its parent's link in 
-$paths$ appended to its $D_h$, essentially failing one of the link in the path.
-If an Equivalence Class has an empty $paths$, then it would have no children.
-
-To compute the functional probability $P_f$ of a given Equivalence Class, we compute the 
-product of the 'up' probability of all the links in $U_h$ and $paths$ and the 
-product of the 'down' probability of all the links in $D_h$.
-
-In our running example, we could see the subset of the exploration tree in Fig. 
-\ref{fig:tree}. 
-We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
-shortest path between A to Z.  
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
-which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
-
-In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
-OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
-The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$
-($0.9^6 \cdot 0.1$).
-
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
-
-\section{Temporal Verification}
-In the temporal verification stage, we will use the path information that we got from 
-functional verification stage and augment them with latency information to 
-produce the final temporal property probability.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/tree}
-    % \includegraphics{ex.eps}
-    \caption{Subset of Modified Exploration Tree $\mathcal{T}$ for our example topology, 
-        $\mathcal{E}_2$ and $\mathcal{E}_3$ will get consolidated}
-    \label{fig:tree}
-\end{figure}
-
-\subsection{Latency Label}
-Similar to link failure $r$, we use edge-labeling technique to represent latency that 
-will get introduced by the connectivity between two nodes. 
-Our model will model two significant sources of latency: \textbf{link propagation} 
-and \textbf{packet queuing in the node}.
-We will encode these latencies by equipping the model with two additional labels.
-
-Encoding link propagation latency is fairly straightforward. 
-We define a function $l_p: E_t \rightarrow \mathcal{D}$ where $\mathcal{D}$ is a set of 
-continuous univariate distribution that has a minimum value of $0$.
-This distribution signifies the time it would take for the respective physical link to transmit 
-a packet from one end to another.
-Because of this, just like $r$, two symmetrical edges will share the same distribution.
-
-To encode queuing latency, we first note that the queuing mechanism in modern switches 
-usually resides in the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively 
-assign the latency to the connectivity between switches.
-To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
-This distribution signifies the output queue latency in the source node that is 
-going to be forwarded to the destination node.
-Unlike $l_p$, two symmetrical edges will have two different distributions since they 
-represent two different output queues.
-
-In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
-
-Given $paths$, $l_p$, and $l_q$, we could then compute $P_t$, the probability of a 
-given temporal property being fulfilled.
-We will describe our algorithm for computing this probability in the next few 
-sections.
-
-\subsection{Weighted Average and Path Convolution}
-Since a convergent behavior of the forwarding plane $paths$ might contain more than 
-one paths, we will compute the temporal probability of $paths$ by computing the 
-weighted average of the temporal probability of each of the individual path.
-The specific weight depends on the load-balancing method used.
-%We have implemented the weight computation for ECMP load-balancing scheme.
-
-To get the temporal probability of a single path, we will use the latency information 
-from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
-To do this, we resort to the methods of \textbf{convolution}.
-For each edge $e$ in the path, we will get $l_p(e)$ and $l_q(e)$ and convolve them 
-all together to get the path latency distribution.
-
-With this path latency distribution, we could get the temporal probability of a 
-path by computing the statistical property of said distribution.
-In this work, given a path latency distribution $\mathcal{L}$, we define two temporal 
-properties:
-\begin{itemize}
-    \item \textbf{Bounded Reachability}: the probability that a packet will get 
-        transmited below $t$ time unit. This will get computed as $cdf(\mathcal{L}, t)$
-    \item \textbf{Tail Reachability}: the probability that a packet will get 
-        transmited above $t$ time unit. This will get computed as 
-        $1 - cdf(\mathcal{L}, t)$
-\end{itemize}
-
-In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
-equally probable path, $ABEZ$ and $ADFZ$.
-We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
-$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
-We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
-desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.
-
-In short, given $paths$, $l_p$, and $l_q$, we will do the following:
-\begin{enumerate}
-    \item Split $paths$ into its individual path
-    \item For each path, compute its latency distribution by convolving the latency distribution 
-        $l_p$ and $l_q$ of each link
-    \item With the resulting path latency distribution, determine the probability of temporal 
-        property by computing the statistical property of said distribution
-    \item Combine the temporal property of each path by computing the weighted average, based 
-        on the load-balancing scheme of the control plane
-\end{enumerate}
-
-We do each of these steps to every Equivalence Class in $\mathcal{T}$, and do a sum-product operation 
-of $P_f$ and $P_t$ to get the final probability of the property in question.
-
-By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
-exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
-In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
-those $n$ Equivalence Class.
-We aim to minimize the overhead of temporal verification by introducing some optimization algorithms
-later.
-
-\subsection{Numeric Convolution}
-However, there is one major problem with a convolution-based technique for computing the latency distribution 
-of a given path: not every distribution pair can be convolved analytically.
-A closed-form solution of a convolution is usually only available for two distributions of the same type.
-
-In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of gamma distribution form.
-Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
-method.
-
-We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
-%TODO: describe DIRECT in one sentence
-We choose this method due to their bounded error property: DIRECT guarantee that the computed 
-distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
-DIRECT has also been implemented in R's popular bayesmeta package.
-%TODO: cite and change font?
-
-One subtle detail about DIRECT is while convolution is defined to be a commutative operation, and 
-DIRECT also achieves this property, the order of operation matters for the algorithm's runtime 
-performance.
-In particular, if we had a chain of DIRECT convolutions, the result of a convolution should not be set 
-as the second input distribution in the later convolutions to avoid performance penalty.
-This is caused by a nested loop from the following two mechanisms:
-\begin{itemize}
-    \item DIRECT works by computing a list of support values by doing many PDF queries of the second input
-        distribution
-    \item Computing the PDF of a DIRECT distribution (the returned distribution of a DIRECT 
-        algorithm) itself involves iterating over the current support values
-\end{itemize}
-
-As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
-to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
-$direct(C, direct(A, B))$.
-
-While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
-DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
-process.
-
-\section{Optimization}
-\subsection{Equivalence Class Consolidation}
-To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
-existing exploration algorithm so that we could reduce the exploration space even further.
-
-In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
-Class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same Equivalence Class will have the same convergent path(s).
-
-However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
-the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
-
-Since we define latency distribution to marginalize over all factors other than the path, we could 
-effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
-verification once.
-
-We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
-We will only do temporal probability computation once, when we first iterate over an equivalence class that 
-has a certain convergent path(s), and we cached the result should another equivalence class with the same 
-convergent path(s) emerged.
-
-\subsection{Path Memoization}
-After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
-bigger, but unique equivalence classes.
-As significant as it is, we could reduce the computation cost further by drawing connections between these 
-unique equivalence classes.
-
-In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
-through multiple possible paths.
-In the context of our framework, we say that an equivalence class might have more than one convergent paths.
-Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
-independent subset.
-In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
-
-Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
-individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
-on the temporal probability of a given path.
-If we were to compute the temporal probability of an equivalence class with previously cached individual 
-path, we only need to calculate the weights that corresponds to the load balancing protocol without 
-doing any convolution.
-
-In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
-cached result when we explore $\mathcal{E}_4$.
-
-With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
-
-\subsection{Convolution Grouping}
-Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
-optimization on the \textit{convolution operation itself}.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/group}
-    % \includegraphics{ex.eps}
-    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
-    we want to start introducing numerical convolution as late as possible}
-    \label{fig:grouping}
-\end{figure}
-
-While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
-in our framework, it comes with two noticeable disadvantages.
-
-The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
-comes with some error. 
-While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
-chain convolution.
-
-The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
-previous subsection, DIRECT is still generally slower than analytical convolution.
-This is expected, since a closed form solution of a convolution tipically boils the operation down into 
-a few arithmetic calculation, rather than complex operation like integration.
-Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
-path latency distribution.
-
-We did this by \textit{grouping} subsets of the path components with the same probability distribution 
-type (e.g. exponential, normal, etc.). 
-We then analytically convolve the distribution within that subset and only numerically convolve the 
-returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
-
-In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
-and then numerically convolve the resulting distribution with $l_q(EZ)$.
-Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
-
-% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{steffen2020probabilistic} have done. 
-% In addition, we also collect additional information from each state (e.g. convergent 
-% paths) to be used in the temporal verification step.
-
-% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-% state with $G_l$.
-% This is done by convolving over the latency distribution of each component in the path. 
-% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-% distribution with a KL divergence error bound.
-% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
-% explored}.
-
-% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
-% traverse from a source to destination below some time unit $T$), we can compute its probability by 
-% integrating the PDF from $0$ to $T$. 
-% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
-% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
-
-% TODO: intro, reformalize graph formulation, citations
-% Does a figure about topology graph and latency graph will be useful?
-% Multimedia context on older networked system Klara Nahrstedts
-% Graph
-
-\section{Evaluation}
-
-Spec
-\begin{itemize}
-    \item Julia
-    \item i7-1065G7 16GB RAM
-\end{itemize}
-
-Network model
-\begin{itemize}
-    \item OSPF weight 1
-    \item 0.1\% link failure rate
-    \item $10^{-8}$ inaccuracy
-    \item Reachability between two edge nodes
-    \item Queuing latency: LogNormal distribution
-    \item Link latency: either Gamma or Chisq (chosen at random)
-    \item Bounded reachability with t=1000.0
-\end{itemize}
-
-\subsection{Runtime Scalability}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{scalability}
-    \caption{Runtime performance of temporal verification, red dots represent the number of convolutions}
-    \label{fig:scalability}
-\end{figure}
-
-\begin{itemize}
-    \item The convolution step takes the most amount of time
-    \item \textbf{Runtime performance correlates with the amount of convolutions} $->$ plot
-    \item The amount of convolution is the result of the optimizations that we've done
-\end{itemize}
-
-\subsection{Equivalence Class Reduction}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{ec}
-    \caption{Amount of temporal EC compared to functional EC, label represents ratio}
-    \label{fig:ec}
-\end{figure}
-
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-\subsection{Optimization Effectiveness}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{optimization}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
-    \label{fig:opt}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-    \item We could see that in the best case, the optimization strategies \textbf{reduce the amount of 
-        convolution by 98\%} (in the case of GtsCe)
-\end{itemize}
-
-\subsection{Effect of Latency Distribution Type}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[scale=0.5]{distribution}
-    \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-    performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-    \label{fig:dist}
-\end{figure}
-
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
-
-\section{Related Works}
-
-\section{Future Works}
-\begin{itemize}
-    \item Addressing independence: convolution treats each probability distribution as independent 
-        with each other. Latency verifier that relaxes this assumption (or a better encoding) would be 
-        interesting future works
-    \item Optimizing functional verification: a functional verifiers that produces functional ECs that 
-        has already been consolidated would be interesting future works
-    \item Extracting Latency Information from Measurement Data: While Tempus serve as a framework for 
-        latency verification, future work needs to be done in translating measurement data into latency 
-        distributions. Julia supports distribution fitting given a set of data. 
-\end{itemize}
+\input{7_future.tex}
 
@@
         length.
-    \item The grouping optimization strategy is an effective optimization strategy, although the exact magnitude 
-        of its effectiveness will depend on the amount of similar distributions that can be convolved analytically
 \end{itemize}

### 2023-01-24 paper/5_eval.tex
[Paper] added fat tree evaluation

@@
 
-To appraise the performance and scalability of Tempus, we seek to answer 5 evaluation questions:
+To appraise the viability of our approach, we seek to answer 5 evaluation questions:
 \begin{enumerate}
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{scalability_fattree}
+    \caption{Fat Tree, red dots represent the number of convolutions}
+    \label{fig:scalabilityfat}
+\end{figure}
+
 \subsubsection{Performance and Scalability}
@@
 To drive this point more precisely, we also evaluate the running time of fat-tree topology in various sizes.
-Our results in Fig. X shows that by keeping the same type of topology and scaling them up, we increased 
-our verification time almost linearly?
+Our results in Fig. \ref{fig:scalabilityfat} shows that by keeping the same type of topology and scaling them up, we increased 
+our verification time almost linearly, while the unoptimized version timed out after 2 hours.
 

### 2023-01-24 paper/5_eval.tex
[Paper] Added 1 fat tree graph and some analysis

@@
 our verification time almost linearly, while the unoptimized version timed out after 2 hours.
+Not only that, our combined optimization strategies actually reduces the time for fat tree topology of $k=10$ compared to 
+equivalent topologies with smaller size.
 
@@
 
-% \begin{figure}[h]
-%     \centering
-%     \includegraphics[scale=0.5]{distribution}
-%     \caption{The effect of distribution type on runtime performance of AttMpls, label represents the 
-%     performance in seconds and red dots represents the amount of \textit{numerical} convolution}
-%     \label{fig:dist}
-% \end{figure}
+The effectiveness of the memoization strategy largely reflects the amount of unique paths that the functional 
+verifier has explored.
+The amount of unique paths (relative to the total explored paths) and the amount of reduction the memoization 
+strategy gains is highly correlated, as shown by Tab. X
 
-\begin{itemize}
-    \item While the amount of convolution gave us a pretty good proxy for performance, we could see 
-        that there's still some variations
-    \item This is due to the fact that we used two kinds of convolutions: analytical and numerical.
-    \item Since numerical convolution is slower, we could see that for the same topology and the same amount of 
-        overall convolution, \textbf{more numerical convolution will result in longer runtime performance}.
-\end{itemize}
+On the other hand, the effectiveness of the consolidation strategy is caused by more factors, such as the 
+topology and the particular load-balancing scheme of the network protocol. 
+As a proxy to this, we count the number of path that crosses each link, and averaged them.
+As we could see from Tab. X, Uninett2010 has a measure of about 108.87, reflecting the relative ishortcoming 
+of this particular strategy compared to another topology like GtsCe, which has a measure of 571.14.
 
-\begin{itemize}
-    \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-        the amount of convolution, we also need to consider the length of each of those path
-    \item By considering the both amount of path and its length to counting the convolution, we could 
-        compare consolidation and memoization strategy to the unoptimized version
-\end{itemize}
+% \begin{itemize}
+%     \item While the amount of path in those states can explain the amount of EC reduction, to get to 
+%         the amount of convolution, we also need to consider the length of each of those path
+%     \item By considering the both amount of path and its length to counting the convolution, we could 
+%         compare consolidation and memoization strategy to the unoptimized version
+% \end{itemize}
 
-\begin{itemize}
-    \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-    \item Corresponds to the consolidation strategy
-    \item \textbf{More than 50\% reduction}
-    \item Doesn't work equally on all topologies, depend on 
-    \begin{itemize}
-        \item How many functional EC are explored (for a given inaccuracy level)
-        \item How many possible paths a given node pair have in those ECs
-    \end{itemize}
-\end{itemize}
-
-% On the correctness side, we want to validate whether the result of our verifier is equivalent to 
-% another verifier that has lower fidelity (i.e. functional property only) if we set the temporal 
-% property to be virtually unconstrained (e.g. bounded reachability with a very high $T$).
-
-% On the performance side, we will evaluate the verifier by measuring the amount of additional 
-% overhead that the temporal verifier would introduce, measured by the amount of states that 
-% need to be re-explored (in the second step) and / or the amount of convolution operation performed. 
-% Early result on Highwinds \cite{knight2011internet} network shows that out of 2344 states 
-% produced by step 1, only 459 got re-explored on step 2.
+% \begin{itemize}
+%     \item Compare the amount of temporal EC to the functional (unconsolidated) EC
+%     \item Corresponds to the consolidation strategy
+%     \item \textbf{More than 50\% reduction}
+%     \item Doesn't work equally on all topologies, depend on 
+%     \begin{itemize}
+%         \item How many functional EC are explored (for a given inaccuracy level)
+%         \item How many possible paths a given node pair have in those ECs
+%     \end{itemize}
+% \end{itemize}

### 2023-01-25 paper/5_eval.tex
[Paper] Updated fat tree figures and analysis

@@
 \subsection{Experiment Setup}
+
+\input{graph/table_top}
+
 We tested Tempus on 5 real-world WAN topologies from the Topology Zoo and 3 datacenter topologies 
@@
 \end{figure}
+\begin{figure}[h]
+    \centering
+    \includegraphics[scale=0.5]{optimization_fattree}
+    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
+    \label{fig:opt_fat}
+\end{figure}
 
@@
 
+\input{graph/table_analysis}
+
 Up to this point, we have demonstrated that consolidation and memoization are two effective strategies in 

### 2023-02-01 paper/1_intro.tex
[Paper] Changed Intro

@@
 \section{Introduction}
-Modern networks consisted of various distributed protocols such as OSPF and BGP that exchange 
-routing information so that a packet can reach their intended destination efficiently, even in the 
-event of a component failure.
-Due to their configuration intricacies however, these protocols are notoriously hard to get right. 
-It is hard for a network engineer to reason about whether their configured network fulfilled some 
-intended property in various possible states of the network. 
-Thus, they are left between the choice of accepting the reality of Murphy's Law or getting a tool 
-to help them in this task, namely, network control plane verifier.
+The intricate configuraton of network devices is notoriously hard to get right, forcing network 
+engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
+correctness, namely, network verifier.
+Over the last decade, there has been a lot of development in the area of network verification 
+generally \cite{hsa}\cite{veriflow}.
+While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
+probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
+\textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
+scenarios -- such as reachability and loop detection.
 
-Popular control plane verifiers, like ARC \cite{arc}, are formulated to answer 
-questions about a given property deterministically. 
-In other words, given a set of states of the network (e.g. $k$-link failures) the verifier are 
-expected to give a yes or no answer about the property (e.g. two nodes are reachable).
-While useful to some extent, this kind of models are sometimes too restrictive since network 
-operators are usually able to tolerate small fraction of failure. 
-For example, a given network might have an availability SLA of 99.999\%. 
-This kind of probabilistic properties have been the focus of a more recent works like NetDice 
-\cite{steffen2020probabilistic}.
+However, network engineers also have the need to verify \textit{quantitative} properties, such as properties 
+related to bandwidth and latency.
+Still, barring some recent works about the verification of link load violation \cite{qarc}, work in this area 
+remain sparse, despite its undoubtable importance.
 
-The network property itself can also be divided into two kinds. 
-Quantitative properties like reachability and loop existence, are the common properties that are 
-studied by most of the existing verifiers. 
-More recent work, like QARC \cite{subramanian2020detecting}, has also explored qualitative 
-properties like link bandwidth violation for a given traffic.
-
-In this work, we're exploring the other side of the network performance metric: latency. 
+In this work, we're exploring the verification of the other side of the network performance metric: latency. 
 Certain network deployment often necessitates some latency requirement such as an ISP that has 
@@
 traversing from a source to destination node under various failure scenarios, by using latency 
-information of each components in the network.
+measurements of the components in the network.
+
+We introduce the design and implementation of \tool, a probabilistic network latency verification 
+framework.
+\tool will use the latency information from relevant component measurements (e.g. router queue length) 
+to infer the latency distribution of said component.
+Assuming that future traffic is i.i.d., we then employ a numerical convolution method to combine the
+latency distribution of each relevant components together to produce the end-to-end latency distribution 
+of an src-dst router pair.
+
+To determine said relevant components, we built \tool on top of an existing qualitative verification 
+framework to efficiently explore the path used for an src-dst router pair given various scenarios of 
+failures.
+By obtaining the end-to-end latency distribution from this information and the latency measurement, 
+we could analyze the statistical properties of the src-dst pair latency distribution to probabilistically 
+argue about the latency properties.
+
+We also introduced two optimization techniques on top of this framework to reduce the verification time 
+by multiple orders of magnitude.
+These two optimization techniques rely on the symmetry that we have found in the quantitative verifier 
+when brought into the context of latency verification.
+
+Our evaluation shows that the verification framework, with the help of our optimization techniques, 
+could accomplish the verification task on various WAN and datacenter topologies in the order of 
+minutes?
+
+With this work, we make the following technical contributions:
+\begin{itemize}
+    \item \textbf{Novel temporal verification framework} using numerical convolution of network 
+        components' latency measurement.
+    \item \textbf{Two optimization techniques} in said verification framework by exploiting 
+        symmetry in the failure exploration.
+    \item \textbf{\tool}, an implementation of our verification framework and its optimization 
+        on Julia.
+\end{itemize}
+
+The code for this work is open-sourced at link. %TODO: change link
+This work does not raise any ethical issues.
 

### 2023-02-01 paper/3_functional.tex
[Paper] Changed Intro

@@
 \section{Functional Verification}
-For functional property verification, NetDice \cite{steffen2020probabilistic} 
+For functional property verification, NetDice \cite{netdice} 
 has laid the way for verifying reachability between two nodes under failure in a 

### 2023-02-01 paper/6_related.tex
[Paper] Changed Intro

@@
 Works like ProbNetKAT \cite{probnetkat} (a probabilistic verifier based on NetKAT 
-\cite{netkat}), NetDice \cite{steffen2020probabilistic}, and SRE \cite{sre} provides 
+\cite{netkat}), NetDice \cite{netdice}, and SRE \cite{sre} provides 
 the user ability to define probabilistic failure model, where a given component 
@@
 Some recent works are an exception to this trend, however.
-Works like QARC \cite{subramanian2020detecting} (a quantitative verifier based on 
+Works like QARC \cite{qarc} (a quantitative verifier based on 
 ARC \cite{arc}) has developed a framework to verify one particular quantitative 

### 2023-02-01 paper/2_overview.tex
[Paper] Updated latency verification background

@@
-\section{Overview}
-% \section{On Latency Modeling}
-\subsection{Latency as Path Property} 
-We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
-a single packet.
-In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
-through the many components of the network (e.g. links, routers, firewall) and each of those 
-components might introduce some latency into the packet transmission process.
-Figuring out which of those components will actually introduce latency into the packet in question, 
-and by how much, is the next logical step that we must figure out.
+% \section{Overview}
+\section{Latency Modeling Background}
+In the context of verification, there are multiple approaches to model latency in a network.
+We will briefly compare 4 possible methods that might be useful to attack this problem and 
+and its limitation to illustrate the methodological challenge to accurately model latency.
 
-Obviously, a given packet doesn't need to visit all network components to reach its destination 
-host.
-A network engineer will configure the network in such a way that a packet will only need to be 
-routed via a specific subset of its components.
-That specific subset is dictated by two things: how those components are connected together 
-(i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
-protocol, ACL).
-Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
-given packet appropriately.
+\subsection{Simulation}
+One of the available methods to reason about latency in a network is by simulating it 
+using a network simulator (e.g. ns3, omnet).
+On a high level, the tool will simulate the lifecycle of packets, created by a load generator, 
+and the user could fail some network components and collect the packets' latency measurement 
+to see how the perturbation affects the latency property in question.
 
-The goal of a classical control plane verifier then, is to use this forwarding table in some 
-form to verify certain properties. %TODO: examples?
-However, looking at the forwarding table alone will not give us a conclusive result regarding 
-which components are going to be visited by a given packet, making verification of latency 
-properties less clear.
-For example, the network might be configured with a load-balancing protocol in which a packet 
-departing from a source host might take multiple different \textit{paths} (with certain 
-probability weights) to arrive at the destination host, possibly resulting in a different 
-end-to-end latency measurement.
+Network simulator has been widely used in our community as an alternative to experimenting on 
+a bare-metal system. 
+While generally accurate, network simulators notoriously take a long time to run.
+Moreover, the simulation workflow is more akin to testing than verification, relying on 
+the user to create failure scenarios and reason about latency in a contrapositive way -- 
+the absence of a negative result does not mean the property is guaranteed to be fulfilled.
 
-Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
-the primary unit of reasoning, rather than forwarding table.
-By being more specific about our unit of reasoning, we could answer verification questions more 
-clearly, and we could design our verifier more efficiently since we could use it to represent 
-multiple different forwarding tables that shares the same path.
+\subsection{Timed Automata}
+Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory that 
+adds a global clock to be used as constraints. 
+The runtime will keep a set of global clock that monotonously increases and can be individually 
+reset.
+We could construct the state machine that constraints each transition with a time limit. 
+By having this additional time constraints, we could verify properties in a similar way 
+as traditional model checking methods (e.g. using CTL formulas).
 
-\subsection{Relation to Classical Verifier}
-Before we analyze the latency of a given packet that propagates through a certain network 
-path however, we must make sure that said path exists in the first place.
-We note that latency is a property that only make sense after connectivity between two hosts 
-has been established. 
-In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
-link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
-making the verification task trivial.
+While this technique offers a proper verification framework and a higher level of abstraction 
+compared to simulation, traditional Timed Automata theory is deterministic, offering no random 
+transitions to express random component failure and a way to reason about latency probabilistically.
 
-Fortunately, there are a rich body of work in the network verification literature regarding 
-functional reachability under failure. %TODO: cite
-We could then design our verifier on top of an existing classical control plane verifier.
-We use it to verify reachability property, and only if the reachability property is fulfilled, 
-we would verify whether the latency between two hosts fulfilled some additional condition.
+A popular Timed Automata modeling tool, UPPAAL, tried to alleviate this determinism problem by 
+introducing Statistical Model Checking (SMC) on top of their Timed Automata model. 
+The feature allows user to introduce random transition (with configurable weights) and distribution 
+based transition, in order to probabilistically reason about the desired property and model 
+random failures of network components.
 
-\subsection{Path Latency Distribution}
-Up to this point, we have talked about measuring the latency of a single packet by figuring out 
-its path; analyzing which exact components it has traversed through.
-However, when we try to generalize this framework and ask questions about the latency of multiple 
-packets, it is apparent that the path alone is not a determinative information, as the latency 
-of two packets propagating through the same path might be different due to a multitude of factors.
+However, they did it by collecting the statistical property of multiple simulation runs, which 
+has the same drawback as the simulation-based method. 
+It is also limited in its choice of distribution, only allowing user to use uniform and 
+exponential distribution.
 
-The natural extension to the framework then, is instead of representing latency of a path with a 
-single number, we instead represent it with a continuous random variable that signifies the 
-possible delay that a given packet traversing through that path might have.
-This random variable will have a distribution that marginalizes over all other factors other than 
-the path.
+\subsection{Queuing Theory}
+Another theoretical framework that is more closely related to networking devices are queuing 
+theory.
+Queuing theory allows you to reason about the equilibrium latency of a given traffic, given the 
+arrival and processing distribution to each component, among other things.
 
-We can then use this latency distribution to verify some temporal properties in a probabilistic 
-manner.
-For example, we could verify the probability that a packet will be delivered in under a time 
-unit by taking the CDF of the distribution.
+One major downside of this framework, however, is the assumptions about what kind of probability 
+distributions being used. 
+There is a closed-form solution for a network where the queues are identified as a poisson process 
+(i.e. Jackson network). 
+Still, a queuing network with a more diverse queuing behavior does not have as strong theoretical 
+answer in this framework.
 
-\subsection{Path Decomposition and Convolution} \label{decomposition}
-The final question that we had to decide on was how do we actually model path latency distribution
-based on real component measurements.
-Considering the complexity of factors that might determine the path latency distribution and the 
-availability of measurement data, we settle on the assumption that the path latency distribution 
-is composed of multiple independent distribution that corresponds to the latency each components 
-in the path might introduce.
-%TODO: confirm about independence?
+\subsection{Convolution}
 
-Since not all components in the path will introduce latency with a non-negligible value, we 
-specifically choose to model two source of latency in the path's components which we deem 
-significant: \textbf{link propagation latency} and \textbf{queuing latency}.
-%TODO: expand on how to get the data later
+% \subsection{Latency as Path Property} 
+% We began our study by first pondering about a basic construct: modeling the end-to-end latency of 
+% a single packet.
+% In a packet-switched network, a packet is sent from one transmiting host to one receiving host 
+% through the many components of the network (e.g. links, routers, firewall) and each of those 
+% components might introduce some latency into the packet transmission process.
+% Figuring out which of those components will actually introduce latency into the packet in question, 
+% and by how much, is the next logical step that we must figure out.
 
-Propagation delay is the latency that is introduced by the links in the network, which is 
-independent of the traffic load in the system.
-Queuing delay is the latency that the queuing process in the node introduced. 
-Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
-packets there are in the queue, the more delay the node will introduce to a subsequent packets.
+% Obviously, a given packet doesn't need to visit all network components to reach its destination 
+% host.
+% A network engineer will configure the network in such a way that a packet will only need to be 
+% routed via a specific subset of its components.
+% That specific subset is dictated by two things: how those components are connected together 
+% (i.e. topology) and how the control plane protocols are configured to route a packet (e.g. routing 
+% protocol, ACL).
+% Based on a variety of these setups, nodes in the network will form a forwarding table to route a 
+% given packet appropriately.
 
-For propagation delay, the semantic of this random variable is relatively straightforward: it is 
-the distribution of latency that a given link will introduce.
-For queuing delay however, this random variable represents the delay that a given queuing process 
-will introduce, marginalized over various traffic pattern that a given network state might have 
-resulted.
+% The goal of a classical control plane verifier then, is to use this forwarding table in some 
+% form to verify certain properties. %TODO: examples?
+% However, looking at the forwarding table alone will not give us a conclusive result regarding 
+% which components are going to be visited by a given packet, making verification of latency 
+% properties less clear.
+% For example, the network might be configured with a load-balancing protocol in which a packet 
+% departing from a source host might take multiple different \textit{paths} (with certain 
+% probability weights) to arrive at the destination host, possibly resulting in a different 
+% end-to-end latency measurement.
 
-In order to obtain the overall path latency distribution from these per-component latency 
-distributions, we do a convolution operations over all the relevant component's latency 
-distributions.
-Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
-convolution technique with a guaranteed error bound. 
-We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
-the lack of a general technique to measure error made us opt for the formerly mentioned method.
+% Therefore, we argue that when it comes to analyzing end-to-end latency, a network path should be 
+% the primary unit of reasoning, rather than forwarding table.
+% By being more specific about our unit of reasoning, we could answer verification questions more 
+% clearly, and we could design our verifier more efficiently since we could use it to represent 
+% multiple different forwarding tables that shares the same path.
+
+% \subsection{Relation to Classical Verifier}
+% Before we analyze the latency of a given packet that propagates through a certain network 
+% path however, we must make sure that said path exists in the first place.
+% We note that latency is a property that only make sense after connectivity between two hosts 
+% has been established. 
+% In other words, if two nodes in a network aren't even functionally connected (e.g. physical 
+% link failure, ACL policy), then the latency between them will \textit{always} be infinite, 
+% making the verification task trivial.
+
+% Fortunately, there are a rich body of work in the network verification literature regarding 
+% functional reachability under failure. %TODO: cite
+% We could then design our verifier on top of an existing classical control plane verifier.
+% We use it to verify reachability property, and only if the reachability property is fulfilled, 
+% we would verify whether the latency between two hosts fulfilled some additional condition.
+
+% \subsection{Path Latency Distribution}
+% Up to this point, we have talked about measuring the latency of a single packet by figuring out 
+% its path; analyzing which exact components it has traversed through.
+% However, when we try to generalize this framework and ask questions about the latency of multiple 
+% packets, it is apparent that the path alone is not a determinative information, as the latency 
+% of two packets propagating through the same path might be different due to a multitude of factors.
+
+% The natural extension to the framework then, is instead of representing latency of a path with a 
+% single number, we instead represent it with a continuous random variable that signifies the 
+% possible delay that a given packet traversing through that path might have.
+% This random variable will have a distribution that marginalizes over all other factors other than 
+% the path.
+
+% We can then use this latency distribution to verify some temporal properties in a probabilistic 
+% manner.
+% For example, we could verify the probability that a packet will be delivered in under a time 
+% unit by taking the CDF of the distribution.
+
+% \subsection{Path Decomposition and Convolution} \label{decomposition}
+% The final question that we had to decide on was how do we actually model path latency distribution
+% based on real component measurements.
+% Considering the complexity of factors that might determine the path latency distribution and the 
+% availability of measurement data, we settle on the assumption that the path latency distribution 
+% is composed of multiple independent distribution that corresponds to the latency each components 
+% in the path might introduce.
+% %TODO: confirm about independence?
+
+% Since not all components in the path will introduce latency with a non-negligible value, we 
+% specifically choose to model two source of latency in the path's components which we deem 
+% significant: \textbf{link propagation latency} and \textbf{queuing latency}.
+% %TODO: expand on how to get the data later
+
+% Propagation delay is the latency that is introduced by the links in the network, which is 
+% independent of the traffic load in the system.
+% Queuing delay is the latency that the queuing process in the node introduced. 
+% Unlike propagation delay, queuing delay might be dependent on load in the system, since the more 
+% packets there are in the queue, the more delay the node will introduce to a subsequent packets.
+
+% For propagation delay, the semantic of this random variable is relatively straightforward: it is 
+% the distribution of latency that a given link will introduce.
+% For queuing delay however, this random variable represents the delay that a given queuing process 
+% will introduce, marginalized over various traffic pattern that a given network state might have 
+% resulted.
+
+% In order to obtain the overall path latency distribution from these per-component latency 
+% distributions, we do a convolution operations over all the relevant component's latency 
+% distributions.
+% Since not all distributions can be convolved in a closed-form fashion, we use a numerical 
+% convolution technique with a guaranteed error bound. 
+% We initially consider a Monte-Carlo simulation in order to approach numerical convolution, but 
+% the lack of a general technique to measure error made us opt for the formerly mentioned method.
 

### 2023-02-01 paper/main.tex
[Paper] Changed Intro
[Paper] Updated latency verification background

@@
 
-\title{Tempus: Probabilistic Network Latency Verifier}
+\newcommand{\tool}[0]{Tempus}
+
+\title{Probabilistic Network Latency Verification}
 \author{WIP}
@@
 \begin{abstract}
-    To combat problems and bugs that a given network might have, network verifiers have emerged as 
-    one of the promising solutions. 
+    To combat configuration errors that a given network might have, network verifiers have emerged as 
+    one of the promising solution. 
     State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under k-link failure. 
+    various scenarios of network failures, such as reachability under loop detection.
+
     However, as modern networks evolved and performance need becomes more stringent -- often 
@@
     nodes in the network.
-    Early evaluation shows that Tempus can verify bounded reachability property -- the probability
+
+    Our evaluation shows that \tool can verify bounded reachability property -- the probability
     that the average delay of packets traversing from a source to destination node is below $T$

### 2023-02-06 paper/2_overview.tex
[Paper] Completed section 2

@@
 Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory that 
-adds a global clock to be used as constraints. 
+adds a global clock to be used as transition constraints. 
 The runtime will keep a set of global clock that monotonously increases and can be individually 
 reset.
-We could construct the state machine that constraints each transition with a time limit. 
-By having this additional time constraints, we could verify properties in a similar way 
+Using these clocks, we could construct the state machine that constraints each transition with 
+a time limit. 
+With this additional time constraints, we could verify properties in a similar way 
 as traditional model checking methods (e.g. using CTL formulas).
@@
 has the same drawback as the simulation-based method. 
-It is also limited in its choice of distribution, only allowing user to use uniform and 
+It is also limited in its choice of probability distribution, only allowing user to use uniform and 
 exponential distribution.
@@
 
-One major downside of this framework, however, is the assumptions about what kind of probability 
-distributions being used. 
-There is a closed-form solution for a network where the queues are identified as a poisson process 
-(i.e. Jackson network). 
-Still, a queuing network with a more diverse queuing behavior does not have as strong theoretical 
-answer in this framework.
+One major downside of this framework, however, is the limitation of assumed probability 
+distributions that is being used in each queue and the arrival process. 
+Where one could derive a closed-form solution for a network where the queues and / or the arrival process 
+are identified as a poisson process (i.e. Jackson network, BCMP network \cite{bcmp}), 
+the same thing cannot be said for a queuing network that posits a more diverse queuing behavior.
+A framework to describe a queuing network with an arbitrary queuing behavior and arrival process in this 
+theory is yet to be discovered.
 
-\subsection{Convolution}
+\subsection{Latency as Random Variable}
+% Verification framework that is
+% - Expressive (congestion control, load-balancing scheme, traffic generation -> probabilistic)
+% - Performant (verification time that makes sense, proper-level of abstraction)
+% - Practical (reasonably accurate while requiring minimal amount of information and assumption)
+
+Considering these available approaches and their respective strength and limitation, we devise a 
+simple yet powerful framework to verify the latency property in a given network.
+
+Our main idea is to model the latency of a given network component as an independent random variable.
+By modeling the latency this way, we could then operate on those random variable to model the latency 
+of a given path, and eventually the latency of a given src-dst pair.
+Using the resulting random variable, we could then determine the temporal property in question by 
+analyzing the statistical property of the resulting probability distribution.
+
+\textbf{This framework is expressive}.
+The packet interarrival time that is represented by the latency distribution in each router encodes the 
+latency information of the various traffic that is going through the network.
+At the same time, it also encodes the effect of congestion and flow control scheme that affects 
+those traffic.
+
+\textbf{This framework is performant}.
+Operating on the high-level of abstraction of the latency distribution in each network components allows 
+this framework to perform the verification task in a matter of minutes?
+We demonstrate our evaluation result on Section \ref{eval}.
+
+\textbf{This framework is practical}.
+Queue length is a common metric that is used to measure the resource utilization of a network 
+\cite{dctcp} \cite{swift}.
+Using this metric, one could easily construct a latency distribution by combining it with the 
+line rate bandwidth.
 
@@
 
-We divide the problem of latency verification into two parts: 
+We divide the problem of latency verification in this framework into two parts: 
 verifying that two nodes are reachable (\textbf{functional verification}) and only if 
@@
 two nodes fulfilled some temporal condition (\textbf{temporal verification}).
+In Section \ref{sec:functional}, we briefly describe the functional verification scheme that 
+our framework is built upon. 
+Next, in Section \ref{sec:temp}, we describe our novel temporal verification framework 
+in more detail.
+Next, in Section \ref{sec:opt}, we describe the optimization techniques that arise from 
+these two steps.
+We then evaluate our implementation in Section \ref{eval}.
+Finally, we will touch on some related works on Section \ref{sec:rel}, limitations and future 
+works on \ref{sec:fut}, and conclude our work with Section \ref{sec:conc}.

### 2023-02-06 paper/3_functional.tex
[Paper] Completed section 2

@@
-\section{Functional Verification}
+\section{Functional Verification} \label{sec:functional}
 For functional property verification, NetDice \cite{netdice} 

### 2023-02-06 paper/4_temporal.tex
[Paper] Changed Intro
[Paper] Completed section 2

@@
-\section{Temporal Verification}
+\section{Temporal Verification} \label{sec:temp}
 In the temporal verification stage, we will use the path information that we got from 
@@
 
-\section{Optimization}
+\section{Optimization} \label{sec:opt}
 \subsection{Equivalence Class Consolidation}
@@
 % \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{steffen2020probabilistic} have done. 
+% \cite{netdice} have done. 
 % In addition, we also collect additional information from each state (e.g. convergent 

### 2023-02-06 test.sh

@@ -1,0 +1,20 @@
+julia --project=. src/Tempus.jl fattree4gen.json conv
+julia --project=. src/Tempus.jl fattree6gen.json conv
+julia --project=. src/Tempus.jl fattree8gen.json conv
+julia --project=. src/Tempus.jl fattree10gen.json conv
+julia --project=. src/Tempus.jl fattree12gen.json conv
+julia --project=. src/Tempus.jl fattree14gen.json conv
+
+julia --project=. src/Tempus.jl fattree4gen.json cons
+julia --project=. src/Tempus.jl fattree6gen.json cons
+julia --project=. src/Tempus.jl fattree8gen.json cons
+julia --project=. src/Tempus.jl fattree10gen.json cons
+julia --project=. src/Tempus.jl fattree12gen.json cons
+julia --project=. src/Tempus.jl fattree14gen.json cons
+
+julia --project=. src/Tempus.jl fattree4gen.json base
+julia --project=. src/Tempus.jl fattree6gen.json base
+julia --project=. src/Tempus.jl fattree8gen.json base
+julia --project=. src/Tempus.jl fattree10gen.json base
+julia --project=. src/Tempus.jl fattree12gen.json base
+julia --project=. src/Tempus.jl fattree14gen.json base

### 2023-02-06 paper/4_temporal.tex
[Paper] Polish

@@
 process.
-
-\section{Optimization} \label{sec:opt}
-\subsection{Equivalence Class Consolidation}
-To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
-existing exploration algorithm so that we could reduce the exploration space even further.
-
-In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
-Class based on its cold edges (edges whose failure won't change the convergent path(s)).
-In other words, network states within the same Equivalence Class will have the same convergent path(s).
-
-However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
-the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
-share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.
-
-Since we define latency distribution to marginalize over all factors other than the path, we could 
-effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
-verification once.
-
-We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
-We will only do temporal probability computation once, when we first iterate over an equivalence class that 
-has a certain convergent path(s), and we cached the result should another equivalence class with the same 
-convergent path(s) emerged.
-
-\subsection{Path Memoization}
-After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
-bigger, but unique equivalence classes.
-As significant as it is, we could reduce the computation cost further by drawing connections between these 
-unique equivalence classes.
-
-In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
-through multiple possible paths.
-In the context of our framework, we say that an equivalence class might have more than one convergent paths.
-Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
-independent subset.
-In other words, \textbf{multiple equivalence classes could share a subset of individual path}.
-
-Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
-individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
-on the temporal probability of a given path.
-If we were to compute the temporal probability of an equivalence class with previously cached individual 
-path, we only need to calculate the weights that corresponds to the load balancing protocol without 
-doing any convolution.
-
-In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
-cached result when we explore $\mathcal{E}_4$.
-
-With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
-
-\subsection{Convolution Grouping}
-Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
-optimization on the \textit{convolution operation itself}.
-
-\begin{figure}[h]
-    \centering
-    % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics[width=\columnwidth]{../tikz/group}
-    % \includegraphics{ex.eps}
-    \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
-    we want to start introducing numerical convolution as late as possible}
-    \label{fig:grouping}
-\end{figure}
-
-While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
-in our framework, it comes with two noticeable disadvantages.
-
-The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
-comes with some error. 
-While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
-chain convolution.
-
-The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
-previous subsection, DIRECT is still generally slower than analytical convolution.
-This is expected, since a closed form solution of a convolution tipically boils the operation down into 
-a few arithmetic calculation, rather than complex operation like integration.
-Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
-path latency distribution.
-
-We did this by \textit{grouping} subsets of the path components with the same probability distribution 
-type (e.g. exponential, normal, etc.). 
-We then analytically convolve the distribution within that subset and only numerically convolve the 
-returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.
-
-In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
-convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
-and then numerically convolve the resulting distribution with $l_q(EZ)$.
-Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.
-
-% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
-% \cite{netdice} have done. 
-% In addition, we also collect additional information from each state (e.g. convergent 
-% paths) to be used in the temporal verification step.
-
-% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
-% state with $G_l$.
-% This is done by convolving over the latency distribution of each component in the path. 
-% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
-% distribution with a KL divergence error bound.
-% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
-% explored}.
-
-% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
-% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
-% traverse from a source to destination below some time unit $T$), we can compute its probability by 
-% integrating the PDF from $0$ to $T$. 
-% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
-% get the final probability: the probability of \textit{the network} fulfilling a temporal property.
-
-% TODO: intro, reformalize graph formulation, citations
-% Does a figure about topology graph and latency graph will be useful?
-% Multimedia context on older networked system Klara Nahrstedts
-% Graph

### 2023-02-06 paper/5_opt.tex
[Paper] Polish

\section{Optimization} \label{sec:opt}
\subsection{Equivalence Class Consolidation}
To minimize the amount of temporal verification overhead, we want to find some symmetry in the 
existing exploration algorithm so that we could reduce the exploration space even further.

In the original NetDice exploration algorithm, multiple network states were merged into an Equivalence 
Class based on its cold edges (edges whose failure won't change the convergent path(s)).
In other words, network states within the same Equivalence Class will have the same convergent path(s).

However, we observe that each of those Equivalence Classes are not \textit{unique}: while network states within 
the same equivalence class shares a convergent path(s), \textbf{multiple equivalence classes could also 
share the exact same convergent path(s)}, as shown in $\mathcal{E}_2$ and $\mathcal{E}_3$ in Fig. \ref{fig:tree}.

Since we define latency distribution to marginalize over all factors other than the path, we could 
effectively \textit{consolidate} these equivalence classes into one big equivalence class and do temporal 
verification once.

We implemented this idea by doing memoization on the temporal probability of a given convergent path(s).
We will only do temporal probability computation once, when we first iterate over an equivalence class that 
has a certain convergent path(s), and we cached the result should another equivalence class with the same 
convergent path(s) emerged.

\subsection{Path Memoization}
After consolidating many equivalence classes that shares the same convergent path(s), we are left with fewer, 
bigger, but unique equivalence classes.
As significant as it is, we could reduce the computation cost further by drawing connections between these 
unique equivalence classes.

In a network with a load-balancing protocol, the convergent state of the data plane might forward a packet 
through multiple possible paths.
In the context of our framework, we say that an equivalence class might have more than one convergent paths.
Nevertheless, we note that two equivalence classes with different convergent paths might not be two 
independent subset.
In other words, \textbf{multiple equivalence classes could share a subset of individual path}.

Since the way we compute convergent path(s) temporal probability is by calculating the weighted average of 
individual path temporal probability, we could reduce the amount of convolution we will need to do by doing further memoization 
on the temporal probability of a given path.
If we were to compute the temporal probability of an equivalence class with previously cached individual 
path, we only need to calculate the weights that corresponds to the load balancing protocol without 
doing any convolution.

In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
an identical path, $ADFZ$.
We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
cached result when we explore $\mathcal{E}_4$.

With this optimization, we will do a chain convolution by the amount of path available in the network between 
the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
and $ADFZ$).

% \subsection{Convolution Grouping}
% Aside from optimizing the \textit{amount} of paths we need to do convolution on, we could also do some 
% optimization on the \textit{convolution operation itself}.

% \begin{figure}[h]
%     \centering
%     % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
%     \includegraphics[width=\columnwidth]{../tikz/group}
%     % \includegraphics{ex.eps}
%     \caption{One example of a convolution tree to get $\mathcal{L}_{ABEZ}$ with optimized grouping,
%     we want to start introducing numerical convolution as late as possible}
%     \label{fig:grouping}
% \end{figure}

% While the use of numerical technique to do convolution enabled us to use arbitrary probability distribution 
% in our framework, it comes with two noticeable disadvantages.

% The first is \textit{error}: numerical method is fundamentally an approximation technique that always 
% comes with some error. 
% While DIRECT algorithm comes with an error-bound guarantee, it is not zero and could expand if we do 
% chain convolution.

% The second is \textit{runtime performance}: even with the input ordering constraints we laid out on the 
% previous subsection, DIRECT is still generally slower than analytical convolution.
% This is expected, since a closed form solution of a convolution tipically boils the operation down into 
% a few arithmetic calculation, rather than complex operation like integration.
% Due to these factors, we feel the need to \textbf{prioritize analytical convolution} in the calculation of 
% path latency distribution.

% We did this by \textit{grouping} subsets of the path components with the same probability distribution 
% type (e.g. exponential, normal, etc.). 
% We then analytically convolve the distribution within that subset and only numerically convolve the 
% returned distribution, which will be fewer in quantity, thus enabling smaller error and faster performance.

% In our running example, in the chain convolution process to get $\mathcal{L}_{ABEZ}$, we will analytically
% convolve $l_p(AB)$, $l_q(AB)$, $l_p(BE)$, $l_q(BE)$, and $l_p(EZ)$ since they are of the same type (gamma) 
% and then numerically convolve the resulting distribution with $l_q(EZ)$.
% Fig. \ref{fig:grouping} illustrates how the chain convolution process will be broken down.

% \textbf{First}, we do the functional verification with $G_t$ similar to what NetDice 
% \cite{netdice} have done. 
% In addition, we also collect additional information from each state (e.g. convergent 
% paths) to be used in the temporal verification step.

% \textbf{Second}, we compute the total path latency distribution from the convergent path(s) in each 
% state with $G_l$.
% This is done by convolving over the latency distribution of each component in the path. 
% Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
% distribution with a KL divergence error bound.
% Multiple states can share the same convergent path(s), so \textit{only a fraction of the states need to be 
% explored}.

% \textbf{Third}, we calculate the probability of a given temporal property being true in that state. 
% For example, given a \textit{bounded reachability property} (probability of whether a packet can 
% traverse from a source to destination below some time unit $T$), we can compute its probability by 
% integrating the PDF from $0$ to $T$. 
% We then combine all of the \textit{temporal} and \textit{functional} probability in each state to 
% get the final probability: the probability of \textit{the network} fulfilling a temporal property.

% TODO: intro, reformalize graph formulation, citations
% Does a figure about topology graph and latency graph will be useful?
% Multimedia context on older networked system Klara Nahrstedts
% Graph

### 2023-02-08 paper/6_eval.tex
[Paper] Polish
[Paper] finished the evaluation result

--- 2023-01-25 paper/5_eval.tex
@@
-\section{Evaluation}
+\section{Evaluation} \label{eval}
 
@@
 
-To examine the factors that affects the effectiveness of each optimization strategy, we compute some properties 
-of the topology that relates to the nature of the optimization.
+\subsubsection{Why is Fat Tree 10 better than Fat Tree 4?}
+To analyze what affects the effectiveness of our optimization technique, we will start by explaining 
+one peculiar result in our evaluation: despite having a bigger size, the temporal verification runtime 
+of fat tree with $k = 10$ is faster than $k = 4$.
 
-The effectiveness of the memoization strategy largely reflects the amount of unique paths that the functional 
-verifier has explored.
-The amount of unique paths (relative to the total explored paths) and the amount of reduction the memoization 
-strategy gains is highly correlated, as shown by Tab. X
+This is due to the fact that the shortest path in $k = 10$ is a lot more diverse than that of $k = 4$.
+In a 3-tier Fat Tree topology, an edge node could reach another pod's edge node in at least 4 hops.
+The amount of this shortest path is $(k/2)^2$, which is the same as the number of core node.
 
-On the other hand, the effectiveness of the consolidation strategy is caused by more factors, such as the 
-topology and the particular load-balancing scheme of the network protocol. 
-As a proxy to this, we count the number of path that crosses each link, and averaged them.
-As we could see from Tab. X, Uninett2010 has a measure of about 108.87, reflecting the relative ishortcoming 
-of this particular strategy compared to another topology like GtsCe, which has a measure of 571.14.
+We could see from Table \ref{tab:analysis} that when $k \geq 10$, the amount of unique path explored by the 
+functional verifier is exactly the same as the number of core node. 
+Combined with the fact that the average amount of convolution per path is 7 (which suggests that the path 
+length is 4), means that the functional verifier only produces equivalence classes that consisted of only 
+different combination of these shortest paths.
 
-% \begin{itemize}
-%     \item While the amount of path in those states can explain the amount of EC reduction, to get to 
-%         the amount of convolution, we also need to consider the length of each of those path
-%     \item By considering the both amount of path and its length to counting the convolution, we could 
-%         compare consolidation and memoization strategy to the unoptimized version
-% \end{itemize}
+When $k < 10$ however, the functional verifier also needs to produce equivalence classes that results 
+in a longer convergent paths in order to reach the same accuracy level.
+in Fat Tree, this is usually marked by a visit to another aggregation node in another pod. 
+We could see from Table \ref{tab:analysis} that when $k < 10$, the amount of of convolution per path is more 
+than 7 (which suggests that a path longer than 4 hops exist).
 
-% \begin{itemize}
-%     \item Compare the amount of temporal EC to the functional (unconsolidated) EC
-%     \item Corresponds to the consolidation strategy
-%     \item \textbf{More than 50\% reduction}
-%     \item Doesn't work equally on all topologies, depend on 
-%     \begin{itemize}
-%         \item How many functional EC are explored (for a given inaccuracy level)
-%         \item How many possible paths a given node pair have in those ECs
-%     \end{itemize}
-% \end{itemize}
+\subsubsection{Generalization}
+From the insight of this particular result, we could draw two properties in a general network that could 
+determine the runtime of our verifier.
+
+The first one is the \textbf{amount of possible paths}.
+If the amount of possible paths between a src-dst pair in a given topology is small, then the functional 
+verification step would produce fewer equivalence classes and end early.
+While most of our evaluated topology had a lot of possible paths, one exception to this is LatNet, which 
+only have 3 possible paths.
+
+The second is whether the src-dst pair in a given topology would \textbf{produce paths that are "symmetric"}.
+By symmetric, we mean that the routing and load balancing method produce a lot of identical equivalence class 
+and / or equivalence class that share the same path.
+As our explanation about Fat Tree 10 suggests, larger Fat Tree has a lot of path with the same length.
+This would result in a lot of symmetrical equivalence classes under ECMP.
+Our optimization techniques could identify this symmetry and reduce the amount of convolution in the temporal 
+verification step.
+
+%TODO: path enumeration?

### 2023-02-08 paper/main.tex
[Paper] Completed section 2
[Paper] Polish
[Paper] finished the evaluation result

@@
     State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability under loop detection.
+    various scenarios of network failures, such as reachability and loop detection.
 
@@
 
-    In this work, we introduce a verifier framework, Tempus, that can answer the probability of 
+    In this work, we introduce a verification framework, \tool, that can answer the probability of 
     a given temporal property being true given latency distributions of individual links and 
     nodes in the network. 
+    We use the convolution operation on these probability distributions to calculate the end-to-end 
+    path latency distribution, and answer property by analyzing the statistical property of this 
+    resulting distribution.
+    We also introduce two effective optimization techniques to exploit the symmetry in the verification 
+    process to reduce the amount of convolution operation needed.
 
-    Our evaluation shows that \tool can verify bounded reachability property -- the probability
-    that the average delay of packets traversing from a source to destination node is below $T$
-    time unit -- by only adding a fraction of the state exploration overhead introduced by the 
-    qualitative verifier it's built upon.
+    Our evaluation shows that \tool can verify temporal property 4 orders of magnitude faster on average than 
+    our baseline, which is devoid of all optimization.
+    We also identify that highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
+    and can cut the amount of convolution up to 6 orders of magnitude.
 \end{abstract}
@@
 
-\input{5_eval.tex}
+\input{5_opt.tex}
 
-\input{6_related.tex}
+\input{6_eval.tex}
 
-\input{7_future.tex}
+\input{8_future.tex}
 
-\section{Conclusion}
+\input{7_related.tex}
+
+\section{Conclusion} \label{sec:conc}
 \begin{itemize}

### 2023-02-13 paper/1_intro.tex
[Paper] Changed abstract, section1-3

@@
 \section{Introduction}
-The intricate configuraton of network devices is notoriously hard to get right, forcing network 
-engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
-correctness, namely, network verifier.
+% The intricate configuraton of network devices is notoriously hard to get right, forcing network 
+% engineers to accept the reality of Murphy's Law or getting some help from a tool to ensure their 
+% correctness, namely, network verifier.
+% Over the last decade, there has been a lot of development in the area of network verification 
+% generally \cite{hsa}\cite{veriflow}.
+% While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
+% probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
+% \textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
+% scenarios -- such as reachability and loop detection.
+
+Many modern network deployments often necessitate tight performance guarantee in order to sustain 
+their increasing demand.
+At the same time, end users also have a rising performance standard with regards to their internet browsing 
+experience, forcing companies to optimize every stack of their deployment in order to increase conversion.
+Hence, network operator must find a way to verify that the network that they configure meets this demand, or 
+accept the reality of Murphy's Law.
+
 Over the last decade, there has been a lot of development in the area of network verification 
 generally \cite{hsa}\cite{veriflow}.
-While their approach varies a lot (e.g. data plane vs. control plane verification, deterministic vs. 
-probabilistic), most of these verifiers shares a common focus, primarily fixated on verifying 
-\textit{qualitative} properties -- the convergent functional behavior of a network under various failure 
-scenarios -- such as reachability and loop detection.
+Still, most of these verifiers are primarily fixated on verifying \textit{qualitative} properties 
+-- the convergent functional behavior of a network under various failure scenarios -- 
+such as reachability and loop detection, and not \textit{quantitative} properties 
+related to performance, such as bandwidth and latency.
+Compared to the verification of reachability-based property, verification of bandwidth and latency is 
+inherently more challenging since it introduces more dimension to the problem.
+In addition to reasoning about the existence of a flow, one needs to reason about the measured 
+detail of said flow and model their interaction.
 
-However, network engineers also have the need to verify \textit{quantitative} properties, such as properties 
-related to bandwidth and latency.
-Still, barring some recent works about the verification of link load violation \cite{qarc}, work in this area 
-remain sparse, despite its undoubtable importance.
+There has been a few notable works that push network verification concept towards this direction \cite{qarc}
+\cite{ccac}.
+These newer tools model a wide range of additional network behavior that might affect performance to answer a
+specific question in mind, namely link-load violation and congestion control behavior.
+However, their approaches so far has been focused on worst-case analysis, modeling and answering whether 
+an unwanted event could happen or not, which requires further statistical interpretation in order to translate 
+them to a percentage guarantee that appear on an SLA.
 
-In this work, we're exploring the verification of the other side of the network performance metric: latency. 
-Certain network deployment often necessitates some latency requirement such as an ISP that has 
-latency SLA \cite{Verizon} or deployments of Time-Sensitive Networking (TSN) \cite{TSN}.
-We proposed a verification framework to probabilistically verify the latency property of packets 
+Moreover, none of these tools were designed to answer questions related to end-to-end latency SLA \cite{Verizon}.
+Ideally, such a tool would answer the challenge of modeling various latency altering network behavior, 
+such as traffic generation, load balancing scheme, and congestion control, and expresses the important 
+statistical property of the resulting end-to-end latency.
+Not just in terms of worst-case behaviors and tails, but also the means.
+
+In this work, we propose a verification framework to probabilistically verify the latency property of packets 
 traversing from a source to destination node under various failure scenarios, by using latency 

### 2023-02-13 paper/3_functional.tex
[Paper] Changed abstract, section1-3

--- 2023-02-01 paper/3_functional.tex
@@
-\section{Functional Verification}
-For functional property verification, NetDice \cite{netdice} 
+\section{Functional Verification} \label{sec:functional}
+For functional verification stage, NetDice \cite{netdice} 
 has laid the way for verifying reachability between two nodes under failure in a 
@@
 control plane when the links in $D_h$ fails.
-Unlike the original algorithm, we explicitly store $paths$ in $\mathcal{E}$ so that 
-it could be used to for the temporal verification stage.
 
@@
 
-After having the algorithm produced $\mathcal{T}$, we then continue to the temporal verification 
-stage.
+After each equivalence class produced its convergent paths, we then continue to verify the 
+temporal property of that equivalence class in the next stage.

### 2023-02-13 paper/4_temporal.tex
[Paper] Changed abstract, section1-3

@@
 Our model will model two significant sources of latency: \textbf{link propagation} 
-and \textbf{packet queuing in the node}.
+and \textbf{packet queuing} in the node.
 We will encode these latencies by equipping the model with two additional labels.
@@
 usually resides in the output port. %TODO: Cite PISA?
-Since a port in a switch is only connected to one other port, we could effectively 
-assign the latency to the connectivity between switches.
+Since one port in a switch is only connected to one port in another switch, we could effectively 
+assign queuing latency to the connectivity between switches.
 To do this, we define another function $l_q: E_t \rightarrow \mathcal{D}$.
@@
 
-In our running example, we will assume that all edges in the topology will have $l_p$
-and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
-have a lognormal distribution with $\alpha = 1$ and $\theta = 1$.
+% In our running example, we will assume that all edges in the topology will have $l_p$
+% and $l_q$ of gamma distribution with $\lambda = 1$, except $l_q(EZ)$ which will 
+% have a lognormal distribution with $\alpha = 1$ and $\theta = 1$
+
+\subsection{Latency Measurement as Distribution}
+% To form the required latency distribution
+
+% iid
+
+% independent
 
@@
 The specific weight depends on the load-balancing method used.
-%We have implemented the weight computation for ECMP load-balancing scheme.
+In our example, we will use ECMP load-balancing scheme.
 
 To get the temporal probability of a single path, we will use the latency information 
-from $l_p$ and $l_q$ to compute the latency distribution of the whole path.
+from $l_p$ and $l_q$ of each edge in the path to compute the latency distribution of the whole path.
 To do this, we resort to the methods of \textbf{convolution}.
@@
 
-By introducing the temporal verification algorithm, we're essentially adding additional overhead to the 
-exploration algorithm (on top of functional verification) that scales to the size of the Exploration Tree.
+At this point, by introducing the temporal verification algorithm, we're essentially adding additional 
+overhead to the exploration algorithm (on top of functional verification) that scales to the size of the 
+Exploration Tree.
 In other words, for $\mathcal{T}$ of size $n$, we will need to do temporal verification in each of 
@@
 \subsection{Numeric Convolution}
-However, there is one major problem with a convolution-based technique for computing the latency distribution 
+There is one major problem with a convolution-based technique for computing the latency distribution 
 of a given path: not every distribution pair can be convolved analytically.

### 2023-02-13 paper/main.tex
[Paper] Changed abstract, section1-3

@@
 \begin{abstract}
-    To combat configuration errors that a given network might have, network verifiers have emerged as 
-    one of the promising solution. 
-    State-of-the-art network verifiers mainly focused on evaluating qualitative properties under 
-    various scenarios of network failures, such as reachability and loop detection.
-
-    However, as modern networks evolved and performance need becomes more stringent -- often 
-    expressed in terms of Service Level Agreement (SLA) -- there is a need to evolve network 
-    verifiers to also reason about quantitative performance properties. 
-    Works in quantitative network verifiers that has arose in recent years mainly focused on one 
-    side of the network performance metric: bandwidth and load violation properties. 
-    Questions about the other side of network performance metric, latency, were left unanswered.
-
-    In this work, we introduce a verification framework, \tool, that can answer the probability of 
-    a given temporal property being true given latency distributions of individual links and 
-    nodes in the network. 
-    We use the convolution operation on these probability distributions to calculate the end-to-end 
-    path latency distribution, and answer property by analyzing the statistical property of this 
-    resulting distribution.
+    As modern networks evolved and performance need became more stringent, often 
+    expressed in terms of Service Level Agreement (SLA), there is a rising need to verify that these 
+    agreements hold true.
+    Designing a network verifier that can practically reason about quantitative properties proves to be 
+    a unique challenge, since it needs to be able to expressively model many quantitative phenomenon, like 
+    congestion control and load balancing scheme, while being reasonably performant and not 
+    requiring too many inputs from its user. 
+    In this work, we introduce a network latency verification framework, \tool, that could probablistically 
+    verify temporal properties of a network given latency measurements of its components. 
     We also introduce two effective optimization techniques to exploit the symmetry in the verification 
-    process to reduce the amount of convolution operation needed.
-
-    Our evaluation shows that \tool can verify temporal property 4 orders of magnitude faster on average than 
+    process to significantly reduce the verification runtime.
+    Our evaluation shows that \tool\ can verify temporal property 90\% faster on average than 
     our baseline, which is devoid of all optimization.

### 2023-02-14 paper/4_temporal.tex
[Paper] updated various things

@@
 \subsection{Latency Measurement as Distribution}
-% To form the required latency distribution
+To have the required latency distribution for $l_p$ and $l_q$, we form an empirical distribution 
+from latency measurement information.
+This measurement information might come in the form of packet inter-arrival time, 
+queue length, and other data that is similar in nature.
 
-% iid
-
-% independent
+Since packet inter-arrival time represent the effect of various traffic shaping mechanism on 
+the many traffics on the network, we could obtain such data from published works like 
+\cite{dctcp}\cite{swift}.
+Central to our model is the assumption that the latency introduced by the covered component is 
+i.i.d. and are independent of one another.
 
@@
 
-In our example, we define $l_q(EZ)$ to be lognormal distribution, which doesn't have a closed form convolution 
-solution with the other latency distribution in the network, which is of gamma distribution form.
+In our model however, we intend to use real-world empirical data to be used as latency distribution, which 
+might be arbitrarily shaped.
 Therefore, in order to convolve two arbitrary distributions, we need to leverage a numerical convolution 
@@
 
-We leverage an existing numerical convolution algorithm, DIRECT, to fulfill this role.
-%TODO: describe DIRECT in one sentence
+We leverage an existing numerical convolution algorithm, DIRECT \cite{direct}, to fulfill this role.
 We choose this method due to their bounded error property: DIRECT guarantee that the computed 
 distribution and the correct theoretical distribution has a KL-divergence below a certain bound.
-DIRECT has also been implemented in R's popular bayesmeta package.
+DIRECT has also been implemented in R's popular bayesmeta package \cite{bayesmeta}.
 %TODO: cite and change font?
+
+DIRECT works by numericaly approximating \textit{mixture distributions} from a large mixture components.
+It is done by ``binning'' the mixture component / mixing distribution and intelligently chose a single 
+reference value for each bin to compute the final CDF of the resulting mixing distribution.
+The amount of bins and the reference value is determined based on the error bound that the user has set.
 
@@
 As an ilustration, if we had 3 random variables $A$, $B$, and  $C$, and we want 
-to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of 
-$direct(C, direct(A, B))$.
-
-While this problem is trivial to solve (swap the input parameter if the latter is detected as a 
-DIRECT distribution), it does limit the type of optimization that we could have in the chain convolution 
-process.
+to convolve them all together, it would be faster to compute $direct(direct(A, B), C)$ instead of \\
+$direct(C, direct(A, B))$. Trivially but importantly, we swap the input parameter if the latter is detected as a 
+DIRECT distribution.

### 2023-02-14 paper/7_related.tex
[Paper] updated various things

--- 2023-02-01 paper/6_related.tex
@@
-\section{Related Works}
+\section{Related Works} \label{sec:rel}
 \textbf{Deterministic Data and Control Plane Verifier}
@@
 \textbf{Network Calculus}
-CCAC?
+One promising concept with regards to performance metric verification is the theoretical 
+framework of network calculus.
+Network calculus aims to provide a framework to analyze the bandwidth and latency guarantee 
+of a flow.
+It could model link capacities, background traffic, and congestion control.
+CCAC \cite{ccac}, for example, has developed a tool based on this framework to help explain 
+the various behavior of a congestion control system.
+
+While providing a strong theoretical background in latency verification, this framework 
+primarily deals with guaranteeing bounds deterministically.
+It also requires the user to provide more information compared to \tool, since the user 
+need to provide the cummulative arrival and departure curve over time.

### 2023-02-14 paper/main.tex
[Paper] updated various things
[Paper] updated graph for fattree

@@
-\documentclass[10pt,sigconf,letterpaper,anonymous,nonacm]{acmart}
+\documentclass[10pt,sigconf,letterpaper]{acmart}
 
@@
 \title{Probabilistic Network Latency Verification}
-\author{WIP}
+\author{\#929}
 
@@
 \section{Conclusion} \label{sec:conc}
-\begin{itemize}
-    \item Tempus is a scalable latency SLA verifier. It could probabilistically verify in the order of minutes
-    \item Consolidation and memoization is an effective optimization strategy, although the exact magnitude of 
-        its effectiveness will depend on the network topology, specifically on the amount of paths and their 
-        length.
-\end{itemize}
+\tool is a scalable latency SLA verifier. 
+It could probabilistically verify temporal properties orders of magnitude faster than its unoptimized 
+baseline.
+Equivalence Class Consolidation and Path Memoization are two effective optimization strategies, 
+although the exact magnitude of its effectiveness will depend on the network topology. 
+At its best, if the network topology allows only a handful of path between src-dst pair or the paths are 
+highly symmetrical, then it could cut temporal verification time up to 6 orders of magnitude.
 

### 2023-02-22 paper/1_intro.tex
[Paper] Added new overview diagram
[Paper] Updated figures and polish

@@
 
-Over the last decade, there has been a lot of development in the area of network verification 
-generally \cite{hsa}\cite{veriflow}.
-Still, most of these verifiers are primarily fixated on verifying \textit{qualitative} properties 
--- the convergent functional behavior of a network under various failure scenarios -- 
-such as reachability and loop detection, and not \textit{quantitative} properties 
-related to performance, such as bandwidth and latency.
-Compared to the verification of reachability-based property, verification of bandwidth and latency is 
-inherently more challenging since it introduces more dimension to the problem.
+Verification of network properties related to performance metric, like bandwidth and latency, is inherently 
+more challenging than a classic reachabality-based property since it introduces more dimension to the problem.
 In addition to reasoning about the existence of a flow, one needs to reason about the measured 
@@
 
-There has been a few notable works that push network verification concept towards this direction \cite{qarc}
-\cite{ccac}.
-These newer tools model a wide range of additional network behavior that might affect performance to answer a
-specific question in mind, namely link-load violation and congestion control behavior.
+Recently, there has been a few notable works that push network verification concept towards this direction 
+\cite{qarc}\cite{pita}\cite{ccac}.
+These \textit{quantitative} verifier model a wide range of additional network behavior that might affect 
+performance to answer a specific question in mind, such as link-load violation and congestion control behavior.
 However, their approaches so far has been focused on worst-case analysis, modeling and answering whether 
@@
 
-Moreover, none of these tools were designed to answer questions related to end-to-end latency SLA \cite{Verizon}.
+Moreover, none of these tools were designed to answer questions related to end-to-end latency, 
+which is one of many important property that needs to be made certain by network providers \cite{Verizon}.
 Ideally, such a tool would answer the challenge of modeling various latency altering network behavior, 
@@
 statistical property of the resulting end-to-end latency.
-Not just in terms of worst-case behaviors and tails, but also the means.
+This is done not just in terms of worst-case behaviors and tails, but also the means.
 
@@
 Our evaluation shows that the verification framework, with the help of our optimization techniques, 
-could accomplish the verification task on various WAN and datacenter topologies in the order of 
-minutes?
+could accomplish the verification task on various WAN and datacenter topologies around 90\% faster than 
+our unoptimized baseline.
+Moreover, highly symmetrical topology, like Fat Tree, benefits from our optimization the most 
+and could reduce the verification time up to 6 orders of magnitude.
 
@@
 
-The code for this work is open-sourced at link. %TODO: change link
+The code for this work is open-sourced on GitHub. %TODO: change link
 This work does not raise any ethical issues.
@@
     \centering
-    \includegraphics{../tikz/process}
+    \includegraphics[scale=0.5]{overall}
     \caption{Overview of Tempus}

### 2023-02-22 paper/2_overview.tex
[Paper] Polish
[Paper] Updated figures and polish

@@
 In the context of verification, there are multiple approaches to model latency in a network.
-We will briefly compare 4 possible methods that might be useful to attack this problem and 
+We will briefly compare 3 possible methods that might be useful to attack this problem and 
 and its limitation to illustrate the methodological challenge to accurately model latency.
@@
 One of the available methods to reason about latency in a network is by simulating it 
-using a network simulator (e.g. ns3, omnet).
+using a network simulator (e.g. ns-3 \cite{ns3}, OMNeT \cite{omnet}).
 On a high level, the tool will simulate the lifecycle of packets, created by a load generator, 
@@
 \subsection{Timed Automata}
-Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory that 
+The other alternative is by using Timed Automata.
+Timed Automata is an extension to the classic Finite State Machine (FSM) in automata theory where it 
 adds a global clock to be used as transition constraints. 
@@
 
-A popular Timed Automata modeling tool, UPPAAL, tried to alleviate this determinism problem by 
+A popular Timed Automata modeling tool, UPPAAL \cite{uppaal}, tried to alleviate this determinism problem by 
 introducing Statistical Model Checking (SMC) on top of their Timed Automata model. 
@@
 
-\subsection{Latency as Random Variable}
+\subsection{\tool}
 % Verification framework that is

### 2023-02-22 paper/3_functional.tex
[Paper] Updated figures and polish

@@
     % \includegraphics[width=0.4\textwidth, trim=0cm 11cm 0cm 5cm]{ex.pdf}
-    \includegraphics{../tikz/ex}
+    \includegraphics[width=\columnwidth]{ex_topology}
     % \includegraphics{ex.eps}
-    \caption{Example topology, we want to check the reachability of A to Z}
+    \caption{Example topology, we want to check the reachability of S to T}
     \label{fig:ex}
@@
 We refine NetDice's model slightly by allowing each links to have different failure rate.
-We label each edge in $E_t$ with a function $r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x 
-\le 1\}$ that represents the failure rate of a given physical link.
+We label each edge in $E_t$ with a function 
+$$r: E_t \rightarrow \{x \in \mathbb{R} \mid 0 \le x \le 1\}$$ 
+that represents the failure rate of a given physical link.
 As a consequence, two symmetrical edges (i.e. two edges that shares the same node pairs but with 
@@
 \ref{fig:tree}. 
-We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $ACZ$ as the 
-shortest path between A to Z.  
+We see that in a perfect network ($\mathcal{E}_1$), OSPF will produce $SAT$ as the 
+shortest path between $S$ to $T$.  
 The probability of this Equivalence Class actually materializing ($P_f$) is $0.81$, 
-which is the probability of $AC$ and $CZ$ being up at the same time ($0.9^2$).
+which is the probability of $SA$ and $AT$ being up at the same time ($0.9^2$).
 
-In the Equivalence Class where the link $AC$ is failing however ($\mathcal{E}_2$), 
-OSPF and ECMP will produce two equaly-weighted paths $ABEZ$ and $ADFZ$.
+In the Equivalence Class where the link $SA$ is failing however ($\mathcal{E}_2$), 
+OSPF and ECMP will produce two equaly-weighted paths $SBCT$ and $SDET$.
 The probability of this Equivalence Class actually materializing ($P_f$) is $0.053$

### 2023-02-22 paper/4_temporal.tex
[Paper] Updated figures and polish

@@
 In our Fig. \ref{fig:tree} Exploration Tree for example, $\mathcal{E}_2$ has two 
-equally probable path, $ABEZ$ and $ADFZ$.
-We will first do a chain convolution of all the distributions in $ABEZ$ ($l_p(AB)$, 
-$l_q(AB)$, ..., $l_q(EZ)$) and do the same thing for $ADFZ$.
-We will then take the $CDF$ of $\mathcal{L}_{ABEZ}$ and $\mathcal{L}_{ADFZ}$ with our 
+equally probable path, $SBCT$ and $SDET$.
+We will first do a chain convolution of all the distributions in $SBCT$ ($l_p(SB)$, 
+$l_q(SB)$, ..., $l_q(CT)$) and do the same thing for $SDET$.
+We will then take the $CDF$ of $\mathcal{L}_{SBCT}$ and $\mathcal{L}_{SDET}$ with our 
 desired $t$ and average them out to get $P_t$ of $\mathcal{E}_2$.

### 2023-02-22 paper/5_opt.tex
[Paper] updated various things
[Paper] Updated figures and polish

@@
 In our running example, we could see that in Fig. \ref{fig:tree}, $\mathcal{E}_2$ and $\mathcal{E}_4$ share 
-an identical path, $ADFZ$.
-We could therefore memoize $cdf(\mathcal{L}_{ADFZ}, t)$ when we explore $\mathcal{E}_2$ and query the 
+an identical path, $SDET$.
+We could therefore memoize $cdf(\mathcal{L}_{SDET}, t)$ when we explore $\mathcal{E}_2$ and query the 
 cached result when we explore $\mathcal{E}_4$.
@@
 With this optimization, we will do a chain convolution by the amount of path available in the network between 
-the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $ACZ$, $ABEF$, 
-and $ADFZ$).
+the source and destination (which is 3 in the case of our running example in Fig. \ref{fig:ex}: $SAT$, $SBCT$, 
+and $SDET$).
 
@@
 % Since not all convolutions can be calculated analytically, we implemented a numerical convolution 
-% method via mixture distribution, DIRECT \cite{rover2017discrete}, which is able to convolve two 
+% method via mixture distribution, DIRECT \cite{direct}, which is able to convolve two 
 % distribution with a KL divergence error bound.

### 2023-02-22 paper/6_eval.tex
[Paper] updated graph for fattree
[Paper] Updated figures and polish

@@
     \centering
-    \includegraphics[scale=0.5]{optimization}
+    \begin{subfigure}[b]{\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{optimization}
         \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
         \label{fig:opt}
-\end{figure}
-\begin{figure}[h]
+    \end{subfigure}
+    \vfill
+    \begin{subfigure}[b]{\columnwidth}
         \centering
-    \includegraphics[scale=0.5]{optimization_fattree}
-    \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied}
+        \includegraphics[width=\textwidth]{optimization_fattree}
+        \caption{Amount of convolution (in log scale) depending on what optimization strategy is applied. 
+            Fattree 12 upwards is extrapolatied from the number of convolution}
         \label{fig:opt_fat}
+    \end{subfigure}
+    \caption{test}
 \end{figure}
@@
 
+% \begin{figure}[h]
+%     \centering
+%     \includegraphics[scale=0.5]{ec}
+%     \caption{Amount of temporal EC compared to functional EC, label represents ratio}
+%     \label{fig:ec}
+% \end{figure}
+
 \begin{figure}[h]
     \centering
-    \includegraphics[scale=0.5]{ec}
+    \begin{subfigure}[b]{\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ec}
+        \caption{Functional EC and Temporal EC for WAN topologies}
+        \label{fig:y equals x}
+    \end{subfigure}
+    \vfill
+    \begin{subfigure}[b]{\columnwidth}
+        \centering
+        \includegraphics[width=\textwidth]{ec_fattree}
+        \caption{Functional EC and Temporal EC for DC topologies}
+        \label{fig:three sin x}
+    \end{subfigure}
        \caption{Amount of temporal EC compared to functional EC, label represents ratio}
@@
 
-The first one is the \textbf{amount of possible paths}.
+The first one is the \textbf{amount of possible paths} (Fig. \ref{fig:opt_analysis_amt}).
 If the amount of possible paths between a src-dst pair in a given topology is small, then the functional 
 verification step would produce fewer equivalence classes and end early.
+In our simplified example in Fig. \label{fig:opt_analysis_amt}, although the network has multiple interconnected
+nodes, S and T has only one possible path.
 While most of our evaluated topology had a lot of possible paths, one exception to this is LatNet, which 
-only have 3 possible paths.
+only have 3 possible paths in our evaluation.
 
-The second is whether the src-dst pair in a given topology would \textbf{produce paths that are "symmetric"}.
+The second is whether the src-dst pair in a given topology would \textbf{produce paths that are "symmetric"}
+(Fig \ref{fig:opt_analysis_symmetry}).
 By symmetric, we mean that the routing and load balancing method produce a lot of identical equivalence class 
 and / or equivalence class that share the same path.
+In our simplified example in Fig. \ref{fig:opt_analysis_symmetry}, the network has five 2-hop paths, which will give a 
+lot of equal-cost path combination under ECMP.
 As our explanation about Fat Tree 10 suggests, larger Fat Tree has a lot of path with the same length.
@@
 
+\begin{figure}
+    \centering
+    \begin{subfigure}[b]{0.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{eval_ex_amt}
+        \caption{S-T only has one path}
+        \label{fig:opt_analysis_amt}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}[b]{0.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{eval_ex_symmetry}
+        \caption{S-T has five 2-hop paths}
+        \label{fig:opt_analysis_symmetry}
+    \end{subfigure}
+       \caption{Illustration of topological property}
+       \label{fig:opt_analysis}
+\end{figure}
+
 %TODO: path enumeration?

### 2023-02-23 src/Direct.jl
[Code] polish some code

@@ -59,14 +59,17 @@ function direct(d1::UnivariateDistribution, d2::UnivariateDistribution, delta::F
     end
     
     # Determine bin half-width from delta
-    # Problem: if d2 has small probability value, then this will be infinite loop?
+    # Problem: infinite loop if sym_kldivergence_shift = 0.0 as step gets larger
+    # e.g. Empirical -> logpdf = 0 everywhere
     # e.g. Normal(1, 0.01) instead of Normal()
     step = sqrt(delta)
     while sym_kldivergence_shift(d2, step) < delta
+        # step == Inf && break # @diwangs: hack
         step = step * 2
     end
     # This line took the longest to finish
     step = find_zero(x -> sym_kldivergence_shift(d2, x) - delta, (0.0, step), A42())
+    # step = find_zero(x -> sym_kldivergence_shift(d2, x) - delta, 0.0)
 
     # Determine grid range from epsilon
     mini = -1.0

### 2023-02-23 src/Exploration.jl

@@ -1,37 +1,55 @@
-mutable struct MetaPath
-    path::Vector{Symbol}
-    dependencies::Vector{Symbol}        # TODO: technically should be a set
-end
-
-function explore(tg::AbstractGraph, src::Symbol, dst::Symbol, threshold::Real)::Float64
+function explore(tg::AbstractGraph, src::Symbol, dst::Symbol, threshold::Real, opt::String)::Float64
     start = now()
 
     # If src and dst are the same node
     src == dst && return 1.0
 
     # Create state tree and exploration queue
-    st = StateTree()
-    e = PriorityQueue(Base.Order.Reverse)
-
+    e = PriorityQueue{State, Float64}(Base.Order.Reverse)
     # Create the perfect state as the initial state
     perfect = State([], [])
-    st[:perfect] = perfect
-    st[:perfect].p_state = get_p_state(tg, st, :perfect)
-    e[:perfect] = st[:perfect].p_state
-    
+    e[perfect] = get_p_state(tg, perfect)
     # Prepare for functional state exploration
     tgcopy = deepcopy(tg)
-    p_explored = 0.0
 
+    # Probabilities
+    p_explored = 0.0
+    p_property = 0.0
+    # Cache for consolidation and memoization
+    p_paths_temporals = Dict{Set{Vector{Symbol}}, Float64}()
+    p_path_temporals = Dict{Vector{Symbol}, Float64}()
+
+    # Logging variable
+    # Equivalence classes
+    cnt_fec = 0
+    cnt_fec_unreachable = 0
+    cnt_tec = 0
+    # Path
+    cnt_path = 0
+    # Convolution
+    cnt_conv::UInt = 0
+    # cnt_nconv::UInt = 0
+    # Analytic: in an optimized run, how much EC, path, and convolution is actually pruned?
+    # The value of these variables might not make sense for consolidation-only and baseline version
+    cnt_analytic_tec = Dict{Set{Vector{Symbol}}, Int}() # Does not include unreachable state
+    cnt_analytic_path = Dict{Vector{Symbol}, Int}() # Not covered by consolidation
+    
+    # Duration
+    dur_functional = zero(now())
+    dur_temporal = zero(now())
+    # dur_path = Dict{Vector{Symbol}, Millisecond}()
+    
+    # cnt_links = Dict{Tuple{Symbol, Symbol}, Int}()
     # Explore functional state
-    cnt = 0
[…]

### 2023-02-23 src/StateTrees.jl

@@ -1,20 +1,20 @@
-mutable struct State 
+struct State 
     # Pre-exploration
     force_enabled::Vector{Tuple{Symbol, Symbol}}
     disabled::Vector{Tuple{Symbol, Symbol}}
     spur_node_idx::UInt # this state shares the same root_path with its parent up until this index (1 == only shares src)
     
-    # Post-exploration
-    converged_paths::Vector{Vector{Symbol}}
-    p_state::Float64                            # The probability of the network arriving at this state (hot edges of this state and its predecesors)
-    p_paths_functional::Float64                 # The probability of _new_ hot edges being up (converged_paths - (force_enabled + disabled))
-                                                # 1.0 == no new hot edges, 0.0 == not reachable
+    # # Post-exploration
+    # converged_paths::Vector{Vector{Symbol}}
+    # p_state::Float64                            # The probability of the network arriving at this state (hot edges of this state and its predecesors)
+    # p_paths_functional::Float64                 # The probability of _new_ hot edges being up (converged_paths - (force_enabled + disabled))
+    #                                             # 1.0 == no new hot edges, 0.0 == not reachable
 
-    # Post-grouping
-    p_paths_temporal::Float64                   # The probability of converged_paths transmitting packets below a threshold
+    # # Post-grouping
+    # p_paths_temporal::Float64                   # The probability of converged_paths transmitting packets below a threshold
 
-    State(force_enabled, disabled, spur_node_idx::Int) = new(force_enabled, disabled, spur_node_idx, [], 0.0, 0.0, 0.0)
-    State(force_enabled, disabled) = new(force_enabled, disabled, 1, [], 0.0, 0.0, 0.0)
+    State(force_enabled, disabled, spur_node_idx::Int) = new(force_enabled, disabled, spur_node_idx)
+    State(force_enabled, disabled) = new(force_enabled, disabled, 1)
 end
 
 const StateTree = typeof(MetaGraph(DiGraph(), VertexData = State))
@@ -64,4 +64,20 @@ function get_p_state(tg::TopologyGraph, st::StateTree, l::Symbol)::Float64
     p_state *= get_p_state(tg, st, parent)
 
     return p_state
+end
+
+function get_p_state(tg::TopologyGraph, s::State)::Float64
+    p_state::Float64 = 1.0
[…]

### 2023-02-23 src/Tempus.jl

@@ -1,13 +1,20 @@
 module Tempus
 
+using Logging
 import JSON
+import CSV
 using DataStructures
 using Random
-using Distributions
-import Distributions: cdf, pdf, logpdf, minimum, maximum, quantile, convolve
 using Dates
 
+using Distributions
+import Distributions: cdf, pdf, logpdf, minimum, maximum, quantile, convolve
+using StatsBase
+using EmpiricalDistributions
+import EmpiricalDistributions: _linear_interpol, _ratio
+
 using Graphs
+import Graphs: weights
 using MetaGraphsNext
 
 using QuadGK # Numerical integration
@@ -18,25 +25,36 @@ include("TopologyGraphs.jl")
 include("StateTrees.jl")
 include("Exploration.jl")
 
+function _linear_interpol(x_lo::Real, x_hi::Real, y_lo::Real, y_hi::Real, x::Real)
+    T = promote_type(typeof(y_lo), typeof(y_hi), typeof(x))
+    w_hi = T(_ratio(x - x_lo, x_hi - x_lo))
+    w_lo = one(w_hi) - w_hi
+    y = y_lo * w_lo + y_hi * w_hi
+    # Original assertion might fail in the case of float rounding error
+    # @assert y_lo <= y <= y_hi 
+    !(y_lo < y) && @assert isapprox(y_lo, y)
+    !(y < y_hi) && @assert isapprox(y, y_hi)
+    y
+end
+
 function main()
-    isempty(ARGS) && (println("Please provide a file name as an input"); return 1)
+    size(ARGS)[1] < 1 && (println("Please provide a file name as an input (and optionally optimization level)"); return 1)
 
     json::String = open("artifacts/" * ARGS[1]) do file
         read(file, String)
     end
     config::Dict{String, Any} = JSON.parse(json)
+    if size(ARGS)[1] == 1
+        opt = "opti"
+    else
+        opt = ARGS[2]
+    end
 
     tg::TopologyGraph = TopologyGraph(config["routers"], config["links"])
-    prob::Float64 = explore(tg, Symbol(config["intent"]["src"]), Symbol(config["intent"]["dst"]), config["intent"]["threshold"])
-    println(prob)
+    # @info ne(tg) / 2
 
-    # start = now()
-    # x = LogNormal()
-    # y = Exponential()
-    # # z = convolve(convolve(x, x), y)
-    # z = convolve(x, x)
-    # println(now() - start)
-    # println(z)
[…]

### 2023-02-23 src/TopologyGraphs.jl

@@ -17,7 +17,6 @@ function TopologyGraph(routers::Vector{Any}, links::Vector{Any})::TopologyGraph
             continue
         end
 
-        # println(link)
         linkdist = gendist(link["delayModel"])
         uvdist = gendist(getdelaymodel(routers, link["u"], link["v"]))
         tg[Symbol(link["u"]), Symbol(link["v"])] = (link["failProb"], linkdist, uvdist, link["w_uv"])
@@ -35,12 +34,20 @@ function getdelaymodel(routers::Vector{Any}, u, v::String)::Dict{String, Any}
 end
 
 function gendist(delaymodel::Dict{String, Any})::Distribution
-    args = Tuple(delaymodel["args"])
-    rawdist = eval(Meta.parse(delaymodel["delayType"] * string(args)))
-    # Truncate to disallow negative delay
-    if cdf(rawdist, 0.0) <= 0.0 # TODO: change this threshold
-        return rawdist 
-    else
-        return truncated(rawdist; lower=0.0)
+    if delaymodel["delayType"] == "Empiric"
+        samples = CSV.File("artifacts/" * delaymodel["args"][1])["sample"]
+        d = UvBinnedDist(fit(Histogram, samples))
+        return d
+    else 
+        args = Tuple(delaymodel["args"])
+        rawdist = eval(Meta.parse(delaymodel["delayType"] * string(args)))
+        # Truncate to disallow negative delay
+        if cdf(rawdist, 0.0) > 0.0 
+            rawdist = truncated(rawdist; lower=0.0)
+        end
+    
+        # Empiric
+        dist = UvBinnedDist(fit(Histogram, rand(rawdist, 1000000)))
+        return dist
     end
 end

### 2023-02-23 src/netdiceparser/FatTreeGenerator.jl

@@ -1,0 +1,115 @@
+module FatTreeGenerator
+
+using Printf
+using JSON
+
+function main() 
+    k = parse(Int, ARGS[1]) # number of pods
+
+    tempusConfig = Dict()
+
+    # Link 
+    links = []
+    for pod in 1:k
+        for idx in 1:(k/2) # For every aggregation switch...
+            for dst in 1:(k/2)
+                # Connect to edge switch in the same pods in a bipartite manner
+                push!(links, Dict(
+                    "u" => @sprintf("aggr_%d_%d", pod, idx),
+                    "v" => @sprintf("edge_%d_%d", pod, dst),
+                    "w_uv" => 1,
+                    "w_vu" => 1,
+                    "failProb" => 0.001,
+                    "delayModel" => Dict(
+                        "isEmpiric" => true,
+                        "delayType" => "Gamma",
+                        "args" => []
+                    )
+                ))
+
+                # Aggr is connected to k/2 core
+                push!(links, Dict(
+                    "u" => @sprintf("aggr_%d_%d", pod, idx), 
+                    "v" => @sprintf("core_%d", (idx - 1) * (k/2) + dst),
+                    "w_uv" => 1,
+                    "w_vu" => 1,
+                    "failProb" => 0.001,
+                    "delayModel" => Dict(
+                        "isEmpiric" => true,
+                        "delayType" => "Gamma",
+                        "args" => []
+                    )
+                ))
+            end
+        end
+    end
+
+    # Routers
+    # k pods
+    #   k/2 edge routers -> edge_1_1: edge, pod 1, index 1
+    #   k/2 aggr routers -> aggr_1_2: aggr, pod 1, index 2
+    # (k/2)^2 core routers -> core_1, core_2, ...
+    routers = []
+    for pod in 1:k
+        for idx in 1:(k/2)
+            name = @sprintf("edge_%d_%d", pod, idx)
+            push!(routers, Dict(
+                "name" => name,
+                "failProb" => 0.0,
+                "outQdelayModel" => [Dict(
+                    "to" => link["u"] == name ? link["v"] : link["u"],
+                    "delayModel" => Dict(
+                        "isEmpiric" => true,
+                        "delayType" => "Empiric",
+                        "args" => ["queue_measurement/dctcp_us.csv"]
+                    )
+                ) for link in filter(x -> x["u"] == name || x["v"] == name, links)]
+            ))
+
[…]

### 2023-02-23 src/netdiceparser/NetDiceTopologyParser.jl

@@ -6,7 +6,7 @@ using Random
 function main()
     tempusConfig = Dict()
 
-    open("artifacts/AttMpls.in") do f 
+    open("artifacts/" * ARGS[1]) do f 
         
         r = readline(f) # Read the first line containing the number of nodes
         
@@ -23,9 +23,11 @@ function main()
                 "w_vu" => parse(Int, array[4]),
                 "failProb" => 0.001,
                 "delayModel" => rand(Bool) ? Dict(
+                    "isEmpiric" => true,
                     "delayType" => "Gamma",
                     "args" => []
                 ) : Dict(
+                    "isEmpiric" => true,
                     "delayType" => "Chisq",
                     "args" => [3]
                 )
@@ -40,12 +42,10 @@ function main()
                 "failProb" => 0.0,
                 "outQdelayModel" => [Dict(
                     "to" => link["u"] == name ? link["v"] : link["u"],
-                    "delayModel" => rand(Bool) ? Dict(
-                        "delayType" => "Gamma",
-                        "args" => []
-                    ) : Dict(
-                        "delayType" => "Chisq",
-                        "args" => [3]
+                    "delayModel" => Dict(
+                        "isEmpiric" => true,
+                        "delayType" => "Empiric",
+                        "args" => ["queue_measurement/dctcp_us.csv"]
                     )
                 ) for link in filter(x -> x["u"] == name || x["v"] == name, links)]
             ))
@@ -54,8 +54,8 @@ function main()
         tempusConfig["routers"] = routers
         tempusConfig["links"] = links
         tempusConfig["intent"] = Dict(
-            "src" => "30",
-            "dst" => "46",
+            "src" => ARGS[3],
+            "dst" => ARGS[4],
             "threshold" => 1000.0
         )
 
@@ -63,7 +63,7 @@ function main()
         # println(s)
     end
 
-    open("artifacts/AttMplsGammaChisq.json", "w") do f
+    open("artifacts/" * ARGS[2], "w") do f
         write(f, JSON.json(tempusConfig))
     end
 end
